---
title: "cDNA processing"
author: "Max Trauernicht"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: 
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---

# knitr document van Steensel lab

# Introduction
I previously processed the raw sequencing data, optimized the barcode clustering, quantified the pDNA data and normalized the cDNA data. In this script, I want to have a detailed look at the cDNA data from a general perspective.

## Description of Data
How to make a good rendering table: 
```{r table1, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- "
| column1	|	column2	|	column3	|	
|----|----|----|
|1	|	2	|	3	|	
|a	|	b	|	c	|	
"
cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
```

# Data processing
## Path, Libraries, Parameters and Useful Functions
```{r setup, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
StartTime <-Sys.time()

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8) 
# libraries:
library(RColorBrewer)
library(ggplot2)
library(dplyr)
library(maditr)
library(tibble)
library(pheatmap)
library(ggpubr)
library(ggbeeswarm)
library(ggforce)
library(viridis)
library(plyr)
library(cowplot)
library(gridExtra)
```

### Custom functions
Functions used thoughout this script.
```{r}
SetFileName <- function(filename, initials) {
  # Set filename with extension and initials to make filename with date integrated.
  filename <- substitute(filename)
  initials <- substitute(initials)
  filename <- paste0(initials, Date, filename)
  filename
}


cbind.fill <- function(...){
    nm <- list(...) 
    nm <- lapply(nm, as.matrix)
    n <- max(sapply(nm, nrow)) 
    do.call(cbind, lapply(nm, function (x) 
        rbind(x, matrix(, n-nrow(x), ncol(x))))) 
}
```


## Data import
```{r}
# Import processed bc counts from the preprocessing step
cDNA_df <- read.csv("/DATA/usr/m.trauernicht/projects/tf_activity_reporter/data/SuRE_TF_1/results/mt20200715_reporter_activity_filt.csv", header = T)
cDNA_df$background <- as.character(cDNA_df$background)
cDNA_df <- na.omit(cDNA_df)
cDNA_df$TF <- revalue(cDNA_df$TF, c("Tcfcp2l1" = "Tfcp2l1", "Tcfcp2l1_neg" = "Tfcp2l1_neg"))
cDNA_df$reporter_id <- gsub("Tcfcp2l1", "Tfcp2l1", cDNA_df$reporter_id)
```

# Analysis

## First insights into data distribution - reporter activity distribution plots
```{r data_density, out.width= "100%", fig.align= "center", echo=FALSE, warning= FALSE}
## reporter activity distribution over all experiments
colors <- c("#1B998B", "#2D3047", "#FF9B71")

## reporter activity distribution for each experiment individually
ggplot(data = cDNA_df, aes(x = log_activity)) + 
  geom_density(fill = "#3D9F83") + xlab("log2(cDNA/pDNA)") + 
  labs(title = "reporter activity distribution") +
  xlim(-5,5) + theme_classic() + facet_wrap(vars(condition))

fill <- c("2i_neg_LIF" = "#CACDD3", 
            "2i_pos_LIF" = "#CACDD3",
            "LIF_pos_CH" = "#CACDD3",
            "LIF_pos_PD" = "#CACDD3",
            "N2B27" = "#CACDD3",
            "NPC" = "#1B998B",
            "vitA" = "#CACDD3"
            )

ggplot(data = cDNA_df, aes(x = log_activity, fill = condition, alpha = 0.1)) + 
  geom_density() + xlab("log2(cDNA/pDNA)") + 
  labs(title = "reporter activity distribution: NPC vs. mESCs") +
  scale_fill_manual(values = fill)+
  xlim(0,5) + theme_classic() +
  theme(text = element_text(size = 12))

## reporter activity distribution for each experiment individually
ggplot(data = cDNA_df, aes(x = log_activity, color = condition)) + 
  stat_ecdf() + xlab("log2(cDNA/pDNA)") + 
  labs(title = "reporter activity distribution") + ylab("fraction") +
  xlim(-5,5) + theme_classic() +
  theme(text = element_text(size = 12))


## reporter activity distribution - highlight negative motif controls
ggplot(cDNA_df[cDNA_df$hPGK == "No",], aes(x = log_activity, fill = neg_ctrls, alpha = 0.4)) + 
  geom_density() + xlab("log2(cDNA/pDNA)") + 
  labs(title = "reporter activity distribution - negative ctrls") +
  scale_fill_manual(values = colors) +
  xlim(-5,5) + theme_classic()+
  theme(text = element_text(size = 12))


## reporter activity distribution - highlight positive hPGK control
ggplot(cDNA_df, aes(x = log_activity, fill = hPGK, alpha = 0.4)) + 
  geom_density() + xlab("log2(cDNA/pDNA)") + 
  labs(title = "reporter activity distribution - hPGK") +
  scale_fill_manual(values = colors) +
  xlim(-5,5) + theme_classic()

## reporter activity distribution - highlight Miguels enhancer control
ggplot(cDNA_df, aes(x = log_activity, fill = native_enhancer, alpha = 0.4)) + 
  geom_density() + xlab("log2(cDNA/pDNA)") + 
  labs(title = "reporter activity distribution - native enhancers") +
  scale_fill_manual(values = colors) +
  xlim(0,5) + theme_classic()+ facet_wrap(vars(condition))


## reporter activity distribution - highlight different promoters
ggplot(cDNA_df, aes(x = log_activity, fill = promoter, alpha = 0.4)) + 
  geom_density() + xlab("log2(cDNA/pDNA)") + 
  labs(title = "reporter activity distribution - promoters") +
  scale_fill_brewer("Dark2") +
  xlim(-5,5) + theme_classic()

## reporter activity distribution - highlight random promoters
ggplot(cDNA_df, aes(x = log_activity, fill = rand_promoter, alpha = 0.4)) + 
  geom_density() + xlab("log2(cDNA/pDNA)") + 
  labs(title = "reporter activity distribution - random promoter") +
  scale_fill_manual(values = colors) +
  xlim(-5,5) + theme_classic()

## reporter activity distribution - highlight different distances
ggplot(cDNA_df[cDNA_df$distance != "",], 
       aes(x = log_activity, fill = distance, alpha = 0.4)) + 
  geom_density() + xlab("log2(cDNA/pDNA)") + 
  labs(title = "reporter activity distribution - distances") +
  scale_fill_manual(values = colors) +
  xlim(-5,5) + theme_classic()

## reporter activity distribution - highlight different spacings
ggplot(cDNA_df[cDNA_df$spacing != "",], aes(x = log_activity, fill = spacing, alpha = 0.4)) + 
  geom_density() + xlab("log2(cDNA/pDNA)") + 
  labs(title = "reporter activity distribution - spacings") +
  scale_fill_manual(values = colors) +
  xlim(-5,5) + theme_classic()

## reporter activity distribution - highlight different backgrounds
ggplot(cDNA_df[cDNA_df$background != "0",], aes(x = log_activity, fill = background, alpha = 0.4)) + 
  geom_density() + xlab("log2(cDNA/pDNA)") + 
  labs(title = "reporter activity distribution - backgrounds") +
  scale_fill_manual(values = colors) +
  xlim(-5,5) + theme_classic()

## normalized bc-counts for each TF - comparing to original pDNA counts
ggplot(cDNA_df[cDNA_df$native_enhancer == "No" &
               cDNA_df$neg_ctrls  == "No" &
               cDNA_df$hPGK == "No" &
               cDNA_df$rand_promoter == "No",]) + 
  geom_density(aes(x = bc_counts_rpm, fill = "#1B998B", alpha = 0.4)) + 
  geom_density(aes(x = pDNA_counts_rpm, fill = "#2D3047", alpha = 0.4)) + 
  xlab("bc-counts per million") + 
  labs(title = "bc-count distribution cDNA (green) vs. pDNA (blue)") +
  scale_fill_manual(values = colors) +
  xlim(0,200) + theme_classic() +
  facet_wrap(~TF)

## correlation bc-counts pDNA vs. cDNA
sp <- ggscatter(cDNA_df[cDNA_df$native_enhancer == "No" &
               cDNA_df$neg_ctrls  == "No" &
               cDNA_df$hPGK == "No" &
               cDNA_df$rand_promoter == "No",],
               x = "pDNA_counts_rpm", y = "bc_counts_rpm",
               size = 0.1, alpha = 0.3, color = "#2D3047", 
   add = "reg.line",
   add.params = list(color = "blue", fill = "lightgray"), title = "normalized barcode counts cDNA vs. pDNA",
   conf.int = TRUE, ylab = "cDNA", xlab = "pDNA")
sp + stat_cor(method = "pearson", label.x = 150, label.y = 2) + 
  geom_abline(linetype = "dashed") + facet_wrap(~TF) + xlim(0,200) + ylim(0,200)


## correlation bc-counts pDNA vs. reporter activity
sp <- ggscatter(cDNA_df[cDNA_df$native_enhancer == "No" &
                    cDNA_df$neg_ctrls  == "No" &
                    cDNA_df$hPGK == "No" &
                    cDNA_df$rand_promoter == "No",],
          size = 0.75, color = "condition", alpha = 0.1,
          y = "log_activity", x = "pDNA_counts_rpm",
          title = "reporter pDNA counts vs log2 reporter activity",
          ylab = "log2 reporter activity",
          xlab = "pDNA counts per million",
          facet.by = "TF")
sp + xlim(0,200) + ylim(-5,5) + scale_color_manual(values = fill) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10), 
          axis.text.y = element_text(size = 10)) +
  theme(text = element_text(size = 12))

## let's zoom into a model TF with high noise at low pDNA: Gbx2
sp <- ggscatter(cDNA_df[cDNA_df$TF == "Gbx2" &
                    cDNA_df$neg_ctrls  == "No" &
                    cDNA_df$hPGK == "No" &
                    cDNA_df$rand_promoter == "No",],
          size = 0.1, color = "#2D3047",
          y = "log_activity", x = "pDNA_counts_rpm",
          title = "Gbx2: reporter pDNA counts vs log2 reporter activity",
          ylab = "log2 reporter activity",
          xlab = "pDNA counts per million",
          facet.by = "condition")
sp + xlim(0,200) + ylim(-5,5)

## let's zoom Stat3 because it's nice
sp <- ggscatter(cDNA_df[cDNA_df$TF == "Stat3" &
                    cDNA_df$neg_ctrls  == "No" &
                    cDNA_df$hPGK == "No" &
                    cDNA_df$rand_promoter == "No",],
          size = 0.1, color = "#2D3047",
          y = "log_activity", x = "pDNA_counts_rpm",
          title = "Stat3: reporter pDNA counts vs log2 reporter activity",
          ylab = "log2 reporter activity",
          xlab = "pDNA counts per million",
          facet.by = "condition")
sp + xlim(0,200) + ylim(-5,5)
```




## Heat map - display mean log2-activity for each TF in each condition
```{r tf_activity_heatmap, out.width= "100%", fig.align= "center", echo=FALSE, warning= FALSE}
## Prepare dataframe
# Caculate mean TF activity per condition
cDNA_df$tf_activity <- ave(cDNA_df$reporter_activity, cDNA_df$condition, cDNA_df$TF, FUN = function(x) mean(x))
cDNA_df$tf_activity <- log2(cDNA_df$tf_activity)
cDNA_df$group  <- "Other"
neural <- c("Gli1", "Homez", "Irx3", "Neurog2", "Otx1", "Pax6", "Rbpj")
cDNA_df$group[grep(paste(neural, collapse = "|"), cDNA_df$TF)] <- "Neural induction"
ES <- c("Esrrb", "Gbx2", "Klf4", "Smad4", "Stat3", "Tcf7l2", "Tfcp2l1", "Zfp42", "Zfx")
cDNA_df$group[grep(paste(ES, collapse = "|"), cDNA_df$TF)] <- "Stemness"
pos <- c("Sp1", "Nfya", "Random")
cDNA_df$group[grep(paste(pos, collapse = "|"), cDNA_df$TF)] <- "Control"
  
tf_activity_heatmap <- cDNA_df[cDNA_df$neg_ctrls == "No" & cDNA_df$hPGK == "No" &
                               cDNA_df$native_enhancer == "No" & cDNA_df$rand_promoter == "No",] %>% 
  select(TF, condition, tf_activity) %>% unique()
tf_activity_heatmap <- dcast(tf_activity_heatmap, TF ~ condition, value.var="tf_activity")
group <- cDNA_df %>% select(TF, group) %>% unique()
tf_activity_heatmap <- merge(tf_activity_heatmap, group)

tf_activity_heatmap <- tf_activity_heatmap %>% 
  remove_rownames %>% column_to_rownames(var="TF")
# Keeping the scale in the pheatmap function
myBreaks1 <- seq(0,3.5,0.035)


# pheatmap function
pheatmap(as.matrix(t(tf_activity_heatmap[,-8])),
         color = colorRampPalette(brewer.pal(n = 7, name = "Greys"))(100),
         breaks = myBreaks1, border_color = "#565656", 
         cellwidth = 12, cellheight = 10,
         angle_col = 90)

# selective representation
tf_activity_heatmap <- tf_activity_heatmap[,c(6,2,1,3,4,5,7,8)]
for (i in unique(tf_activity_heatmap$group)) {
  p <- pheatmap(as.matrix(t(tf_activity_heatmap[tf_activity_heatmap$group == i,-8])),
         color = colorRampPalette(brewer.pal(n = 7, name = "Greys"))(100),
         breaks = myBreaks1, border_color = "black", 
         cellwidth = 12, cellheight = 10, cluster_rows = F,
         angle_col = 90)
  print(p) 
}

```
## Only ~10 of the 26 TFs in the library show promising activity at the first glimpse

## Heatmap for native enhancers
```{r}
## Prepare dataframe
# Caculate mean TF activity per condition
tf_activity_heatmap <- cDNA_df[cDNA_df$native_enhancer == "Yes",] %>% 
  select(TF, condition, tf_activity) %>% unique()
tf_activity_heatmap <- dcast(tf_activity_heatmap, condition ~ TF, value.var="tf_activity")

tf_activity_heatmap <- tf_activity_heatmap %>% 
  remove_rownames %>% column_to_rownames(var="condition")

comb_activity <- data.frame("pos" = colnames(tf_activity_heatmap), 
                             "activity" = colSums(tf_activity_heatmap)) 

ggplot(comb_activity, aes(x = pos, y = activity)) +
  geom_bar(stat = "identity", aes(fill = activity)) + theme_classic() +
    theme(axis.title.y = element_blank(), axis.text.y = element_blank(), 
          axis.ticks.y = element_blank(), axis.ticks.x = element_blank(),
          axis.text.x = element_blank(), 
          axis.title.x = element_blank(),
          legend.position = "none") +
    scale_fill_distiller(name = "activity", palette = "Reds", direction = 1)



# Keeping the scale in the pheatmap function
myBreaks1 <- seq(-3,3,0.06)


# pheatmap function
pheatmap(as.matrix(tf_activity_heatmap),
         color = colorRampPalette(rev(brewer.pal(n = 7, name = "RdBu")))(100),
         breaks = myBreaks1, border_color = "#565656", 
         cellwidth = 10, cellheight = 10, cluster_rows = F, cluster_cols = F)

```



## Heatmap per TF - comparing design activities mutated vs. non-mutated
```{r heatmap_4, out.width= "100%", fig.align= "center", echo=FALSE, warning= FALSE}
## Prepare dataframe
# Caculate mean TF activity per condition
cDNA_df$reporter <- gsub("_[0-9]{1}$", "", cDNA_df$reporter_id)
cDNA_df$tf_design_activity <- ave(cDNA_df$log_activity, 
                                  cDNA_df$condition, 
                                  cDNA_df$reporter,
                                  FUN = function(x) mean(x))


tf_activity_heatmap <- cDNA_df[cDNA_df$native_enhancer == "No" & 
                                 cDNA_df$rand_promoter == "No" &
                                 cDNA_df$hPGK == "No",] %>% 
  select(reporter, condition, tf_design_activity) %>% unique()

tf_activity_heatmap$tf_design_activity[!is.finite(tf_activity_heatmap$tf_design_activity)] <- 0
tf_activity_heatmap <- dcast(tf_activity_heatmap, reporter ~ condition,
                             value.var="tf_design_activity")
tf_activity_heatmap$TF <- gsub("(^.*?)_[0-9]{1,2}bp.*", "\\1", tf_activity_heatmap$reporter)
tf_activity_heatmap$TF <- gsub("_neg$", "", tf_activity_heatmap$TF)
tf_activity_heatmap <- tf_activity_heatmap %>% 
  remove_rownames %>% column_to_rownames(var="reporter")
# Keeping the scale in the pheatmap function
myBreaks1 <- seq(-3,3,0.06)

# Manually annotate heatmaps
annotation_col <- data.frame("TF" = cDNA_df$reporter[cDNA_df$native_enhancer == "No" & 
                                 cDNA_df$rand_promoter == "No" &
                                 cDNA_df$hPGK == "No"], 
                             "neg_ctrls" = cDNA_df$neg_ctrls[cDNA_df$native_enhancer == "No" & 
                                 cDNA_df$rand_promoter == "No" &
                                 cDNA_df$hPGK == "No"]) %>%
  unique() %>%
  remove_rownames %>% column_to_rownames(var="TF")

annotation_colors = list(
    neg_ctrls = c(No ="#1B998B", Yes = "#2D3047"))


# pheatmap function
for(i in unique(tf_activity_heatmap$TF)){
  pheatmap(as.matrix(t(tf_activity_heatmap[tf_activity_heatmap$TF == i, -8])),
         color = colorRampPalette(rev(brewer.pal(n = 7, name = "RdBu")))(100),
         breaks = myBreaks1, border_color = "#565656", 
         cellwidth = 10, cellheight = 10,
         annotation_col = annotation_col, 
         annotation_colors = annotation_colors)
}
```



## Heatmap per TF - only WT TF activities
```{r heatmap_5, out.width= "100%", fig.align= "center", echo=FALSE, warning= FALSE}
## Prepare dataframe
# Caculate mean TF activity per condition
tf_activity_heatmap <- cDNA_df[cDNA_df$native_enhancer == "No" & 
                                 cDNA_df$hPGK == "No" &
                                 cDNA_df$neg_ctrls == "No",] %>% 
  select(reporter, condition, tf_design_activity) %>% unique()

tf_activity_heatmap$tf_design_activity[!is.finite(tf_activity_heatmap$tf_design_activity)] <- 0
tf_activity_heatmap <- dcast(tf_activity_heatmap, reporter ~ condition,
                             value.var="tf_design_activity")
tf_activity_heatmap$TF <- gsub("(^.*?)_[0-9]{1,2}bp.*", "\\1", tf_activity_heatmap$reporter)
tf_activity_heatmap <- tf_activity_heatmap %>% 
  remove_rownames %>% column_to_rownames(var="reporter")
# Keeping the scale in the pheatmap function
myBreaks1 <- seq(-3,3,0.06)

# pheatmap function
for(i in unique(tf_activity_heatmap$TF)){
  pheatmap(as.matrix(tf_activity_heatmap[tf_activity_heatmap$TF == i,-8]),
         color = colorRampPalette(rev(brewer.pal(n = 7, name = "RdBu")))(100),
         breaks = myBreaks1, border_color = "#565656", 
         cellwidth = 30, cellheight = 10)
}
```




## Compute activity changes relative to their negative controls
```{r, out.width= "100%", fig.align= "center", echo=FALSE, warning= FALSE}
## TF activity relative to negative controls 
cDNA_df_ctrl <- cDNA_df[cDNA_df$native_enhancer == "No" &
                          cDNA_df$hPGK == "No" & cDNA_df$TF != "Random1" &
                          cDNA_df$TF != "Random2" & cDNA_df$TF != "Random3",] %>% 
  mutate(reporter_2 = gsub("_neg", "", reporter_id)) %>%
  select(reporter_2, condition, reporter_activity, neg_ctrls) %>% 
  mutate(reporter_activity = ave(reporter_activity, reporter_2, condition,
                                 neg_ctrls, FUN = function(x) mean(x))) %>%
  unique()
  

## Divide TF activity by negative control activity for each reporter
for (i in unique(cDNA_df_ctrl$reporter_2)) {
  for (j in unique(cDNA_df_ctrl$condition)) {
    if (length(cDNA_df_ctrl$reporter_activity[cDNA_df_ctrl$reporter_2 == i & 
                                     cDNA_df_ctrl$neg_ctrls == "Yes" &
                                     cDNA_df_ctrl$condition == j]) != 0 &
       length(cDNA_df_ctrl$reporter_activity[cDNA_df_ctrl$reporter_2 == i & 
                                    cDNA_df_ctrl$neg_ctrls == "No" &
                                     cDNA_df_ctrl$condition == j]) != 0) { 
        
    cDNA_df_ctrl$reporter_activity_relative[cDNA_df_ctrl$reporter_2 == i & 
                                     cDNA_df_ctrl$condition == j] <- 
      
     cDNA_df_ctrl$reporter_activity[cDNA_df_ctrl$reporter_2 == i & 
                             cDNA_df_ctrl$neg_ctrls == "No" & 
                              cDNA_df_ctrl$condition == j] /
      
      cDNA_df_ctrl$reporter_activity[cDNA_df_ctrl$reporter_2 == i & 
                              cDNA_df_ctrl$neg_ctrls == "Yes" & 
                               cDNA_df_ctrl$condition == j]
   }
    else {
     cDNA_df_ctrl$reporter_activity_relative[cDNA_df_ctrl$reporter_2 == i & 
                                       cDNA_df_ctrl$condition == j] <- 0
    }
  }
}

## Log2 transformation
cDNA_df_ctrl <- cDNA_df_ctrl %>% 
  mutate(log_reporter_activity_relative = log2(reporter_activity_relative))
  
cDNA_df_ctrl$log_reporter_activity_relative[!is.finite(cDNA_df_ctrl$log_reporter_activity_relative)] <- 0

cDNA_df_ctrl$TF <- gsub("(^.*?)_[0-9]{1,2}bp.*", "\\1", cDNA_df_ctrl$reporter_2)

## Generate heatmaps with relative log2 activities 
# Caculate mean TF activity per condition
tf_activity_heatmap <- cDNA_df_ctrl[cDNA_df_ctrl$neg_ctrls == "No",] %>% 
  select(reporter_2, TF, condition, log_reporter_activity_relative) %>% unique()
tf_activity_heatmap2 <- tf_activity_heatmap %>% select(-reporter_2) %>%
  mutate(log_reporter_activity_relative = 
           ave(log_reporter_activity_relative, condition, TF, FUN = function(x) mean(x))) %>%
  unique()
tf_activity_heatmap2 <- tf_activity_heatmap2 %>% 
  dcast(TF ~ condition, value.var = "log_reporter_activity_relative")
tf_activity_heatmap2 <- tf_activity_heatmap2 %>% 
  remove_rownames %>% column_to_rownames(var="TF")
tf_activity_heatmap <- tf_activity_heatmap %>% 
  mutate(reporter_2 = gsub("_[1-3]$", "", reporter_2),
         log_reporter_activity_relative = 
           ave(log_reporter_activity_relative, reporter_2, condition, FUN = function(x) mean(x))) %>%
  unique()
tf_activity_heatmap <- tf_activity_heatmap %>% 
  dcast(reporter_2+TF ~ condition, value.var = "log_reporter_activity_relative")
tf_activity_heatmap <- tf_activity_heatmap %>% 
  remove_rownames %>% column_to_rownames(var="reporter_2")
# Keeping the scale in the pheatmap function
myBreaks1 <- seq(0, 3.5, 0.035)


pheatmap(as.matrix(t(tf_activity_heatmap2)),
         color = colorRampPalette(brewer.pal(n = 7, name = "Greys"))(100),
         breaks = myBreaks1, border_color = "#565656", 
         cellwidth = 12, cellheight = 10,
         angle_col = 45)

myBreaks1 <- seq(0, 3.5, 0.035)

# pheatmap function
tf_activity_heatmap <- tf_activity_heatmap[,c(1,7,3,2,4,5,6,8)]
for(i in unique(tf_activity_heatmap$TF)){
  pheatmap(as.matrix(t(tf_activity_heatmap[tf_activity_heatmap$TF == i,-1])),
         color = colorRampPalette(brewer.pal(n = 7, name = "Greys"))(100),
         border_color = "black",
         breaks = myBreaks1, cluster_rows = F,
         cellwidth = 10, cellheight = 10)
}
```
### All of these heatmaps conclude that there we have informative reporters for ~10 TFs, and that the TF reporter design matters for some but not all TFs 





## SuperPlot of TF activity per condition - this way we can plot not only the mean, but the complete data distribution across technical and biological replicates
```{r SuperPlot, out.width= "100%", fig.align= "center", echo=FALSE, warning= FALSE}
# # Plot for each tf-activity per condition per biological replicate the techical replicate distribution and means
# ## Compute relevant mean and sd values
# activity_mean <- cDNA_df %>% 
#   select(activity, condition, rep, TF, promoter, neg_ctrls, hPGK, native_enhancer, rand_promoter) %>% 
#   group_by(condition, rep, TF, promoter) %>% 
#   mutate(activity = mean(activity))
# 
# activity_mean <- activity_mean %>% 
#   group_by(condition, TF, promoter) %>% 
#   mutate(mean = mean(activity),
#          se = sd(activity)/sqrt(length(unique((activity)))))
# 
# activity_mean <- activity_mean %>% 
#   mutate(upper = mean + se, lower = mean - se)
# 
# 
# ## All TFs
# activity_mean_1 <- activity_mean[activity_mean$promoter == "minP" & 
#                         activity_mean$neg_ctrls == "No" &
#                         activity_mean$native_enhancer == "No",] %>% unique()
# 
# ## hPGK controls
# activity_mean_2 <- activity_mean[activity_mean$promoter == "hPGK" & 
#                         activity_mean$neg_ctrls == "No" & 
#                         activity_mean$native_enhancer == "No" ,] %>% unique()
# 
# ## Native enhancers
# activity_mean_3 <- activity_mean[activity_mean$promoter == "minP" & 
#                         activity_mean$native_enhancer == "Yes",] %>% unique()
# 
# 
# ## Random promoters
# activity_mean_4 <- activity_mean[activity_mean$promoter == "Random" & 
#                         activity_mean$neg_ctrls == "No" & 
#                         activity_mean$native_enhancer == "No",] %>% unique()
# 
# ## Negative controls
# activity_mean_5 <- activity_mean[activity_mean$promoter == "minP" & 
#                         activity_mean$neg_ctrls == "Yes" & 
#                         activity_mean$native_enhancer == "No",] %>% unique()
# 
# 
# ## Generate SuperPlot
# ### Selection 1 - All TFs
# for (i in 1:4) {
#   p <- ggplot(cDNA_df[cDNA_df$neg_ctrls == "No" &
#                         cDNA_df$native_enhancer == "No",], 
#               aes(x=condition, y=activity, color=factor(rep))) + 
#     scale_color_manual(values = c("#1B998B", "#2D3047", "#FF9B71")) +
#     geom_quasirandom(aes(alpha = 0.9), dodge.width = 0.4) +
#     geom_errorbar(data = activity_mean_1, 
#                   aes(ymin = lower, 
#                       ymax = upper), colour = "black") + 
#     geom_quasirandom(data = activity_mean_1, aes(alpha = 1), size=4) + 
#     stat_compare_means(data=activity_mean_1, 
#                        comparisons = list(c(1:2)), 
#                    method="t.test", paired=TRUE) + 
#     labs(title = "TF activity distribution per condition", 
#          subtitle = paste("Page", i)) +
#     facet_wrap_paginate(~TF, ncol = 3, nrow = 3, page = i) +
#     theme_bw() +
#     ylim(0,6) +
#     theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10), 
#           axis.text.y = element_text(size = 10)) +
#     ylab("TF reporter activity") + xlab("") + 
#     theme(text = element_text(size = 12))
#   print(p)
# }
# 
# 
# 
# ### Selection 2 - hPGK controls
# ggplot(cDNA_df[cDNA_df$promoter == "hPGK" & 
#                         cDNA_df$neg_ctrls == "No" &
#                         cDNA_df$native_enhancer == "No",], 
#               aes(x=condition, y=activity, color=factor(rep))) + 
#     scale_color_manual(values = c("#1B998B", "#2D3047", "#FF9B71")) +
#     geom_quasirandom(aes(alpha = 0.9), dodge.width = 0.4) +
#     geom_errorbar(data = activity_mean_2, 
#                   aes(ymin = lower, 
#                       ymax = upper), colour = "black") + 
#     geom_quasirandom(data = activity_mean_2, aes(alpha = 1), size=4) + 
#     stat_compare_means(data=activity_mean_2, 
#                        comparisons = list(c(1:2)), 
#                    method="t.test", paired=TRUE) + 
#     labs(title = "hPGK controls: activity distribution per condition", 
#          subtitle = p) +
#     theme_bw() +
#     theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10), 
#           axis.text.y = element_text(size = 10)) +
#     ylab("TF reporter activity") + xlab("") + 
#     theme(text = element_text(size = 12))
# 
# 
# 
# ### Selection 3 - Native enhancers
# for (i in 1:5) {
#   p <- ggplot(cDNA_df[cDNA_df$native_enhancer == "Yes",], 
#               aes(x=condition, y=activity, color=factor(rep))) + 
#     scale_color_manual(values = c("#1B998B", "#2D3047", "#FF9B71")) +
#     geom_quasirandom(aes(alpha = 0.9), dodge.width = 0.4) +
#     geom_errorbar(data = activity_mean_3, 
#                   aes(ymin = lower, 
#                       ymax = upper), colour = "black") + 
#     geom_quasirandom(data = activity_mean_3, aes(alpha = 1), size=4) + 
#     stat_compare_means(data=activity_mean_3, 
#                        comparisons = list(c(1:2)), 
#                    method="t.test", paired=TRUE) + 
#     labs(title = "Miguels native enhancer chunks per condition", 
#          subtitle = paste("Page", i)) +
#     facet_wrap_paginate(~TF, ncol = 3, nrow = 3, page = i) +
#     ylim(0,6) +
#     theme_bw() +
#     theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10), 
#           axis.text.y = element_text(size = 10)) +
#     ylab("TF reporter activity") + xlab("") + 
#     theme(text = element_text(size = 12))
#   print(p)
# }
# 
# 
# ### Selection 5 - All TF-negative controls
# for (i in 1:3) {
#   p <- ggplot(cDNA_df[cDNA_df$promoter == "minP" & 
#                         cDNA_df$neg_ctrls == "Yes" &
#                         cDNA_df$native_enhancer == "No",], 
#               aes(x=condition, y=activity, color=factor(rep))) + 
#     scale_color_manual(values = c("#1B998B", "#2D3047", "#FF9B71")) +
#     geom_quasirandom(aes(alpha = 0.9), dodge.width = 0.4) +
#     geom_errorbar(data = activity_mean_5, 
#                   aes(ymin = lower, 
#                       ymax = upper), colour = "black") + 
#     geom_quasirandom(data = activity_mean_5, aes(alpha = 1), size=4) + 
#     stat_compare_means(data=activity_mean_5, 
#                        comparisons = list(c(1:2)), 
#                    method="t.test", paired=TRUE) + 
#     labs(title = "TF activity distribution per condition - only negative controls", 
#          subtitle = paste("Page", i)) +
#     facet_wrap_paginate(~TF, ncol = 3, nrow = 3, page = i) +
#     theme_bw() +
#     ylim(0,6) +
#     theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10), 
#           axis.text.y = element_text(size = 10)) +
#     ylab("TF reporter activity") + xlab("") + 
#     theme(text = element_text(size = 12))
#   print(p)
# }
```


# SuperPlots comparing different designs 
```{r superplot_2, out.width= "100%", fig.align= "center", echo=FALSE, warning= FALSE}
# # Same as before - now compare different designs
# ## Generate facet_wrap plot for each TF - each facet is a different design 
# ## Compute values for single and combinatorial designs
# 
# 
# ## Compute relevant mean and sd values
# activity_mean <- cDNA_df %>% 
#   select(activity, condition, rep, TF, promoter, neg_ctrls, 
#          hPGK, native_enhancer, rand_promoter, reporter_id, spacing, 
#          distance, background) %>% 
#   group_by(condition, rep, reporter_id) %>% 
#   mutate(activity = mean(activity))
# 
# activity_mean <- activity_mean %>% 
#   group_by(condition, reporter_id) %>% 
#   mutate(mean = mean(activity),
#          se = sd(activity)/sqrt(length(unique((activity)))))
# 
# activity_mean <- activity_mean %>% 
#   mutate(upper = mean + se, lower = mean - se)
# 
# activity_mean <- activity_mean[activity_mean$neg_ctrls == "No" &
#                         activity_mean$native_enhancer == "No",] %>% unique()
# 
# 
# ## Generate SuperPlot
# ### Selection 1 - All Designs separately
# for (i in 1:4) {
#   p <- ggplot(cDNA_df[cDNA_df$TF == "Trp53" & 
#                         cDNA_df$neg_ctrls == "No" &
#                         cDNA_df$native_enhancer == "No",], 
#               aes(x=condition, y=activity, color=factor(rep))) + 
#     scale_color_manual(values = c("#1B998B", "#2D3047", "#FF9B71")) +
#     geom_quasirandom(aes(alpha = 0.9), dodge.width = 0.4) +
#     geom_errorbar(data = activity_mean[activity_mean$TF == "Trp53",], 
#                   aes(ymin = lower, 
#                       ymax = upper), colour = "black") + 
#     geom_quasirandom(data = activity_mean[activity_mean$TF == "Trp53",], aes(alpha = 1), size=4) + 
#     stat_compare_means(data=activity_mean[activity_mean$TF == "Trp53",], 
#                        comparisons = list(c(1:2)), 
#                    method="t.test", paired=TRUE) + 
#     labs(title = "Trp53 activity distribution per design and condition", 
#          subtitle = paste("Page", i)) +
#     facet_wrap_paginate(~reporter_id, ncol = 3, nrow = 3, page = i) +
#     theme_bw() +
#     ylim(0,16) +
#     theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10), 
#           axis.text.y = element_text(size = 10)) +
#     ylab("TF reporter activity") + xlab("") + 
#     theme(text = element_text(size = 12))
#   print(p)
# }
# 
# 
# ### All reporters next to each other
# ggplot(cDNA_df[cDNA_df$TF == "Trp53" & 
#                         cDNA_df$neg_ctrls == "No" &
#                         cDNA_df$native_enhancer == "No",], 
#               aes(x=reporter_id, y=log_activity)) + 
#     scale_color_manual(values = c("#1B998B", "#2D3047", "#FF9B71")) +
#     geom_boxplot() +
#     labs(title = "Trp53 activity distribution per design and condition") +
#     theme_bw() +
#     ylim(-2,6) +
#     theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10), 
#           axis.text.y = element_text(size = 10)) +
#     ylab("TF reporter activity") + xlab("") + 
#     theme(text = element_text(size = 12))
# 
# 
# 
# 
# 
# ## Subdivide in different designs 
# ### Spacing and distance
# cDNA_df2 <- cDNA_df %>% 
#   mutate(space_dist = paste(spacing, distance)) %>% 
#   select(activity, condition, TF, space_dist, neg_ctrls, native_enhancer, promoter) 
# 
# cDNA_df2 <- cDNA_df2[cDNA_df2$neg_ctrls == "No" &
#                        cDNA_df2$native_enhancer == "No" & 
#                        cDNA_df2$promoter != "Random" &
#                        cDNA_df2$promoter != "hPGK",] %>% unique()
# 
# 
# 
# for(i in 1:8){
#   p <- ggplot(cDNA_df2, 
#         aes(x=condition, y=activity, color=space_dist)) + 
#       scale_color_manual(values = c("#1B998B", "#2D3047", "#FF9B71", "#ECDD7B")) +
#       geom_quasirandom(dodge.width = 0.9) +
#       geom_boxplot(position = position_dodge(0.9), aes(alpha = 0.5)) + 
#       labs(title = "Comparing spacings: TF activity distribution per condition", 
#            subtitle = paste("Page", i)) +
#       facet_wrap_paginate(~TF, nrow = 2, ncol = 2, page = i) +
#       theme_bw() +
#       ylim(0,6) +
#       theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10), 
#            axis.text.y = element_text(size = 10)) +
#       ylab("TF reporter activity") + xlab("") + 
#       theme(text = element_text(size = 12))
#   print(p)
# }
# 
# 
# 
# ### Background
# cDNA_df2 <- cDNA_df %>% 
#   select(activity, condition, TF, background, neg_ctrls, native_enhancer, promoter) 
# 
# 
# cDNA_df2 <- cDNA_df2[cDNA_df2$neg_ctrls == "No" &
#                        cDNA_df2$native_enhancer == "No" & 
#                        cDNA_df2$promoter != "Random" &
#                        cDNA_df2$promoter != "hPGK",] %>% unique()
# 
# 
# 
# for(i in 1:8){
#   p <- ggplot(cDNA_df2, 
#         aes(x=condition, y=activity, color=background)) + 
#       scale_color_manual(values = c("#1B998B", "#2D3047", "#FF9B71", "#ECDD7B")) +
#       geom_quasirandom(dodge.width = 0.9) +
#       geom_boxplot(position = position_dodge(0.9), aes(alpha = 0.5)) + 
#       labs(title = "Comparing backgrounds: TF activity distribution per condition", 
#            subtitle = paste("Page", i)) +
#       facet_wrap_paginate(~TF, nrow = 2, ncol = 2, page = i) +
#       theme_bw() +
#       ylim(0,6) +
#       theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10), 
#            axis.text.y = element_text(size = 10)) +
#       ylab("TF reporter activity") + xlab("") + 
#       theme(text = element_text(size = 12))
#   print(p)
# }
# 
# 
# ### Promoter
# cDNA_df2 <- cDNA_df %>% 
#   select(activity, condition, TF, promoter, neg_ctrls, native_enhancer, promoter) 
# 
# 
# cDNA_df2 <- cDNA_df2[cDNA_df2$neg_ctrls == "No" &
#                        cDNA_df2$native_enhancer == "No" &
#                        cDNA_df2$promoter != "hPGK",] %>% unique()
# 
# 
# 
# for(i in 1:8){
#   p <- ggplot(cDNA_df2, 
#         aes(x=condition, y=activity, color=promoter)) + 
#       scale_color_manual(values = c("#1B998B", "#2D3047", "#FF9B71", "#ECDD7B")) +
#       geom_quasirandom(dodge.width = 0.9) +
#       geom_boxplot(position = position_dodge(0.9), aes(alpha = 0.5)) + 
#       labs(title = "Comparing promoters: TF activity distribution per condition", 
#            subtitle = paste("Page", i)) +
#       facet_wrap_paginate(~TF, nrow = 2, ncol = 2, page = i) +
#       theme_bw() +
#       ylim(0,6) +
#       theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10), 
#            axis.text.y = element_text(size = 10)) +
#       ylab("TF reporter activity") + xlab("") + 
#       theme(text = element_text(size = 12))
#   print(p)
# }
# 
# 
# ### Promoter - native enhancers
# cDNA_df2 <- cDNA_df %>% 
#   select(activity, condition, TF, background, neg_ctrls, native_enhancer, promoter) 
# 
# 
# cDNA_df2 <- cDNA_df2[cDNA_df2$neg_ctrls == "No" &
#                        cDNA_df2$native_enhancer == "Yes" &
#                        cDNA_df2$promoter != "hPGK",] %>% unique()
# 
# 
# 
# for(i in 1:10){
#   p <- ggplot(cDNA_df2, 
#         aes(x=condition, y=activity, color=promoter)) + 
#       scale_color_manual(values = c("#1B998B", "#2D3047", "#FF9B71", "#ECDD7B")) +
#       geom_quasirandom(dodge.width = 0.9) +
#       geom_boxplot(position = position_dodge(0.9), aes(alpha = 0.5)) + 
#       labs(title = "Miguels native enhancers per promoter", 
#            subtitle = paste("Page", i)) +
#       facet_wrap_paginate(~TF, nrow = 2, ncol = 2, page = i) +
#       theme_bw() +
#       ylim(0,6) +
#       theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10), 
#            axis.text.y = element_text(size = 10)) +
#       ylab("TF reporter activity") + xlab("") + 
#       theme(text = element_text(size = 12))
#   print(p)
# }
```








## Log-linear expression modelling to explain variance - model for each TF
```{r log-linear-model-3, out.width= "100%", fig.align= "center", echo=FALSE, warning= FALSE}
### For each TF separately - only the non-mutated 
cDNA_df_TF <- cDNA_df[cDNA_df$native_enhancer == "No" &
                        cDNA_df$hPGK == "No" & 
                        cDNA_df$neg_ctrls == "No" &
                        cDNA_df$TF != "Random1"&
                        cDNA_df$TF != "Random2"&
                        cDNA_df$TF != "Random3",] %>% 
  mutate(log_reporter_activity = ave(log_reporter_activity, reporter_id, condition, FUN = function(x) mean(x))) %>%
  select(log_reporter_activity, TF, condition, promoter, spacing, distance, background) %>%
  unique()

cDNA_df_rel <- cDNA_df_ctrl[cDNA_df_ctrl$neg_ctrls == "No",] %>% 
  mutate(TF = gsub("(^.*)_[0-9]{1,2}bp_[0-9]{1,2}bp_.*", "\\1", reporter_2),
         promoter = gsub(".*(minP|mCMV|hBGm|Random)_.*", "\\1", reporter_2),
         distance = gsub(".*_([0-9]{1,2}bp)_.*", "\\1", reporter_2),
         spacing = gsub(".*_([0-9]{1,2}bp)_[0-9]{1,2}bp_.*", "\\1", reporter_2),
         background = gsub(".*([0-9]{1}$)", "\\1", reporter_2)) %>%
  select(condition, log_reporter_activity_relative, TF, promoter, distance, spacing, background)

cDNA_df_TF <- merge(cDNA_df_TF, cDNA_df_rel, all = T)

### Calculate expression variance
exp_variance <- data.frame(c("condition", "promoter", 
                             "spacing", "distance", "background", "Residuals"))
prediction <- data.frame(1:504)
names(prediction) <- "id"
names(exp_variance) <- "feature"
weight <- data.frame(c("(Intercept)", "condition2i_pos_LIF", 
                  "conditionLIF_pos_CH", "conditionLIF_pos_PD", "conditionN2B27", 
                  "conditionNPC", "conditionvitA", "promotermCMV", "promoterminP", 
                  "promoterRandom",
                  "spacing10bp", "distance21bp", "background2", "background3"))
names(weight) <- "feature"


cDNA_df_TF$row <- rownames(cDNA_df_TF)



for (i in unique(cDNA_df_TF$TF)) {
  x <- lm(log_reporter_activity_relative ~ condition + 
                       promoter + spacing + distance + background, 
                     cDNA_df_TF[cDNA_df_TF$TF == i,])
  y <- data.frame(x$fitted.values) %>% rownames_to_column()
  par(mfrow=c(2,2))
  plot(x, main = i)
  names(y) <- c("row", i)
  y$id <- 1:nrow(y)
  prediction <- merge(prediction,y, all = T)
  
  w <- data.frame(x$coefficients) %>% rownames_to_column()
  names(w) <- c("feature", i)
  weight <- merge(weight,w, all = T)
  
  x <- anova(x) %>% rownames_to_column("feature") %>%
    setnames("Sum Sq", "sum_sq") %>%
    select(feature, sum_sq) %>%
    mutate(sum = sum(sum_sq)) %>%
    mutate(rel_sum_sq = (sum_sq/sum)*100) %>%
    setnames("rel_sum_sq", i)
  x <- x[,c(1,4)]
  
  exp_variance <- merge(exp_variance, x)
}


## Prepare lm parameters for plotting
exp_variance <- melt(exp_variance, variable.name = "TF")
exp_variance$group <- "no residual"
exp_variance$group[grep("Residual", exp_variance$feature)] <- "residual"
prediction <- prediction %>% select(-id) %>% melt(variable.name = "TF") %>% na.omit() %>%
  select(-TF) %>% setnames("value", "activity_predicted")
cDNA_df_TF <- merge(cDNA_df_TF, prediction)
weight <- weight %>% filter(feature != "(Intercept)")
weight <- melt(weight, variable.name = "TF")
categorie <- c("mutated", "condition", "promoter", 
               "spacing", "distance", "background")
weight$features <- gsub(paste(categorie, collapse="|"), "",weight$feature)
feature <- weight$features
weight$condition <- gsub(paste(feature, collapse="|"), "", weight$feature)
weight$value[is.na(weight$value)] <- 0


# Further plotting preparations
level_order <- c("2i_pos_LIF", "LIF_pos_CH", 
                 "LIF_pos_PD", "vitA", "N2B27", 
                 "NPC", "mCMV", "minP", "Random",
                 "2", "3", "10bp", "21bp", "5bp")


# Model parameters
exp_variance$value[exp_variance$feature == "Residuals"] <- 100 - exp_variance$value[exp_variance$feature == "Residuals"]
exp_variance$feature <- revalue(exp_variance$feature, c("Residuals"="total"))



ggplot(exp_variance, 
       aes(x = reorder(feature, -value), y = value, fill = group)) +
  scale_fill_manual(values = c("#1B998B", "#2D3047")) +
  geom_bar(stat = "identity") + 
  ylab("Variance explained (%)") + xlab("") + 
  labs(title = "Linear variance modelling", 
       subtitle = "Expression variance predicted by features") +
  theme_bw() +
  theme(text = element_text(size = 12)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0, size = 12), 
        axis.text.y = element_text(size = 10)) +
    guides(fill = F) + facet_wrap(~TF)



for (i in unique(cDNA_df_TF$TF)) {
  
  # Plot to show prediction accuracy
  sp <- ggscatter(cDNA_df_TF[cDNA_df_TF$TF == i,], x = "log_reporter_activity", y = "activity_predicted",
   add = "reg.line", 
   add.params = list(color = "blue", fill = "lightgray"),
   conf.int = TRUE, ylab = "Predicted log2-activity", 
   title = i,
   xlab = "Average log2-activity") + theme_bw()
  
  
  # Feature importance overview for each TF (including mean activities per feature)
  fp <- ggplot(exp_variance[exp_variance$TF == i,], 
       aes(x = reorder(feature, -value), y = value, fill = group)) +
  scale_fill_manual(values = c("#1B998B", "#2D3047")) +
  geom_bar(stat = "identity") + 
  ylab("Variance explained (%)") + xlab("") +
  ggtitle(i)+
  theme_bw() +
  theme(text = element_text(size = 12)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0, size = 12), 
        axis.text.y = element_text(size = 10)) +
    guides(fill = F)
  
  # Plot weights of parameters in model - this helps to understand the model 
  p <- ggplot(weight[weight$TF == i,], 
       aes(x = factor(features, level = level_order), y = value, fill = condition)) +
  scale_fill_manual(values = c("#1B998B", "#2D3047", "#FF9B71", "#ECDD7B", "#C2B2B4", "#A54657")) +
  ggtitle(i)+
  geom_bar(stat = "identity") + 
  ylab("Weight") + xlab("") +
  theme_bw() +
  theme(text = element_text(size = 12)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 1, size = 10), 
        axis.text.y = element_text(size = 10)) +
    guides(fill = F) 

  # Combine all plots
  sp.plot <- arrangeGrob(sp,fp,p, nrow = 2, ncol = 2)
  grid.arrange(sp.plot)
}
```


















## Log-linear expression modelling to explain variance - model for each TF - only WT - without condition
```{r log-linear-model-4, out.width= "100%", fig.align= "center", echo=FALSE, warning= FALSE}
### For each TF separately - only the non-mutated 
cDNA_df_TF <- cDNA_df[cDNA_df$native_enhancer == "No" &
                        cDNA_df$hPGK == "No" & 
                        cDNA_df$neg_ctrls == "No" &
                        cDNA_df$TF != "Random1"&
                        cDNA_df$TF != "Random2"&
                        cDNA_df$TF != "Random3",] %>% 
  mutate(log_reporter_activity = ave(log_reporter_activity, reporter_id, FUN = function(x) mean(x))) %>%
  select(log_reporter_activity, TF, promoter, spacing, distance, background) %>%
  unique()

cDNA_df_rel <- cDNA_df_ctrl[cDNA_df_ctrl$neg_ctrls == "No",] %>% 
  mutate(TF = gsub("(^.*)_[0-9]{1,2}bp_[0-9]{1,2}bp_.*", "\\1", reporter_2),
         promoter = gsub(".*(minP|mCMV|hBGm|Random)_.*", "\\1", reporter_2),
         distance = gsub(".*_([0-9]{1,2}bp)_.*", "\\1", reporter_2),
         spacing = gsub(".*_([0-9]{1,2}bp)_[0-9]{1,2}bp_.*", "\\1", reporter_2),
         background = gsub(".*([0-9]{1}$)", "\\1", reporter_2)) %>%
  select(log_reporter_activity_relative, TF, promoter, distance, spacing, background) %>% unique() %>%
  mutate(log_reporter_activity_relative = ave(log_reporter_activity_relative, 
                                              TF, promoter, distance, spacing, background, 
                                              FUN = function(x) mean(x)))

cDNA_df_TF <- merge(cDNA_df_TF, cDNA_df_rel, all = T)



### Calculate expression variance
exp_variance <- data.frame(c("promoter", "spacing", "distance", "background", "Residuals"))
prediction <- data.frame(1:504)
names(prediction) <- "id"
names(exp_variance) <- "feature"
weight <- data.frame(c("(Intercept)", "promotermCMV", "promoterminP", 
                  "spacing5bp", "distance21bp", "background2", "background3"))
names(weight) <- "feature"


cDNA_df_TF$row <- rownames(cDNA_df_TF)

for (i in unique(cDNA_df_TF$TF)) {
  x <- lm(log_reporter_activity ~ promoter + spacing + distance + background, 
                     cDNA_df_TF[cDNA_df_TF$TF == i,])
  y <- data.frame(x$fitted.values) %>% rownames_to_column()
  names(y) <- c("row", i)
  y$id <- 1:nrow(y)
  prediction <- merge(prediction,y, all = T)
  
  w <- data.frame(x$coefficients) %>% rownames_to_column()
  names(w) <- c("feature", i)
  weight <- merge(weight,w, all = T)
  
  x <- anova(x) %>% rownames_to_column("feature") %>%
    setnames("Sum Sq", "sum_sq") %>%
    select(feature, sum_sq) %>%
    mutate(sum = sum(sum_sq)) %>%
    mutate(rel_sum_sq = (sum_sq/sum)*100) %>%
    setnames("rel_sum_sq", i)
  x <- x[,c(1,4)]
  
  exp_variance <- merge(exp_variance, x)
}


## Fine-tune dfs
exp_variance <- melt(exp_variance, variable.name = "TF")
exp_variance$group <- "no residual"
exp_variance$group[grep("Residual", exp_variance$feature)] <- "residual"
prediction <- prediction %>% select(-id) %>% melt(variable.name = "TF") %>% na.omit() %>%
  select(-TF) %>% setnames("value", "activity_predicted")
cDNA_df_TF <- merge(cDNA_df_TF, prediction)
weight <- weight %>% filter(feature != "(Intercept)")
weight <- melt(weight, variable.name = "TF")
categorie <- c("mutated", "condition", "promoter", 
               "spacing", "distance", "background")
weight$features <- gsub(paste(categorie, collapse="|"), "",weight$feature)
feature <- weight$features
weight$condition <- gsub(paste(feature, collapse="|"), "", weight$feature)
weight <- weight %>% select(-feature)



# Model parameters
exp_variance$value[exp_variance$feature == "Residuals"] <- 100 - exp_variance$value[exp_variance$feature == "Residuals"]
exp_variance$feature <- revalue(exp_variance$feature, c("Residuals"="total"))

ggplot(exp_variance, 
       aes(x = reorder(feature, -value), y = value, fill = group)) +
  scale_fill_manual(values = c("#1B998B", "#2D3047")) +
  geom_bar(stat = "identity") + 
  ylab("Variance explained (%)") + xlab("") + 
  labs(title = "Linear variance modelling", 
       subtitle = "Expression variance predicted by features") +
  theme_bw() +
  theme(text = element_text(size = 12)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0, size = 12), 
        axis.text.y = element_text(size = 10)) +
  facet_wrap(~TF)




## Plot weights of parameters in model - this helps to understand the model 
level_order <- c("mCMV", "minP","Random", 
                 "2", "3", "5bp", "21bp")





## Final overview - model fit for each TF
# Feature importance overview for each TF in a single plot (bar-chart including mean activities per feature)


for (i in unique(cDNA_df_TF$TF)) {
  
  # Plot to show prediction accuracy
  sp <- ggscatter(cDNA_df_TF[cDNA_df_TF$TF == i,], x = "log_reporter_activity", y = "activity_predicted",
   add = "reg.line", 
   add.params = list(color = "blue", fill = "lightgray"),
   conf.int = TRUE, ylab = "Predicted log2-activity", 
   title = i,
   xlab = "Average log2-activity") + theme_bw()
  
  
  # Feature importance overview for each TF (including mean activities per feature)
  fp <- ggplot(exp_variance[exp_variance$TF == i,], 
       aes(x = reorder(feature, -value), y = value, fill = group)) +
  scale_fill_manual(values = c("#1B998B", "#2D3047")) +
  geom_bar(stat = "identity") + 
  ylab("Variance explained (%)") + xlab("") +
  theme_bw() +
  theme(text = element_text(size = 12)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0, size = 12), 
        axis.text.y = element_text(size = 10)) +
    guides(fill = F)
  
  # Plot weights of parameters in model - this helps to understand the model 
  p <- ggplot(weight[weight$TF == i,], 
       aes(x = factor(features, level = level_order), y = value, fill = condition)) +
  scale_fill_manual(values = c("#1B998B", "#2D3047", "#FF9B71", "#ECDD7B", "#C2B2B4", "#A54657")) +
  geom_bar(stat = "identity") + 
  ylab("Variance explained (%)") + xlab("") +
  theme_bw() +
  theme(text = element_text(size = 12)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 1, size = 10), 
        axis.text.y = element_text(size = 10)) +
    guides(fill = F) 

  # Combine all plots
  sp.plot <- arrangeGrob(sp,fp,p, nrow = 2, ncol = 2)
  grid.arrange(sp.plot)
}
```





## Can expression variance be explained by the TF properties?
```{r tf-properties, out.width= "100%", fig.align= "center", echo=FALSE, warning= FALSE}
# Import tf properties dataframe
tf_properties <- read.csv2("/DATA/usr/m.trauernicht/projects/tf_activity_reporter/data/SuRE_TF_1/tf_properties.csv", header = T)

convertletter <- function(x) {
  substr(x, 2, 20) <- tolower(substr(x, 2, 20))
  x
}

tf_properties$TF <- convertletter(tf_properties$TF)

tf_properties <- tf_properties[tf_properties$TF %in% cDNA_df$TF,]

variance <- exp_variance[exp_variance$feature == "total",] %>% select(TF, value) %>% unique()

expression <- cDNA_df_TF %>% select(log_reporter_activity_relative, TF) %>% 
  mutate(log_reporter_activity_relative = ave(log_reporter_activity_relative, TF, FUN = function(x) mean(x))) %>%
  unique()

tf_properties <- merge(tf_properties, variance)
tf_properties <- merge(tf_properties, expression)

x1 <- ggplot(tf_properties, aes(x = chromatin_regulation_mode, y = value)) +
  geom_quasirandom(width = 0.2) +
  theme_bw() +
  xlab("Chromatin regulation mode") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1.0, vjust = 0))+
  ylab("TF reporter quality")

x2 <- ggplot(tf_properties, aes(x = chromatin_regulation_mode, y = log_reporter_activity_relative)) +
  geom_quasirandom(width = 0.2) +
  theme_bw() +
  xlab("Chromatin regulation mode") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1.0, vjust = 0))+
  ylab("TF reporter expression")


#Grid the plots to one panel
sp.plot <- arrangeGrob(x1,x2, nrow = 1, ncol = 2)
grid.arrange(sp.plot)


y1 <- ggplot(tf_properties, aes(x = TF_superclass, y = value)) +
  geom_quasirandom(width = 0.2) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1.0, vjust = 0))+
  xlab("TF superclass") +
  ylab("TF reporter quality")

y2 <- ggplot(tf_properties, aes(x = TF_superclass, y = log_reporter_activity_relative)) +
  geom_quasirandom(width = 0.2) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1.0, vjust = 0))+
  xlab("TF superclass") +
  ylab("TF reporter expression")

sp.plot <- arrangeGrob(y1,y2, nrow = 1, ncol = 2)
grid.arrange(sp.plot)


z1 <- ggplot(tf_properties, aes(x = mode_of_regulation, y = value)) +
  geom_quasirandom(width = 0.2) +
  theme_bw() +
  xlab("Mode of regulation") +
  ylab("TF reporter quality")

z2 <- ggplot(tf_properties, aes(x = mode_of_regulation, y = log_reporter_activity_relative)) +
  geom_quasirandom(width = 0.2) +
  theme_bw() +
  xlab("Mode of regulation") +
  ylab("TF reporter expression")

sp.plot <- arrangeGrob(z1,z2, nrow = 1, ncol = 2)
grid.arrange(sp.plot)


x4 <- ggplot(tf_properties, aes(x = tissue_of_expression, y = value)) +
  geom_quasirandom(width = 0.2) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1.0, vjust = 0))+
  xlab("Tissue specificity") +
  ylab("TF reporter quality")

x5 <- ggplot(tf_properties, aes(x = tissue_of_expression, y = log_reporter_activity_relative)) +
  geom_quasirandom(width = 0.2) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1.0, vjust = 0))+
  xlab("Tissue specificity") +
  ylab("TF reporter expression")

sp.plot <- arrangeGrob(x4,x5, nrow = 1, ncol = 2)
grid.arrange(sp.plot)



```





## Finding the best reporters for a single TF
## Make the same models as before - but now per TF and per condition
```{r log-linear-model-4, out.width= "100%", fig.align= "center", echo=FALSE, warning= FALSE}
### For each TF separately - only the non-mutated 
cDNA_df_TF <- cDNA_df[cDNA_df$native_enhancer == "No" &
                        cDNA_df$hPGK == "No" & 
                        cDNA_df$neg_ctrls == "No" &
                        cDNA_df$TF != "Random1"&
                        cDNA_df$TF != "Random2"&
                        cDNA_df$TF != "Random3",] %>% 
  mutate(log_reporter_activity = ave(log_reporter_activity, reporter_id, condition, FUN = function(x) mean(x)),
         TF = paste(TF, condition, sep = "_")) %>%
  select(log_reporter_activity, TF, promoter, spacing, distance, background) %>%
  unique()

cDNA_df_rel <- cDNA_df_ctrl[cDNA_df_ctrl$neg_ctrls == "No",] %>% 
  mutate(TF = gsub("(^.*)_[0-9]{1,2}bp_[0-9]{1,2}bp_.*", "\\1", reporter_2),
         TF = paste(TF, condition, sep = "_"),
         promoter = gsub(".*(minP|mCMV|hBGm|Random)_.*", "\\1", reporter_2),
         distance = gsub(".*_([0-9]{1,2}bp)_.*", "\\1", reporter_2),
         spacing = gsub(".*_([0-9]{1,2}bp)_[0-9]{1,2}bp_.*", "\\1", reporter_2),
         background = gsub(".*([0-9]{1}$)", "\\1", reporter_2)) %>%
  select(log_reporter_activity_relative, TF, promoter, distance, spacing, background) %>% unique() %>%
  mutate(log_reporter_activity_relative = ave(log_reporter_activity_relative, 
                                              TF, promoter, distance, spacing, background, 
                                              FUN = function(x) mean(x)))

cDNA_df_TF <- merge(cDNA_df_TF, cDNA_df_rel, all = T)



### Calculate expression variance
exp_variance <- data.frame(c("promoter", "spacing", "distance", "background", "Residuals"))
prediction <- data.frame(1:504)
names(prediction) <- "id"
names(exp_variance) <- "feature"
weight <- data.frame(c("(Intercept)", "promotermCMV", "promoterminP", 
                  "spacing5bp", "distance21bp", "background2", "background3"))
names(weight) <- "feature"


cDNA_df_TF$row <- rownames(cDNA_df_TF)

for (i in unique(cDNA_df_TF$TF)) {
  x <- lm(log_reporter_activity ~ promoter + spacing + distance + background, 
                     cDNA_df_TF[cDNA_df_TF$TF == i,])
  y <- data.frame(x$fitted.values) %>% rownames_to_column()
  names(y) <- c("row", i)
  y$id <- 1:nrow(y)
  prediction <- merge(prediction,y, all = T)
  
  w <- data.frame(x$coefficients) %>% rownames_to_column()
  names(w) <- c("feature", i)
  weight <- merge(weight,w, all = T)
  
  x <- anova(x) %>% rownames_to_column("feature") %>%
    setnames("Sum Sq", "sum_sq") %>%
    select(feature, sum_sq) %>%
    mutate(sum = sum(sum_sq)) %>%
    mutate(rel_sum_sq = (sum_sq/sum)*100) %>%
    setnames("rel_sum_sq", i)
  x <- x[,c(1,4)]
  
  exp_variance <- merge(exp_variance, x)
}


## Fine-tune dfs
exp_variance <- melt(exp_variance, variable.name = "TF")
exp_variance$group <- "no residual"
exp_variance$group[grep("Residual", exp_variance$feature)] <- "residual"
prediction <- prediction %>% select(-id) %>% melt(variable.name = "TF") %>% na.omit() %>%
  select(-TF) %>% setnames("value", "activity_predicted")
cDNA_df_TF <- merge(cDNA_df_TF, prediction)
weight <- weight %>% filter(feature != "(Intercept)")
weight <- melt(weight, variable.name = "TF")
categorie <- c("mutated", "condition", "promoter", 
               "spacing", "distance", "background")
weight$features <- gsub(paste(categorie, collapse="|"), "",weight$feature)
ml <- length(weight$features)
feature <- weight$features
weight$condition <- gsub(paste(head(feature, n = ml/2), collapse = "|"), "", weight$feature)
weight <- weight %>% select(-feature)



# Model parameters
exp_variance$value[exp_variance$feature == "Residuals"] <- 100 - exp_variance$value[exp_variance$feature == "Residuals"]
exp_variance$feature <- revalue(exp_variance$feature, c("Residuals"="total"))

ggplot(exp_variance[grep("Creb1", exp_variance$TF),], 
       aes(x = reorder(feature, -value), y = value, fill = group)) +
  scale_fill_manual(values = c("#1B998B", "#2D3047")) +
  geom_bar(stat = "identity") + 
  ylab("Variance explained (%)") + xlab("") + 
  labs(title = "Linear variance modelling", 
       subtitle = "Expression variance predicted by features") +
  theme_bw() +
  theme(text = element_text(size = 12)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0, size = 12), 
        axis.text.y = element_text(size = 10)) +
  facet_wrap(~TF)




## Plot weights of parameters in model - this helps to understand the model 
level_order <- c("mCMV", "minP","Random", 
                 "2", "3", "5bp", "21bp")





## Final overview - model fit for each TF
# Feature importance overview for each TF in a single plot (bar-chart including mean activities per feature)


for (i in unique(cDNA_df_TF$TF[grep("Tfcp2l1", cDNA_df_TF$TF)])) {
  
  # Plot to show prediction accuracy
  sp <- ggscatter(cDNA_df_TF[cDNA_df_TF$TF == i,], x = "log_reporter_activity", y = "activity_predicted",
   add = "reg.line", 
   add.params = list(color = "blue", fill = "lightgray"),
   conf.int = TRUE, ylab = "Predicted log2-activity", 
   title = i,
   xlab = "Average log2-activity") + theme_bw()
  
  
  # Feature importance overview for each TF (including mean activities per feature)
  fp <- ggplot(exp_variance[exp_variance$TF == i,], 
       aes(x = reorder(feature, -value), y = value, fill = group)) +
  scale_fill_manual(values = c("#1B998B", "#2D3047")) +
  geom_bar(stat = "identity") + 
  ylab("Variance explained (%)") + xlab("") +
  theme_bw() +
  theme(text = element_text(size = 12)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0, size = 12), 
        axis.text.y = element_text(size = 10)) +
    guides(fill = F)
  
  # Plot weights of parameters in model - this helps to understand the model 
  p <- ggplot(weight[weight$TF == i,], 
       aes(x = factor(features, level = level_order), y = value, fill = condition)) +
  scale_fill_manual(values = c("#1B998B", "#2D3047", "#FF9B71", "#ECDD7B", "#C2B2B4", "#A54657")) +
  geom_bar(stat = "identity") + 
  ylab("Variance explained (%)") + xlab("") +
  theme_bw() +
  theme(text = element_text(size = 12)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 1, size = 10), 
        axis.text.y = element_text(size = 10)) +
    guides(fill = F) 

  # Combine all plots
  sp.plot <- arrangeGrob(sp,fp,p, nrow = 2, ncol = 2)
  grid.arrange(sp.plot)
}
```














## Exporting potential data. 
```{r}
```

# Session Info
```{r}
paste("Run time: ",format(Sys.time()-StartTime))
getwd()
date()
sessionInfo()
```

