---
title: "TF reporter activity modelling"
author: "Max Trauernicht"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: 
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---

# knitr document van Steensel lab


# Introduction

## Description of Data
How to make a good rendering table: 
```{r table1, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- "
| column1	|	column2	|	column3	|	
|----|----|----|
|1	|	2	|	3	|	
|a	|	b	|	c	|	
"
cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
```

# Data processing
## Path, Libraries, Parameters and Useful Functions
```{r setup}
knitr::opts_chunk$set(echo = TRUE)
StartTime <-Sys.time()

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8) 
# libraries:
```

### Custom functions
Functions used thoughout this script.
```{r}
SetFileName <- function(filename, initials) {
  # Set filename with extension and initials to make filename with date integrated.
  filename <- substitute(filename)
  initials <- substitute(initials)
  filename <- paste0(initials, Date, filename)
  filename
}
```


## Data import
```{r}

```

# Analysis
Following could be done on my data to explain activity variation between designs:
A model was fit using lm()in R stats package to predict average expression from the
independent contribution of background and the 6 CRE positions in the site number and affinity library (expression ~ background + site1 + site2 + site3 + site4 + site5 + site6). Background and each position was represented as categorical variables according to the 3 backgrounds used and 3 possible affinities per position (consensus CRE, weak CRE, and no CRE). The percent of variance explained per model term was obtained using the sum of squares from anova()

## Visualisation
```{r}
```

# Results
```{r}
```

# Conclusions
```{r}
```

## Exporting potential data. 
```{r}
```

# Session Info
```{r}
paste("Run time: ",format(Sys.time()-StartTime))
getwd()
date()
sessionInfo()
```

