---
title: "cDNA counts functional analysis - stimulation 1"
author: "Max Trauernicht"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: 
  prettydoc::html_pretty:
    theme: leonids
    highlight: github
  #   toc: true
  #   toc_float: true
  #   code_folding: show
  # editor_options:
  #   chunk_output_type: console
---

*knitr document van Steensel lab*

# TF reporter cDNA-count processing - stimulation 1

## Introduction
I previously processed the raw sequencing data, optimized the barcode clustering, quantified the pDNA data and normalized the cDNA data. In this script, I want to have a detailed look at the cDNA data from a general perspective.




```{r setup, out.width= "80%", fig.align= "center", echo=FALSE, warning= FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
StartTime <-Sys.time()

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8) 
# libraries:
library(RColorBrewer)
library(ggplot2)
library(dplyr)
library(maditr)
library(tibble)
library(pheatmap)
library(ggpubr)
library(ggbeeswarm)
library(ggforce)
library(viridis)
library(plyr)
library(cowplot)
library(gridExtra)
```


```{r out.width= "80%", fig.align= "center", echo=FALSE, warning= FALSE, message = FALSE}
### Custom functions
SetFileName <- function(filename, initials) {
  # Set filename with extension and initials to make filename with date integrated.
  filename <- substitute(filename)
  initials <- substitute(initials)
  filename <- paste0(initials, Date, filename)
  filename
}


cbind.fill <- function(...){
    nm <- list(...) 
    nm <- lapply(nm, as.matrix)
    n <- max(sapply(nm, nrow)) 
    do.call(cbind, lapply(nm, function (x) 
        rbind(x, matrix(, n-nrow(x), ncol(x))))) 
}

# Function to substring the right part of the motif
substrRight <- function(x, n){
  substr(x, nchar(x)-n+1, nchar(x))
}


# Function to load PWM matrix
get_pwm_feature_matrix <- function(motif_meta_fn, fimo_fn, db = 2) {

  # validate args
  valid_dbs <- 1:2
  if(!db %in% valid_dbs)
    stop('Invalid db (database version). Please use db=1 (maintained for backward compatibility only) or db=2')

  # db=1 is maintained for backward compatibility only
  if(db == 1) {

    # read in motif metadata
    motif_meta    <- read.csv(motif_meta_fn)

    # check whether motif metadata contain essential annotations
    if(!all(c('PWM.ID', 'Cognate.TF') %in% colnames(motif_meta))) {
      message('The motif metadata file does not contain the essential columns PWM.ID and Cognate.TF')
    }

    motif_minimal <- motif_meta[, c('PWM.ID', 'Cognate.TF')]

    # load fimo output --> extract motif id, sequence id and p-value
    df <- read.table(fimo_fn)
    df <- df[, c(1, 2, 7)]

    colnames(df) <- c('PWM.ID', 'seqid', 'pval')

    # add TF id
    df <- merge(df, motif_minimal, by = 'PWM.ID')

    # group motif hits by sequence id
    l <- split(df, df[['seqid']])

    # multiple PWM and multiple hits possible. Reduce hits to one per TF, keeping best p-val only
    l <- lapply(l, function(x) {
      x_by_tf <- split(x, x[['Cognate.TF']], drop = TRUE)
      x_by_tf <- lapply(x_by_tf, function(y) y[which.min(y$pval), ])
      do.call('rbind', x_by_tf)
    })

    # initialize feature matrix
    n_tf          <- motif_minimal[['Cognate.TF']] %>%
      unique %>%
      length
    n_seq         <- length(l)
    pwm           <- matrix(1, nrow = n_seq, ncol = n_tf)
    colnames(pwm) <- (motif_minimal[['Cognate.TF']] %>% unique)

    # replace :: from names of composite motifs
    colnames(pwm) <- str_replace_all(colnames(pwm), '::', '_')

    # fill in feature matrix
    for(i in 1 : n_seq) {
      pwm[i, l[[i]][['Cognate.TF']]] <- l[[i]]$pval
    }

    # -log10 transform
    pwm           <- -1 * log10(pwm)

    # coerce to tib and return
    tib_fimo <- as_data_frame(pwm) %>%
      mutate(id = names(l))
      dplyr::select(id, everything())

  }

  # db = 2 (default)
  else {

    # load metadata
    tib_meta    <- read_csv(motif_meta_fn) %>%
      # extract tf symbol from motif id (Cognate_TF unsafe, it can be empty) and replace :: occurrences
      mutate(tf_symbol = str_remove(ID, '_[0-9]*'),
             tf_symbol = str_replace(tf_symbol, '::', '_')) %>%
      dplyr::select(motif_id = `PWM ID`, tf_symbol)

    # load fimo results
    tib_fimo <- read_tsv(fimo_fn) %>%
      # extract motif id, sequence id and p-value
      dplyr::select(motif_id, sequence_name, pval = `p-value`)

    # add tf symbol to fimo results
    tib_fimo <- tib_fimo %>%
      left_join(tib_meta, by = 'motif_id') %>%
      # remove hits with missing motif id (composite pwms)
      filter(!is.na(tf_symbol))

    # select best hit for each motif and sequence
    tib_fimo <- tib_fimo %>%
      group_by(sequence_name, tf_symbol) %>%
      dplyr::slice(which.min(pval)) %>%
      ungroup()

    # spread into feature matrix
    tib_fimo <- tib_fimo %>%
      mutate(pval = -1 * log10(pval)) %>%
      dplyr::select(-motif_id) %>%
      spread(key = tf_symbol, value = pval, fill = 0, drop = TRUE) %>%
      # perform cosmetics on the id
      mutate(id = sequence_name) %>%
      dplyr::select(-c(sequence_name)) %>%
      dplyr::select(id, everything())

  }

  return(tib_fimo)

}

```


```{r out.width= "80%", fig.align= "center", echo=FALSE, warning= FALSE, message = FALSE}
# Import processed bc counts from the preprocessing step
cDNA_df <- read.csv("/DATA/usr/m.trauernicht/projects/SuRE-TF/data/gcf5927_stimulation-1/results/mt20210121_reporter_activity_filt.csv", header = T)
cDNA_df$background <- as.character(cDNA_df$background)
cDNA_df$background[is.na(cDNA_df$background)] <- ""
cDNA_df$TF <- revalue(cDNA_df$TF, c("Tcfcp2l1" = "Tfcp2l1", "Tcfcp2l1_neg" = "Tfcp2l1_neg"))
cDNA_df$reporter_id <- gsub("Tcfcp2l1", "Tfcp2l1", cDNA_df$reporter_id)
```

## Analysis

### First insights into data distribution - reporter activity distribution plots
```{r data_density, out.width= "80%", fig.align= "center", echo=FALSE, warning= FALSE, message = FALSE}
## reporter activity distribution over all experiments
colors <- c("#1B998B", "#2D3047", "#FF9B71")

## reporter activity distribution for each experiment individually
ggplot(data = cDNA_df %>% dplyr::select(condition, log_reporter_activity, reporter_id) %>% unique(), aes(y = log_reporter_activity, x = "x")) + 
  geom_quasirandom(alpha = 0.4) + xlab("log2(cDNA/pDNA)") + 
  labs(title = "reporter activity distribution per condition") + theme_bw() + facet_wrap(vars(condition))

## reporter activity distribution - highlight negative motif controls
ggplot(cDNA_df[cDNA_df$hPGK == "No",], aes(x = log_activity, fill = neg_ctrls, alpha = 0.4)) + 
  geom_density() + xlab("log2(cDNA/pDNA)") + 
  scale_fill_manual(values = colors)+
  labs(title = "reporter activity distribution - negative ctrls", subtitle = "Green = TF-mediated activity") +
  xlim(-10,10) + theme_bw()+
  theme(text = element_text(size = 12))

## reporter activity distribution - highlight Miguels enhancer control vs negative controls
ggplot(data = cDNA_df %>% 
         filter(native_enhancer == "Yes" | neg_ctrls == "Yes", condition == "2i_pos_LIF") %>%
         dplyr::select(condition, log_reporter_activity, reporter_id, native_enhancer, promoter) %>% unique(), 
       aes(y = log_reporter_activity, x = promoter, color = native_enhancer, alpha = 0.4)) + 
  geom_quasirandom(dodge.width = .75) + xlab("minimal promoter") + 
  labs(title = "enhancer chunks (green), shuffled") +
  scale_color_manual(values = colors) +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 1)) 


## reporter activity distribution - highlight different promoters
ggplot(data = cDNA_df %>% dplyr::select(log_reporter_activity, reporter_id, promoter) %>% unique(), 
       aes(y = log_reporter_activity, x = promoter, color = promoter, alpha = 0.4)) + 
  geom_quasirandom() + xlab("log2(cDNA/pDNA)") + 
  labs(title = "reporter activity distribution - different promoters") +
  scale_color_brewer(palette = "Set2") +
  theme_bw()

## reporter activity distribution - highlight different promoters - and negative controls (does the activity actually come from the TFs or from the promoter itself?)
ggplot(cDNA_df %>% 
         filter(promoter != "hPGK") %>%
         mutate(TF = gsub("_neg", "", TF)) %>%
         dplyr::select(reporter_id, log_reporter_activity, neg_ctrls, promoter, TF) %>% 
         unique(), aes(y = log_reporter_activity, x = neg_ctrls, color = promoter, alpha = 0.4)) + 
  geom_quasirandom(dodge.width = 0.75) + xlab("mutated motif?") + 
  labs(title = "reporter activity distribution - promoters") +
  scale_color_brewer(palette = "Set2") +
  theme_bw() + facet_wrap(~TF)

## reporter activity distribution - highlight different distances individually per TF
ggplot(data = cDNA_df[cDNA_df$distance != "",] %>% 
         filter(neg_ctrls == "No", hPGK == "No", rand_promoter == "No") %>%
         dplyr::select(log_reporter_activity, reporter_id, distance, TF) %>% unique(), 
       aes(y = log_reporter_activity, x = distance, color = distance, alpha = 0.4)) + 
  geom_quasirandom() + ylab("log2(cDNA/pDNA)") + 
  labs(title = "reporter activity distribution - distances") +
  scale_color_brewer(palette = "Set2") +
  theme_bw()+ facet_wrap(~TF)

## reporter activity distribution - highlight different spacings individually per TF
ggplot(data = cDNA_df[cDNA_df$spacing != "",] %>% 
         filter(neg_ctrls == "No", hPGK == "No", rand_promoter == "No") %>%
         dplyr::select(log_reporter_activity, reporter_id, spacing, TF, condition) %>% 
         unique(), 
       aes(y = log_reporter_activity, x = condition, color = spacing, alpha = 0.4)) + 
  geom_quasirandom(dodge.width = 0.75) + ylab("log2(cDNA/pDNA)") + 
  labs(title = "reporter activity distribution - spacings") +
  theme_bw() + facet_wrap(~TF) +
  scale_color_manual(values = colors)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 1)) 


## highlight spacings for p53 in different conditions
ggplot(data = cDNA_df[cDNA_df$spacing != "" & cDNA_df$TF == "Trp53",] %>% 
         filter(neg_ctrls == "No", promoter == "mCMV", rand_promoter == "No", str_detect(condition, "2i_pos_LIF|NPC")) %>%
         dplyr::select(log_activity, reporter_id, spacing, TF, condition) %>% 
         unique(), 
       aes(y = log_activity, x = condition, color = spacing, alpha = 0.4)) + 
  geom_quasirandom(dodge.width = 0.75) + 
  geom_boxplot(position = position_dodge(0.75)) +
  ylab("log2(cDNA/pDNA)") + 
  labs(title = "reporter activity distribution - spacings") +
  scale_color_manual(values = colors)+
  theme_bw() +
  ylim(-1,5)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 1)) 


## reporter activity distribution - highlight different backgrounds
ggplot(data = cDNA_df[cDNA_df$background != "",] %>% dplyr::select(log_reporter_activity, reporter_id, background, condition) %>% unique(), 
       aes(y = log_reporter_activity, x = background, color = background, alpha = 0.4)) + 
  geom_quasirandom() + xlab("log2(cDNA/pDNA)") + 
  labs(title = "reporter activity distribution - backgrounds") +
  scale_color_brewer(palette = "Set2") +
  theme_bw()

## reporter activity distribution - highlight different backgrounds - each TF and each condition
ggplot(data = cDNA_df[cDNA_df$spacing != "",] %>% 
         filter(neg_ctrls == "No", hPGK == "No", rand_promoter == "No") %>%
         dplyr::select(log_reporter_activity, reporter_id, background, TF, condition) %>% 
         unique(), 
       aes(y = log_reporter_activity, x = condition, color = background, alpha = 0.4)) + 
  geom_quasirandom(dodge.width = 0.75) + ylab("log2(cDNA/pDNA)") + 
  labs(title = "reporter activity distribution - backgrounds") +
  theme_bw() + facet_wrap(~TF) +
  scale_color_manual(values = colors)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 1)) 

## reporter activity distribution - highlight different backgrounds - zooming in on Stat3
ggplot(data = cDNA_df[cDNA_df$background != "" & cDNA_df$TF == "Stat3",] %>% 
         filter(neg_ctrls == "No", rand_promoter == "No") %>%
         dplyr::select(log_activity, reporter_id, background, TF, condition) %>% 
         unique(), 
       aes(y = log_activity, x = condition, color = background, alpha = 0.4)) + 
  geom_quasirandom(dodge.width = 0.75) + 
  geom_boxplot(position = position_dodge(0.75)) +
  ylab("log2(cDNA/pDNA)") + 
  labs(title = "reporter activity distribution - backgrounds") +
  scale_color_manual(values = colors)+
  theme_bw() +
  ylim(-1,5)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 1)) 


## correlation bc-counts pDNA vs. cDNA
sp <- ggscatter(cDNA_df[cDNA_df$native_enhancer == "No" &
               cDNA_df$neg_ctrls  == "No" &
               cDNA_df$hPGK == "No" &
               cDNA_df$rand_promoter == "No",],
               x = "pDNA_counts_rpm", y = "rpm",
               size = 0.1, alpha = 0.3, color = "#2D3047", 
   add = "reg.line",
   add.params = list(color = "blue", fill = "lightgray"), title = "normalized barcode counts cDNA vs. pDNA",
   conf.int = TRUE, ylab = "cDNA", xlab = "pDNA")
sp + stat_cor(method = "pearson", label.x = 150, label.y = 2) + 
  geom_abline(linetype = "dashed") + facet_wrap(~TF) + xlim(0,200) + ylim(0,200)


## correlation bc-counts pDNA vs. reporter activity
sp <- ggscatter(cDNA_df[cDNA_df$native_enhancer == "No" &
                    cDNA_df$neg_ctrls  == "No" &
                    cDNA_df$hPGK == "No" &
                    cDNA_df$rand_promoter == "No",],
          size = 0.75, color = "condition", alpha = 0.1,
          y = "log_activity", x = "pDNA_counts_rpm",
          title = "reporter pDNA counts vs log2 reporter activity",
          ylab = "log2 reporter activity",
          xlab = "pDNA counts per million",
          facet.by = "TF")
sp + xlim(0,200) + ylim(-5,5) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10), 
          axis.text.y = element_text(size = 10)) +
  theme(text = element_text(size = 12))

## let's zoom into a model TF with high noise at low pDNA: Gbx2
sp <- ggscatter(cDNA_df[cDNA_df$TF == "Gbx2" &
                    cDNA_df$neg_ctrls  == "No" &
                    cDNA_df$hPGK == "No" &
                    cDNA_df$rand_promoter == "No",],
          size = 0.1, color = "#2D3047",
          y = "log_activity", x = "pDNA_counts_rpm",
          title = "Gbx2: reporter pDNA counts vs log2 reporter activity",
          ylab = "log2 reporter activity",
          xlab = "pDNA counts per million",
          facet.by = "condition")
sp + xlim(0,200) + ylim(-5,5)

## let's zoom Stat3 because it's nice
sp <- ggscatter(cDNA_df[cDNA_df$TF == "Stat3" &
                    cDNA_df$neg_ctrls  == "No" &
                    cDNA_df$hPGK == "No" &
                    cDNA_df$rand_promoter == "No",],
          size = 0.1, color = "#2D3047",
          y = "log_activity", x = "pDNA_counts_rpm",
          title = "Stat3: reporter pDNA counts vs log2 reporter activity",
          ylab = "log2 reporter activity",
          xlab = "pDNA counts per million",
          facet.by = "condition")
sp + xlim(0,200) + ylim(-5,5)
```




### Heat map - display mean log2-activity for each TF in each condition
```{r tf_activity_heatmap, out.width= "80%", fig.align= "center", echo=FALSE, warning= FALSE, message = FALSE}
## Prepare dataframe
# Caculate mean TF activity per condition
cDNA_df$tf_activity <- ave(cDNA_df$reporter_activity, cDNA_df$condition, cDNA_df$TF, FUN = function(x) mean(x))
cDNA_df$tf_activity <- log2(cDNA_df$tf_activity)
cDNA_df$group  <- "Other"
neural <- c("Gli1", "Homez", "Irx3", "Neurog2", "Otx1", "Pax6", "Rbpj")
cDNA_df$group[grep(paste(neural, collapse = "|"), cDNA_df$TF)] <- "Neural induction"
ES <- c("Esrrb", "Gbx2", "Klf4", "Smad4", "Stat3", "Tcf7l2", "Tfcp2l1", "Zfp42", "Zfx")
cDNA_df$group[grep(paste(ES, collapse = "|"), cDNA_df$TF)] <- "Stemness"
pos <- c("Sp1", "Nfya", "Random")
cDNA_df$group[grep(paste(pos, collapse = "|"), cDNA_df$TF)] <- "Control"
  
tf_activity_heatmap <- cDNA_df[cDNA_df$neg_ctrls == "No" & cDNA_df$hPGK == "No" &
                               cDNA_df$native_enhancer == "No" & cDNA_df$rand_promoter == "No",] %>% 
  dplyr::select(TF, condition, tf_activity) %>% unique()
tf_activity_heatmap <- dcast(tf_activity_heatmap, TF ~ condition, value.var="tf_activity")
group <- cDNA_df %>% dplyr::select(TF, group) %>% unique()
tf_activity_heatmap <- merge(tf_activity_heatmap, group)

tf_activity_heatmap <- tf_activity_heatmap %>% 
  remove_rownames %>% column_to_rownames(var="TF")
# Keeping the scale in the pheatmap function
myBreaks1 <- seq(0,3.5,0.035)


# pheatmap function
pheatmap(as.matrix(t(tf_activity_heatmap[,-8])),
         color = colorRampPalette(brewer.pal(n = 7, name = "Greys"))(100),
         breaks = myBreaks1, border_color = "#565656", 
         cellwidth = 12, cellheight = 10,
         angle_col = 90)

# selective representation
tf_activity_heatmap <- tf_activity_heatmap[,c(6,2,1,3,4,5,7,8)]
for (i in unique(tf_activity_heatmap$group)) {
  p <- pheatmap(as.matrix(t(tf_activity_heatmap[tf_activity_heatmap$group == i,-8])),
         color = colorRampPalette(brewer.pal(n = 7, name = "Greys"))(100),
         breaks = myBreaks1, border_color = "black", 
         cellwidth = 12, cellheight = 10, cluster_rows = F,
         angle_col = 90)
  print(p) 
}

```
*~10 of the 26 TFs in the library show promising activity at the first glimpse*

### Heatmap for native enhancers
```{r out.width= "80%", fig.align= "center", echo=FALSE, warning= FALSE, message = FALSE}
## Prepare dataframe
# Caculate mean TF activity per condition
tf_activity_heatmap <- cDNA_df[cDNA_df$native_enhancer == "Yes",] %>% 
  dplyr::select(TF, condition, tf_activity) %>% unique()
tf_activity_heatmap <- dcast(tf_activity_heatmap, condition ~ TF, value.var="tf_activity")

tf_activity_heatmap <- tf_activity_heatmap %>% 
  remove_rownames %>% column_to_rownames(var="condition")

comb_activity <- data.frame("pos" = colnames(tf_activity_heatmap), 
                             "activity" = colSums(tf_activity_heatmap)) 

ggplot(comb_activity, aes(x = pos, y = activity)) +
  geom_bar(stat = "identity", aes(fill = activity)) + theme_classic() +
    theme(axis.title.y = element_blank(), axis.text.y = element_blank(), 
          axis.ticks.y = element_blank(), axis.ticks.x = element_blank(),
          axis.text.x = element_blank(), 
          axis.title.x = element_blank(),
          legend.position = "none") +
    scale_fill_distiller(name = "activity", palette = "Greys", direction = 1)



# Keeping the scale in the pheatmap function
myBreaks1 <- seq(0, 4, 0.04)


# pheatmap function
tf_activity_heatmap <- tf_activity_heatmap[c(6,2,1,3,4,7,5),]
pheatmap(as.matrix(tf_activity_heatmap),
         color = colorRampPalette(brewer.pal(n = 7, name = "Greys"))(100),
         breaks = myBreaks1, border_color = "black", 
         cellwidth = 10, cellheight = 10, cluster_rows = F, cluster_cols = F)

# Based on heatmap: are there Tfcp2l1 or Tcf enrichments in e11_klf2_7 or Stat3 enrichments in e11_klf2_5 or Tcf enrichments in e97_klf2_5/6

# Based on https://molotool.autosome.ru/:
## 1xStat5 motif in e11_klf2_5 
## no candidate motif in e11_klf2_7 
## 1xTcf7l2 in e97_klf2_5 (but also in e97_klf2_4 -> why is this one not expressed)

## I should do a more systematic motif enrichment analysis
## Generate FIMO matrix 
promoters <- c("TAGAGGGTATATAATGGAAGCTCGACTTCCAG",
               "GGCGTTTACTATGGGAGGTCTATATAAGCAGAGCTCGTTTAGTGAACCGTCAGATC",
               "GGGCTGGGCATAAAAGTCAGGGCAGAGCCATCTATTGCTTACATTTGCTTCT",
               "GGTTAGCGATCCAATTCAGCTAGATTTTAAGC")

cDNA_df_native <- cDNA_df[cDNA_df$native_enhancer == "Yes",] %>% dplyr::select(TF, seq) %>%
  mutate(seq = gsub("CACGACGCTCTTCCGATCT.*", "", seq)) %>% 
  mutate(seq = gsub(paste(promoters, collapse = "|"), "", seq)) %>%
  unique()

promoters <- data.frame(TF = c("minP", "mCMV", "hBGm", "Random"),
                        seq = promoters)

cDNA_df_native <- rbind(cDNA_df_native, promoters) %>%
  setnames(c("TF", "seq"), c("seq.name", "seq.text"))

#dat2fasta(cDNA_df_native, outfile = "/DATA/usr/m.trauernicht/projects/tf_activity_reporter/data/SuRE_TF_1/results/native-enhancer/cDNA_df_native.fasta")  

```


```{bash run fimo db2_1, eval = FALSE}
# motfn=/home/f.comoglio/mydata/Annotations/TFDB/Curated_Natoli/update_2017/20170320_pwms_selected.meme
# odir=/home/m.trauernicht/mydata/projects/tf_activity_reporter/data/SuRE_TF_1/results/native-enhancer/fimo
# query=/home/m.trauernicht/mydata/projects/tf_activity_reporter/data/SuRE_TF_1/results/native-enhancer/cDNA_df_native.fasta

# nice -n 19 fimo --no-qvalue --thresh 1e-4 --verbosity 1 --o $odir $motfn $query 
```


```{r build tf motif matrices db2: random_promoter, out.width= "80%", fig.align= "center", echo=FALSE, warning= FALSE, message = FALSE}
# # load motif Metadata --> PWM feature matrix
# tib_pwm_native_enhancer  <- get_pwm_feature_matrix(motif_meta_fn = '/home/f.comoglio/mydata/Annotations/TFDB/Curated_Natoli/update_2017/fc181127_curated_metadata_no_composite_filt.csv', 
#                                         fimo_fn       = '/home/m.trauernicht/mydata/projects/tf_activity_reporter/data/SuRE_TF_1/results/native-enhancer/fimo/fimo.tsv',
#                                         db            = 2)
```


```{r out.width= "80%", fig.align= "center", echo=FALSE, warning= FALSE, message = FALSE}
# ## Add groups so I can plot them individually
# tib_pwm_native_enhancer$group <- gsub("(^e[0-9]{1,2}).*", "\\1", tib_pwm_native_enhancer$id)
# controls <- c("minP", "mCMV", "hBGm")
# tib_pwm_native_enhancer$group <- gsub(paste(controls, collapse = "|"), "Control", tib_pwm_native_enhancer$group)
```


```{r out.width= "80%", fig.align= "center", echo=FALSE}
# load('/DATA/usr/m.martinez.ara/GitLab/gurten/gurten/data/fc181121_epsure_tadec_rnaseq_expr_joshi.RData')
# 
# # Take mean of 2i expression
# tib_fpkm_joshi <- tib_fpkm_joshi %>%
#   dplyr::select(-contains('serum')) %>%
#   rowwise() %>%
#   mutate(expr_mean = mean(c(twoi_1, twoi_2))) %>%
#   ungroup()
```

```{r out.width= "80%", fig.align= "center", echo=FALSE, warning= FALSE, message = FALSE}
# # only select fpkm>0
# tib_fpkm_joshi_expressed <- tib_fpkm_joshi %>% filter(expr_mean>10)
# 
# # Change to mouse name of Tp53
# tib_fpkm_joshi_expressed$symbol <- gsub("TRP53", "TP53", tib_fpkm_joshi_expressed$symbol)
# 
# tib_pwm_native_enhancer_ES <- tib_pwm_native_enhancer[,colnames(tib_pwm_native_enhancer) %in% tib_fpkm_joshi_expressed$symbol | colnames(tib_pwm_native_enhancer) %in% c("id", "group")]
```


```{r, fig.height=10, fig.width=10, out.width= "80%", fig.align= "center", echo=FALSE, warning= FALSE, message = FALSE}
# # remove columns with only 0s and plot heatmap for each native enhancer individually
# for (i in unique(tib_pwm_native_enhancer_ES$group)) {
#   data <- tib_pwm_native_enhancer_ES[tib_pwm_native_enhancer_ES$group == i,] %>% 
#     column_to_rownames("id") %>% dplyr::select(-group)
#   p <- pheatmap(as.matrix(data[, colSums(data != 0) > 0]),
#                 color = colorRampPalette(brewer.pal(n = 7, name = "Oranges"))(100),
#                 border_color = "black",
#                 cellwidth = 8, cellheight = 10)
#   print(p)
# }
# 
# ## Zic2/Hic2 bind to mCMV/hBGm -> Does this explain increased cooperativity?
# 
# # make list of relevant TFs
# relevant_TFs <- c("KLF", "SP1", "SP3", "POU", "RAR", "SOX", "STAT", "^TCF", "TFCP", "TP53")
# relevant_TFs_2 <- c("KLF4", "POU", "SOX", "STAT", "^TCF", "TFCP")
# relevant_TFs_3 <- "TFCP"
# 
# tib_pwm_native_enhancer_ES <- tib_pwm_native_enhancer_ES[-grep(paste(promoters$TF, collapse = "|"), tib_pwm_native_enhancer_ES$id),]
# 
# data <- tib_pwm_native_enhancer_ES %>%
#   dplyr::select(KLF4, POU5F1, SOX2, TCF7, TFCP2L1, CTCF, ZIC2,id, group) %>% 
#   column_to_rownames("id") %>% dplyr::select(-group)
# 
# pheatmap(as.matrix(t(data[, colSums(data != 0) > 0])),
#                 color = colorRampPalette(brewer.pal(n = 7, name = "Greys"))(100),
#                 border_color = "black", cluster_cols = F, 
#                 cellwidth = 10, cellheight = 10)

```


```{r fimo_model, out.width= "80%", fig.align= "center", echo=FALSE, warning= FALSE, message = FALSE}
# # Subselect interesting TFs
# tib_pwm_native_enhancer_ES <- tib_pwm_native_enhancer_ES %>%
#   dplyr::select(KLF4, POU5F1, SOX2, TCF7, TFCP2L1, CTCF, ZIC2,id, group)
# 
# # Transform to binary
# tib_pwm_native_enhancer_ES <- tib_pwm_native_enhancer_ES %>% mutate_if(is.numeric, function(x) ifelse((x==0 | is.na(x)),"No","Yes"))
# 
# # Combine with expression data
# native_expression <- tf_activity_heatmap %>% rownames_to_column("condition") %>%
#   melt(id.vars = "condition", variable.name = "id", value.name = "expression")
# 
# tib_pwm_native_enhancer_ES_merge <- merge(tib_pwm_native_enhancer_ES, native_expression, all = T) %>%
#   dplyr::select(-group)
# 
# 
# 
# ### Calculate expression variance
# exp_variance <- data.frame(c("KLF4", "POU5F1", 
#                              "SOX2", "TCF7", "TFCP2L1", "CTCF", "ZIC2", "Residuals"))
# prediction <- data.frame(1:280)
# names(prediction) <- "id"
# names(exp_variance) <- "feature"
# weight <- data.frame(c("(Intercept)", "KLF4Yes", "POU5F1Yes", 
#                              "SOX2Yes", "TCF7Yes", "TFCP2L1Yes", "CTCFYes", "ZIC2Yes"))
# names(weight) <- "feature"
# 
# 
# tib_pwm_native_enhancer_ES_merge$row <- rownames(tib_pwm_native_enhancer_ES_merge)
# 
# 
# 
# for (i in unique(tib_pwm_native_enhancer_ES_merge$condition)) {
#   x <- lm(expression ~ KLF4 + POU5F1 + SOX2 + TCF7 + TFCP2L1 + CTCF + ZIC2, 
#         tib_pwm_native_enhancer_ES_merge[tib_pwm_native_enhancer_ES_merge$condition == i,])
#   y <- data.frame(x$fitted.values) %>% rownames_to_column()
#   par(mfrow=c(2,2))
#   plot(x, main = i)
#   names(y) <- c("row", i)
#   y$id <- 1:nrow(y)
#   prediction <- merge(prediction,y, all = T)
#   
#  
#   w <- data.frame(x$coefficients) %>% rownames_to_column()
#   names(w) <- c("feature", i)
#   weight <- merge(weight,w, all = T)
#   
#   x <- anova(x) %>% rownames_to_column("feature") %>%
#     setnames("Sum Sq", "sum_sq") %>%
#     dplyr::select(feature, sum_sq) %>%
#     mutate(sum = sum(sum_sq)) %>%
#     mutate(rel_sum_sq = (sum_sq/sum)*100) %>%
#     setnames("rel_sum_sq", i)
#   x <- x[,c(1,4)]
# 
#   exp_variance <- merge(exp_variance, x)
# }
# 
# 
# 
# ## Prepare lm parameters for plotting
# exp_variance <- melt(exp_variance, variable.name = "id")
# exp_variance$group <- "no residual"
# exp_variance$group[grep("Residual", exp_variance$feature)] <- "residual"
# prediction <- prediction %>% dplyr::select(-id) %>% melt(variable.name = "id") %>% na.omit() %>%
#   dplyr::select(-id) %>% setnames("value", "activity_predicted")
# tib_pwm_native_enhancer_ES_merge <- merge(tib_pwm_native_enhancer_ES_merge, prediction)
# weight <- weight %>% filter(feature != "(Intercept)")
# weight <- melt(weight, variable.name = "id")
# categorie <- c("KLF4", "POU5F1", "SOX2", "TCF7", "TFCP2L1", "CTCF", "ZIC2")
# #weight$features <- gsub(paste(categorie, collapse="|"), "",weight$feature)
# #feature <- weight$features
# weight$condition <- gsub(paste(feature, collapse="|"), "", weight$feature)
# weight$value[is.na(weight$value)] <- 0
# 
# 
# # Further plotting preparations
# level_order <- c("KLF4Yes", "POU5F1Yes", "SOX2Yes", "CTCFYes", "ZIC2Yes", "TFCP2L1Yes", "TCF7Yes")
# 
# 
# # Model parameters
# exp_variance$value[exp_variance$feature == "Residuals"] <- 100 - exp_variance$value[exp_variance$feature == "Residuals"]
# exp_variance$feature <- revalue(exp_variance$feature, c("Residuals"="total"))
# 
# 
# 
# ggplot(exp_variance, 
#        aes(x = reorder(feature, -value), y = value, fill = group)) +
#   scale_fill_manual(values = c("#1B998B", "#2D3047")) +
#   geom_bar(stat = "identity") + 
#   ylab("Variance explained (%)") + xlab("") + 
#   labs(title = "Linear variance modelling", 
#        subtitle = "Expression variance predicted by features") +
#   theme_bw() +
#   theme(text = element_text(size = 12)) +
#   theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0, size = 12), 
#         axis.text.y = element_text(size = 10)) +
#     guides(fill = F) + facet_wrap(~id)
# 
# ggscatter(tib_pwm_native_enhancer_ES_merge, x = "expression", y = "activity_predicted",
#           add = "reg.line", 
#           add.params = list(color = "blue", fill = "lightgray"),
#           conf.int = TRUE, ylab = "Predicted log2-activity", size = 0.5,
#           xlab = "Average log2-activity") + theme_bw() + facet_wrap(~condition) +
#   stat_cor(aes(label = ..r.label..), method = "pearson", label.x = -2, label.y = 4)
# 
# 
# 
# for (i in unique(tib_pwm_native_enhancer_ES_merge$condition)) {
#   
#   # Plot to show prediction accuracy
#   sp <- ggscatter(tib_pwm_native_enhancer_ES_merge[tib_pwm_native_enhancer_ES_merge$condition == i,], x = "expression", y = "activity_predicted",
#    add = "reg.line", 
#    add.params = list(color = "blue", fill = "lightgray"),
#    conf.int = TRUE, ylab = "Predicted log2-activity", 
#    title = i,
#    xlab = "Average log2-activity") + theme_bw()
#   
#   
#   # Feature importance overview for each TF (including mean activities per feature)
#   fp <- ggplot(exp_variance[exp_variance$id == i,], 
#        aes(x = reorder(feature, -value), y = value, fill = group)) +
#   scale_fill_manual(values = c("#1B998B", "#2D3047")) +
#   geom_bar(stat = "identity") + 
#   ylab("Variance explained (%)") + xlab("") +
#   ggtitle(i)+
#   theme_bw() +
#   theme(text = element_text(size = 12)) +
#   theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0, size = 12), 
#         axis.text.y = element_text(size = 10)) +
#     guides(fill = F)
#   
#   # Plot weights of parameters in model - this helps to understand the model 
#   p <- ggplot(weight[weight$id == i,], 
#        aes(x = factor(feature, level = level_order), y = value, fill = condition)) +
#   scale_fill_manual(values = c("#DD6B48", "#1B998B", "#1B998B", "#1B998B", "#1B998B", "#1B998B", "1B998B")) +
#   ggtitle(i)+
#   geom_bar(stat = "identity") + 
#   ylab("Weight") + xlab("") +
#   theme_bw() +
#   theme(text = element_text(size = 12)) +
#   theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 1, size = 10), 
#         axis.text.y = element_text(size = 10)) +
#     guides(fill = F) 
# 
#   # Combine all plots
#   sp.plot <- arrangeGrob(sp,fp,p, nrow = 2, ncol = 2)
#   grid.arrange(sp.plot)
# }
```






### Heatmap per TF - comparing design activities mutated vs. non-mutated
```{r heatmap_4, out.width= "80%", fig.align= "center", echo=FALSE, warning= FALSE, message = FALSE}
## Prepare dataframe
# Caculate mean TF activity per condition
cDNA_df$reporter <- gsub("_[0-9]{1}$", "", cDNA_df$reporter_id)
cDNA_df$tf_design_activity <- ave(cDNA_df$log_activity, 
                                  cDNA_df$condition, 
                                  cDNA_df$reporter,
                                  FUN = function(x) mean(x))


tf_activity_heatmap <- cDNA_df[cDNA_df$native_enhancer == "No" & 
                                 cDNA_df$rand_promoter == "No" &
                                 cDNA_df$hPGK == "No",] %>% 
  dplyr::select(reporter, condition, tf_design_activity) %>% unique()

tf_activity_heatmap$tf_design_activity[!is.finite(tf_activity_heatmap$tf_design_activity)] <- 0
tf_activity_heatmap <- dcast(tf_activity_heatmap, reporter ~ condition,
                             value.var="tf_design_activity")
tf_activity_heatmap$TF <- gsub("(^.*?)_[0-9]{1,2}bp.*", "\\1", tf_activity_heatmap$reporter)
tf_activity_heatmap$TF <- gsub("_neg$", "", tf_activity_heatmap$TF)
tf_activity_heatmap <- tf_activity_heatmap %>% 
  remove_rownames %>% column_to_rownames(var="reporter")
# Keeping the scale in the pheatmap function
myBreaks1 <- seq(-3,3,0.06)

# Manually annotate heatmaps
annotation_col <- data.frame("TF" = cDNA_df$reporter[cDNA_df$native_enhancer == "No" & 
                                 cDNA_df$rand_promoter == "No" &
                                 cDNA_df$hPGK == "No"], 
                             "neg_ctrls" = cDNA_df$neg_ctrls[cDNA_df$native_enhancer == "No" & 
                                 cDNA_df$rand_promoter == "No" &
                                 cDNA_df$hPGK == "No"]) %>%
  unique() %>%
  remove_rownames %>% column_to_rownames(var="TF")

annotation_colors = list(
    neg_ctrls = c(No ="#1B998B", Yes = "#2D3047"))


# pheatmap function
for(i in unique(tf_activity_heatmap$TF)){
  pheatmap(as.matrix(t(tf_activity_heatmap[tf_activity_heatmap$TF == i, -8])),
         color = colorRampPalette(rev(brewer.pal(n = 7, name = "RdBu")))(100),
         breaks = myBreaks1, border_color = "#565656", 
         cellwidth = 10, cellheight = 10,
         annotation_col = annotation_col, 
         annotation_colors = annotation_colors)
}
```


### Plotting quasirandom plots instead of heatmaps
```{r}
# Plot activity per reporter
tf_activity_df <- tf_activity_heatmap %>%
  rownames_to_column(var = "reporter") %>%
  dplyr::select(`2i_pos_LIF`, TF, reporter)

tf_activity_df <- tf_activity_df %>%
  mutate(mean = ave(`2i_pos_LIF`, TF, FUN = function(x) mean(x)),
         sd = ave(`2i_pos_LIF`, TF, FUN = function(x) sd(x)))

ggplot(tf_activity_df, aes(x = reorder(TF, -sd), y = `2i_pos_LIF`, color = `2i_pos_LIF`)) +
  geom_quasirandom() +
  theme_bw() +
  geom_hline(yintercept = 1, linetype = "dashed") +
  ylab("reporter activity (log2)") +
  coord_flip() +
  scale_color_viridis()

ggplot(tf_activity_df, aes(x = reorder(TF, -mean), y = `2i_pos_LIF`)) +
  geom_bar(data = tf_activity_df %>% dplyr::select(TF, mean, sd) %>% unique(), 
           mapping = aes(x = reorder(TF, -sd), y = sd), stat = "identity", alpha = 0.5) +
  theme_bw() +
  ylim(0,4) +
  ylab("SD reporter activity") +
  coord_flip()

ggplot(tf_activity_df %>% filter(TF == "Stat3"), aes(x = TF, y = `2i_pos_LIF`, color = `2i_pos_LIF`)) +
  geom_quasirandom(width = 0.1) +
  theme_bw() +
  geom_hline(yintercept = 1, linetype = "dashed") +
  ylab("reporter activity (log2)") +
  coord_flip() +
  scale_color_viridis()



# Plot mean activity per TF vs SD within reporters
ggplot(tf_activity_df, aes(x = mean, y = sd)) +
  geom_point() +
  theme_bw()
```




### Heatmap per TF - only WT TF activities
```{r heatmap_5, out.width= "80%", fig.align= "center", echo=FALSE, warning= FALSE, message = FALSE}
## Prepare dataframe
# Caculate mean TF activity per condition
tf_activity_heatmap <- cDNA_df[cDNA_df$native_enhancer == "No" & 
                                 cDNA_df$hPGK == "No" &
                                 cDNA_df$neg_ctrls == "No",] %>% 
  dplyr::select(reporter, condition, tf_design_activity) %>% unique()

tf_activity_heatmap$tf_design_activity[!is.finite(tf_activity_heatmap$tf_design_activity)] <- 0
tf_activity_heatmap <- dcast(tf_activity_heatmap, reporter ~ condition,
                             value.var="tf_design_activity")
tf_activity_heatmap$TF <- gsub("(^.*?)_[0-9]{1,2}bp.*", "\\1", tf_activity_heatmap$reporter)
tf_activity_heatmap <- tf_activity_heatmap %>% 
  remove_rownames %>% column_to_rownames(var="reporter")
# Keeping the scale in the pheatmap function
myBreaks1 <- seq(-3,3,0.06)

# pheatmap function
for(i in unique(tf_activity_heatmap$TF)){
  pheatmap(as.matrix(tf_activity_heatmap[tf_activity_heatmap$TF == i,-8]),
         color = colorRampPalette(rev(brewer.pal(n = 7, name = "RdBu")))(100),
         breaks = myBreaks1, border_color = "#565656", 
         cellwidth = 30, cellheight = 10)
}
```




### Compute activity changes relative to their negative controls
```{r, out.width= "80%", fig.align= "center", echo=FALSE, warning= FALSE, message = FALSE}
## TF activity relative to negative controls 
cDNA_df_ctrl <- cDNA_df[cDNA_df$native_enhancer == "No" &
                          cDNA_df$hPGK == "No" & cDNA_df$TF != "Random1" &
                          cDNA_df$TF != "Random2" & cDNA_df$TF != "Random3",] %>% 
  mutate(reporter_2 = gsub("_neg", "", reporter_id)) %>%
  dplyr::select(reporter_2, condition, reporter_activity, neg_ctrls) %>% 
  mutate(reporter_activity = ave(reporter_activity, reporter_2, condition,
                                 neg_ctrls, FUN = function(x) mean(x))) %>%
  unique()
  

## Divide TF activity by negative control activity for each reporter
for (i in unique(cDNA_df_ctrl$reporter_2)) {
  for (j in unique(cDNA_df_ctrl$condition)) {
    if (length(cDNA_df_ctrl$reporter_activity[cDNA_df_ctrl$reporter_2 == i & 
                                     cDNA_df_ctrl$neg_ctrls == "Yes" &
                                     cDNA_df_ctrl$condition == j]) != 0 &
       length(cDNA_df_ctrl$reporter_activity[cDNA_df_ctrl$reporter_2 == i & 
                                    cDNA_df_ctrl$neg_ctrls == "No" &
                                     cDNA_df_ctrl$condition == j]) != 0) { 
        
    cDNA_df_ctrl$reporter_activity_relative[cDNA_df_ctrl$reporter_2 == i & 
                                     cDNA_df_ctrl$condition == j] <- 
      
     cDNA_df_ctrl$reporter_activity[cDNA_df_ctrl$reporter_2 == i & 
                             cDNA_df_ctrl$neg_ctrls == "No" & 
                              cDNA_df_ctrl$condition == j] /
      
      cDNA_df_ctrl$reporter_activity[cDNA_df_ctrl$reporter_2 == i & 
                              cDNA_df_ctrl$neg_ctrls == "Yes" & 
                               cDNA_df_ctrl$condition == j]
   }
    else {
     cDNA_df_ctrl$reporter_activity_relative[cDNA_df_ctrl$reporter_2 == i & 
                                       cDNA_df_ctrl$condition == j] <- 0
    }
  }
}

## Log2 transformation
cDNA_df_ctrl <- cDNA_df_ctrl %>% 
  mutate(log_reporter_activity_relative = log2(reporter_activity_relative))
  
cDNA_df_ctrl$log_reporter_activity_relative[!is.finite(cDNA_df_ctrl$log_reporter_activity_relative)] <- 0

cDNA_df_ctrl$TF <- gsub("(^.*?)_[0-9]{1,2}bp.*", "\\1", cDNA_df_ctrl$reporter_2)

## Generate heatmaps with relative log2 activities 
# Caculate mean TF activity per condition
tf_activity_heatmap <- cDNA_df_ctrl[cDNA_df_ctrl$neg_ctrls == "No",] %>% 
  dplyr::select(reporter_2, TF, condition, log_reporter_activity_relative) %>% unique()
tf_activity_heatmap2 <- tf_activity_heatmap %>% dplyr::select(-reporter_2) %>%
  mutate(log_reporter_activity_relative = 
           ave(log_reporter_activity_relative, condition, TF, FUN = function(x) mean(x))) %>%
  unique()
tf_activity_heatmap2 <- tf_activity_heatmap2 %>% 
  dcast(TF ~ condition, value.var = "log_reporter_activity_relative")
tf_activity_heatmap2 <- tf_activity_heatmap2 %>% 
  remove_rownames %>% column_to_rownames(var="TF")
tf_activity_heatmap <- tf_activity_heatmap %>% 
  mutate(log_reporter_activity_relative = 
           ave(log_reporter_activity_relative, reporter_2, condition, FUN = function(x) mean(x))) %>%
  unique()
tf_activity_heatmap <- tf_activity_heatmap %>% 
  dcast(reporter_2+TF ~ condition, value.var = "log_reporter_activity_relative")
tf_activity_heatmap <- tf_activity_heatmap %>% 
  remove_rownames %>% column_to_rownames(var="reporter_2")
# Keeping the scale in the pheatmap function
myBreaks1 <- seq(0, 3.5, 0.035)


pheatmap(as.matrix(t(tf_activity_heatmap2)),
         color = colorRampPalette(brewer.pal(n = 7, name = "Greys"))(100),
         breaks = myBreaks1, border_color = "#565656", 
         cellwidth = 12, cellheight = 10,
         angle_col = 45)

myBreaks1 <- seq(0, 3.5, 0.035)

# pheatmap function
tf_activity_heatmap <- tf_activity_heatmap[,c(1,7,3,2,4,5,6,8)]
for(i in unique(tf_activity_heatmap$TF)){
  pheatmap(as.matrix(t(tf_activity_heatmap[tf_activity_heatmap$TF == i,-1])),
         color = colorRampPalette(brewer.pal(n = 7, name = "Greys"))(100),
         border_color = "black",
         breaks = myBreaks1, cluster_rows = F,
         cellwidth = 10, cellheight = 10)
}
```
*All of these heatmaps conclude that there we have informative reporters for ~10 TFs, and that the TF reporter design matters for some but not all TFs *


```{r SuperPlot, out.width= "80%", fig.align= "center", echo=FALSE, warning= FALSE, message = FALSE}
# # Plot for each tf-activity per condition per biological replicate the techical replicate distribution and means
# ## Compute relevant mean and sd values
# activity_mean <- cDNA_df %>% 
#   dplyr::select(activity, condition, rep, TF, promoter, neg_ctrls, hPGK, native_enhancer, rand_promoter) %>% 
#   group_by(condition, rep, TF, promoter) %>% 
#   mutate(activity = mean(activity))
# 
# activity_mean <- activity_mean %>% 
#   group_by(condition, TF, promoter) %>% 
#   mutate(mean = mean(activity),
#          se = sd(activity)/sqrt(length(unique((activity)))))
# 
# activity_mean <- activity_mean %>% 
#   mutate(upper = mean + se, lower = mean - se)
# 
# 
# ## All TFs
# activity_mean_1 <- activity_mean[activity_mean$promoter == "minP" & 
#                         activity_mean$neg_ctrls == "No" &
#                         activity_mean$native_enhancer == "No",] %>% unique()
# 
# ## hPGK controls
# activity_mean_2 <- activity_mean[activity_mean$promoter == "hPGK" & 
#                         activity_mean$neg_ctrls == "No" & 
#                         activity_mean$native_enhancer == "No" ,] %>% unique()
# 
# ## Native enhancers
# activity_mean_3 <- activity_mean[activity_mean$promoter == "minP" & 
#                         activity_mean$native_enhancer == "Yes",] %>% unique()
# 
# 
# ## Random promoters
# activity_mean_4 <- activity_mean[activity_mean$promoter == "Random" & 
#                         activity_mean$neg_ctrls == "No" & 
#                         activity_mean$native_enhancer == "No",] %>% unique()
# 
# ## Negative controls
# activity_mean_5 <- activity_mean[activity_mean$promoter == "minP" & 
#                         activity_mean$neg_ctrls == "Yes" & 
#                         activity_mean$native_enhancer == "No",] %>% unique()
# 
# 
# ## Generate SuperPlot
# ### Selection 1 - All TFs
# for (i in 1:4) {
#   p <- ggplot(cDNA_df[cDNA_df$neg_ctrls == "No" &
#                         cDNA_df$native_enhancer == "No",], 
#               aes(x=condition, y=activity, color=factor(rep))) + 
#     scale_color_manual(values = c("#1B998B", "#2D3047", "#FF9B71")) +
#     geom_quasirandom(aes(alpha = 0.9), dodge.width = 0.4) +
#     geom_errorbar(data = activity_mean_1, 
#                   aes(ymin = lower, 
#                       ymax = upper), colour = "black") + 
#     geom_quasirandom(data = activity_mean_1, aes(alpha = 1), size=4) + 
#     stat_compare_means(data=activity_mean_1, 
#                        comparisons = list(c(1:2)), 
#                    method="t.test", paired=TRUE) + 
#     labs(title = "TF activity distribution per condition", 
#          subtitle = paste("Page", i)) +
#     facet_wrap_paginate(~TF, ncol = 3, nrow = 3, page = i) +
#     theme_bw() +
#     ylim(0,6) +
#     theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10), 
#           axis.text.y = element_text(size = 10)) +
#     ylab("TF reporter activity") + xlab("") + 
#     theme(text = element_text(size = 12))
#   print(p)
# }
# 
# 
# 
# ### Selection 2 - hPGK controls
# ggplot(cDNA_df[cDNA_df$promoter == "hPGK" & 
#                         cDNA_df$neg_ctrls == "No" &
#                         cDNA_df$native_enhancer == "No",], 
#               aes(x=condition, y=activity, color=factor(rep))) + 
#     scale_color_manual(values = c("#1B998B", "#2D3047", "#FF9B71")) +
#     geom_quasirandom(aes(alpha = 0.9), dodge.width = 0.4) +
#     geom_errorbar(data = activity_mean_2, 
#                   aes(ymin = lower, 
#                       ymax = upper), colour = "black") + 
#     geom_quasirandom(data = activity_mean_2, aes(alpha = 1), size=4) + 
#     stat_compare_means(data=activity_mean_2, 
#                        comparisons = list(c(1:2)), 
#                    method="t.test", paired=TRUE) + 
#     labs(title = "hPGK controls: activity distribution per condition", 
#          subtitle = p) +
#     theme_bw() +
#     theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10), 
#           axis.text.y = element_text(size = 10)) +
#     ylab("TF reporter activity") + xlab("") + 
#     theme(text = element_text(size = 12))
# 
# 
# 
# ### Selection 3 - Native enhancers
# for (i in 1:5) {
#   p <- ggplot(cDNA_df[cDNA_df$native_enhancer == "Yes",], 
#               aes(x=condition, y=activity, color=factor(rep))) + 
#     scale_color_manual(values = c("#1B998B", "#2D3047", "#FF9B71")) +
#     geom_quasirandom(aes(alpha = 0.9), dodge.width = 0.4) +
#     geom_errorbar(data = activity_mean_3, 
#                   aes(ymin = lower, 
#                       ymax = upper), colour = "black") + 
#     geom_quasirandom(data = activity_mean_3, aes(alpha = 1), size=4) + 
#     stat_compare_means(data=activity_mean_3, 
#                        comparisons = list(c(1:2)), 
#                    method="t.test", paired=TRUE) + 
#     labs(title = "Miguels native enhancer chunks per condition", 
#          subtitle = paste("Page", i)) +
#     facet_wrap_paginate(~TF, ncol = 3, nrow = 3, page = i) +
#     ylim(0,6) +
#     theme_bw() +
#     theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10), 
#           axis.text.y = element_text(size = 10)) +
#     ylab("TF reporter activity") + xlab("") + 
#     theme(text = element_text(size = 12))
#   print(p)
# }
# 
# 
# ### Selection 5 - All TF-negative controls
# for (i in 1:3) {
#   p <- ggplot(cDNA_df[cDNA_df$promoter == "minP" & 
#                         cDNA_df$neg_ctrls == "Yes" &
#                         cDNA_df$native_enhancer == "No",], 
#               aes(x=condition, y=activity, color=factor(rep))) + 
#     scale_color_manual(values = c("#1B998B", "#2D3047", "#FF9B71")) +
#     geom_quasirandom(aes(alpha = 0.9), dodge.width = 0.4) +
#     geom_errorbar(data = activity_mean_5, 
#                   aes(ymin = lower, 
#                       ymax = upper), colour = "black") + 
#     geom_quasirandom(data = activity_mean_5, aes(alpha = 1), size=4) + 
#     stat_compare_means(data=activity_mean_5, 
#                        comparisons = list(c(1:2)), 
#                    method="t.test", paired=TRUE) + 
#     labs(title = "TF activity distribution per condition - only negative controls", 
#          subtitle = paste("Page", i)) +
#     facet_wrap_paginate(~TF, ncol = 3, nrow = 3, page = i) +
#     theme_bw() +
#     ylim(0,6) +
#     theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10), 
#           axis.text.y = element_text(size = 10)) +
#     ylab("TF reporter activity") + xlab("") + 
#     theme(text = element_text(size = 12))
#   print(p)
# }
```


```{r superplot_2, out.width= "80%", fig.align= "center", echo=FALSE, warning= FALSE, message = FALSE}
# # Same as before - now compare different designs
# ## Generate facet_wrap plot for each TF - each facet is a different design 
# ## Compute values for single and combinatorial designs
# 
# 
# ## Compute relevant mean and sd values
# activity_mean <- cDNA_df %>% 
#   dplyr::select(activity, condition, rep, TF, promoter, neg_ctrls, 
#          hPGK, native_enhancer, rand_promoter, reporter_id, spacing, 
#          distance, background) %>% 
#   group_by(condition, rep, reporter_id) %>% 
#   mutate(activity = mean(activity))
# 
# activity_mean <- activity_mean %>% 
#   group_by(condition, reporter_id) %>% 
#   mutate(mean = mean(activity),
#          se = sd(activity)/sqrt(length(unique((activity)))))
# 
# activity_mean <- activity_mean %>% 
#   mutate(upper = mean + se, lower = mean - se)
# 
# activity_mean <- activity_mean[activity_mean$neg_ctrls == "No" &
#                         activity_mean$native_enhancer == "No",] %>% unique()
# 
# 
# ## Generate SuperPlot
# ### Selection 1 - All Designs separately
# for (i in 1:4) {
#   p <- ggplot(cDNA_df[cDNA_df$TF == "Trp53" & 
#                         cDNA_df$neg_ctrls == "No" &
#                         cDNA_df$native_enhancer == "No",], 
#               aes(x=condition, y=activity, color=factor(rep))) + 
#     scale_color_manual(values = c("#1B998B", "#2D3047", "#FF9B71")) +
#     geom_quasirandom(aes(alpha = 0.9), dodge.width = 0.4) +
#     geom_errorbar(data = activity_mean[activity_mean$TF == "Trp53",], 
#                   aes(ymin = lower, 
#                       ymax = upper), colour = "black") + 
#     geom_quasirandom(data = activity_mean[activity_mean$TF == "Trp53",], aes(alpha = 1), size=4) + 
#     stat_compare_means(data=activity_mean[activity_mean$TF == "Trp53",], 
#                        comparisons = list(c(1:2)), 
#                    method="t.test", paired=TRUE) + 
#     labs(title = "Trp53 activity distribution per design and condition", 
#          subtitle = paste("Page", i)) +
#     facet_wrap_paginate(~reporter_id, ncol = 3, nrow = 3, page = i) +
#     theme_bw() +
#     ylim(0,16) +
#     theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10), 
#           axis.text.y = element_text(size = 10)) +
#     ylab("TF reporter activity") + xlab("") + 
#     theme(text = element_text(size = 12))
#   print(p)
# }
# 
# 
# ### All reporters next to each other
# ggplot(cDNA_df[cDNA_df$TF == "Trp53" & 
#                         cDNA_df$neg_ctrls == "No" &
#                         cDNA_df$native_enhancer == "No",], 
#               aes(x=reporter_id, y=log_activity)) + 
#     scale_color_manual(values = c("#1B998B", "#2D3047", "#FF9B71")) +
#     geom_boxplot() +
#     labs(title = "Trp53 activity distribution per design and condition") +
#     theme_bw() +
#     ylim(-2,6) +
#     theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10), 
#           axis.text.y = element_text(size = 10)) +
#     ylab("TF reporter activity") + xlab("") + 
#     theme(text = element_text(size = 12))
# 
# 
# 
# 
# 
# ## Subdivide in different designs 
# ### Spacing and distance
# cDNA_df2 <- cDNA_df %>% 
#   mutate(space_dist = paste(spacing, distance)) %>% 
#   dplyr::select(activity, condition, TF, space_dist, neg_ctrls, native_enhancer, promoter) 
# 
# cDNA_df2 <- cDNA_df2[cDNA_df2$neg_ctrls == "No" &
#                        cDNA_df2$native_enhancer == "No" & 
#                        cDNA_df2$promoter != "Random" &
#                        cDNA_df2$promoter != "hPGK",] %>% unique()
# 
# 
# 
# for(i in 1:8){
#   p <- ggplot(cDNA_df2, 
#         aes(x=condition, y=activity, color=space_dist)) + 
#       scale_color_manual(values = c("#1B998B", "#2D3047", "#FF9B71", "#ECDD7B")) +
#       geom_quasirandom(dodge.width = 0.9) +
#       geom_boxplot(position = position_dodge(0.9), aes(alpha = 0.5)) + 
#       labs(title = "Comparing spacings: TF activity distribution per condition", 
#            subtitle = paste("Page", i)) +
#       facet_wrap_paginate(~TF, nrow = 2, ncol = 2, page = i) +
#       theme_bw() +
#       ylim(0,6) +
#       theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10), 
#            axis.text.y = element_text(size = 10)) +
#       ylab("TF reporter activity") + xlab("") + 
#       theme(text = element_text(size = 12))
#   print(p)
# }
# 
# 
# 
# ### Background
# cDNA_df2 <- cDNA_df %>% 
#   dplyr::select(activity, condition, TF, background, neg_ctrls, native_enhancer, promoter) 
# 
# 
# cDNA_df2 <- cDNA_df2[cDNA_df2$neg_ctrls == "No" &
#                        cDNA_df2$native_enhancer == "No" & 
#                        cDNA_df2$promoter != "Random" &
#                        cDNA_df2$promoter != "hPGK",] %>% unique()
# 
# 
# 
# for(i in 1:8){
#   p <- ggplot(cDNA_df2, 
#         aes(x=condition, y=activity, color=background)) + 
#       scale_color_manual(values = c("#1B998B", "#2D3047", "#FF9B71", "#ECDD7B")) +
#       geom_quasirandom(dodge.width = 0.9) +
#       geom_boxplot(position = position_dodge(0.9), aes(alpha = 0.5)) + 
#       labs(title = "Comparing backgrounds: TF activity distribution per condition", 
#            subtitle = paste("Page", i)) +
#       facet_wrap_paginate(~TF, nrow = 2, ncol = 2, page = i) +
#       theme_bw() +
#       ylim(0,6) +
#       theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10), 
#            axis.text.y = element_text(size = 10)) +
#       ylab("TF reporter activity") + xlab("") + 
#       theme(text = element_text(size = 12))
#   print(p)
# }
# 
# 
# ### Promoter
# cDNA_df2 <- cDNA_df %>% 
#   dplyr::select(activity, condition, TF, promoter, neg_ctrls, native_enhancer, promoter) 
# 
# 
# cDNA_df2 <- cDNA_df2[cDNA_df2$neg_ctrls == "No" &
#                        cDNA_df2$native_enhancer == "No" &
#                        cDNA_df2$promoter != "hPGK",] %>% unique()
# 
# 
# 
# for(i in 1:8){
#   p <- ggplot(cDNA_df2, 
#         aes(x=condition, y=activity, color=promoter)) + 
#       scale_color_manual(values = c("#1B998B", "#2D3047", "#FF9B71", "#ECDD7B")) +
#       geom_quasirandom(dodge.width = 0.9) +
#       geom_boxplot(position = position_dodge(0.9), aes(alpha = 0.5)) + 
#       labs(title = "Comparing promoters: TF activity distribution per condition", 
#            subtitle = paste("Page", i)) +
#       facet_wrap_paginate(~TF, nrow = 2, ncol = 2, page = i) +
#       theme_bw() +
#       ylim(0,6) +
#       theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10), 
#            axis.text.y = element_text(size = 10)) +
#       ylab("TF reporter activity") + xlab("") + 
#       theme(text = element_text(size = 12))
#   print(p)
# }
# 
# 
# ### Promoter - native enhancers
# cDNA_df2 <- cDNA_df %>% 
#   dplyr::select(activity, condition, TF, background, neg_ctrls, native_enhancer, promoter) 
# 
# 
# cDNA_df2 <- cDNA_df2[cDNA_df2$neg_ctrls == "No" &
#                        cDNA_df2$native_enhancer == "Yes" &
#                        cDNA_df2$promoter != "hPGK",] %>% unique()
# 
# 
# 
# for(i in 1:10){
#   p <- ggplot(cDNA_df2, 
#         aes(x=condition, y=activity, color=promoter)) + 
#       scale_color_manual(values = c("#1B998B", "#2D3047", "#FF9B71", "#ECDD7B")) +
#       geom_quasirandom(dodge.width = 0.9) +
#       geom_boxplot(position = position_dodge(0.9), aes(alpha = 0.5)) + 
#       labs(title = "Miguels native enhancers per promoter", 
#            subtitle = paste("Page", i)) +
#       facet_wrap_paginate(~TF, nrow = 2, ncol = 2, page = i) +
#       theme_bw() +
#       ylim(0,6) +
#       theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10), 
#            axis.text.y = element_text(size = 10)) +
#       ylab("TF reporter activity") + xlab("") + 
#       theme(text = element_text(size = 12))
#   print(p)
# }
```








### Log-linear expression modelling to explain variance - model for each TF
```{r log-linear-model-3, out.width= "80%", fig.align= "center", echo=FALSE, warning= FALSE, message = FALSE}
### For each TF separately - only the non-mutated 
cDNA_df_TF <- cDNA_df[cDNA_df$native_enhancer == "No" &
                        cDNA_df$hPGK == "No" & 
                        cDNA_df$neg_ctrls == "No" &
                        cDNA_df$TF != "Random1"&
                        cDNA_df$TF != "Random2"&
                        cDNA_df$TF != "Random3",] %>% 
  mutate(log_reporter_activity = ave(log_reporter_activity, reporter_id, condition, FUN = function(x) mean(x))) %>%
  dplyr::select(log_reporter_activity, TF, condition, promoter, spacing, distance, background) %>%
  unique()

cDNA_df_rel <- cDNA_df_ctrl[cDNA_df_ctrl$neg_ctrls == "No",] %>% 
  mutate(TF = gsub("(^.*)_[0-9]{1,2}bp_[0-9]{1,2}bp_.*", "\\1", reporter_2),
         promoter = gsub(".*(minP|mCMV|hBGm|Random)_.*", "\\1", reporter_2),
         distance = gsub(".*_([0-9]{1,2}bp)_.*", "\\1", reporter_2),
         spacing = gsub(".*_([0-9]{1,2}bp)_[0-9]{1,2}bp_.*", "\\1", reporter_2),
         background = gsub(".*([0-9]{1}$)", "\\1", reporter_2)) %>%
  dplyr::select(condition, log_reporter_activity_relative, TF, promoter, distance, spacing, background)

cDNA_df_TF <- merge(cDNA_df_TF, cDNA_df_rel, all = T)

### Calculate expression variance
exp_variance <- data.frame(c("condition", "promoter", 
                             "spacing", "distance", "background", "Residuals"))
prediction <- data.frame(1:504)
names(prediction) <- "id"
names(exp_variance) <- "feature"
weight <- data.frame(c("(Intercept)", "condition2i_pos_LIF", 
                  "conditionLIF_pos_CH", "conditionLIF_pos_PD", "conditionN2B27", 
                  "conditionNPC", "conditionvitA", "promotermCMV", "promoterminP", 
                  "promoterRandom",
                  "spacing10bp", "distance21bp", "background2", "background3"))
names(weight) <- "feature"


cDNA_df_TF$row <- rownames(cDNA_df_TF)


for (i in unique(cDNA_df_TF$TF[cDNA_df_TF$TF != "Homez" & cDNA_df_TF$TF != "Irx3" & cDNA_df_TF$TF != "Nr3c1(GR)"])) {
  x <- lm(log_reporter_activity_relative ~ condition + 
                       promoter + spacing + distance + background, 
                     cDNA_df_TF[cDNA_df_TF$TF == i,])
  y <- data.frame(x$fitted.values) %>% rownames_to_column()
  par(mfrow=c(2,2))
  plot(x, main = i)
  names(y) <- c("row", i)
  y$id <- 1:nrow(y)
  prediction <- merge(prediction,y, all = T)
  
  w <- data.frame(x$coefficients) %>% rownames_to_column()
  names(w) <- c("feature", i)
  weight <- merge(weight,w, all = T)
  
  x <- anova(x) %>% rownames_to_column("feature") %>%
    setnames("Sum Sq", "sum_sq") %>%
    dplyr::select(feature, sum_sq) %>%
    mutate(sum = sum(sum_sq)) %>%
    mutate(rel_sum_sq = (sum_sq/sum)*100) %>%
    setnames("rel_sum_sq", i)
  x <- x[,c(1,4)]
  
  exp_variance <- merge(exp_variance, x)
}


## Prepare lm parameters for plotting
exp_variance <- melt(exp_variance, variable.name = "TF")
exp_variance$group <- "no residual"
exp_variance$group[grep("Residual", exp_variance$feature)] <- "residual"
prediction <- prediction %>% dplyr::select(-id) %>% melt(variable.name = "TF") %>% na.omit() %>%
  dplyr::select(-TF) %>% setnames("value", "activity_predicted")
cDNA_df_TF <- merge(cDNA_df_TF, prediction)
weight <- weight %>% filter(feature != "(Intercept)")
weight <- melt(weight, variable.name = "TF")
categorie <- c("mutated", "condition", "promoter", 
               "spacing", "distance", "background")
weight$features <- gsub(paste(categorie, collapse="|"), "",weight$feature)
feature <- weight$features
weight$condition <- gsub(paste(feature, collapse="|"), "", weight$feature)
weight$value[is.na(weight$value)] <- 0


# Further plotting preparations
level_order <- c("2i_pos_LIF", "LIF_pos_CH", 
                 "LIF_pos_PD", "vitA", "N2B27", 
                 "NPC", "mCMV", "minP", "Random",
                 "2", "3", "10bp", "21bp", "5bp")


# Model parameters
exp_variance$value[exp_variance$feature == "Residuals"] <- 100 - exp_variance$value[exp_variance$feature == "Residuals"]
exp_variance$feature <- revalue(exp_variance$feature, c("Residuals"="total"))



ggplot(exp_variance, 
       aes(x = reorder(feature, -value), y = value, fill = group)) +
  scale_fill_manual(values = c("#1B998B", "#2D3047")) +
  geom_bar(stat = "identity") + 
  ylab("Variance explained (%)") + xlab("") + 
  labs(title = "Linear variance modelling", 
       subtitle = "Expression variance predicted by features") +
  theme_bw() +
  theme(text = element_text(size = 12)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0, size = 12), 
        axis.text.y = element_text(size = 10)) +
    guides(fill = F) + facet_wrap(~TF)

ggscatter(cDNA_df_TF, x = "log_reporter_activity", y = "activity_predicted",
          add = "reg.line", 
          add.params = list(color = "blue", fill = "lightgray"),
          conf.int = TRUE, ylab = "Predicted log2-activity", size = 0.5,
          xlab = "Average log2-activity") + theme_bw() + facet_wrap(~TF) + 
  stat_cor(aes(label = ..r.label..), method = "pearson", label.x = -2, label.y = 4)



for (i in unique(cDNA_df_TF$TF)) {
  
  # Plot to show prediction accuracy
  sp <- ggscatter(cDNA_df_TF[cDNA_df_TF$TF == i,], x = "log_reporter_activity", y = "activity_predicted",
   add = "reg.line", 
   add.params = list(color = "blue", fill = "lightgray"),
   conf.int = TRUE, ylab = "Predicted log2-activity", 
   title = i,
   xlab = "Average log2-activity") + theme_bw()
  
  
  # Feature importance overview for each TF (including mean activities per feature)
  fp <- ggplot(exp_variance[exp_variance$TF == i,], 
       aes(x = reorder(feature, -value), y = value, fill = group)) +
  scale_fill_manual(values = c("#1B998B", "#2D3047")) +
  geom_bar(stat = "identity") + 
  ylab("Variance explained (%)") + xlab("") +
  ggtitle(i)+
  theme_bw() +
  theme(text = element_text(size = 12)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0, size = 12), 
        axis.text.y = element_text(size = 10)) +
    guides(fill = F)
  
  # Plot weights of parameters in model - this helps to understand the model 
  p <- ggplot(weight[weight$TF == i,], 
       aes(x = factor(features, level = level_order), y = value, fill = condition)) +
  scale_fill_manual(values = c("#1B998B", "#2D3047", "#FF9B71", "#ECDD7B", "#C2B2B4", "#A54657")) +
  ggtitle(i)+
  geom_bar(stat = "identity") + 
  ylab("Weight") + xlab("") +
  theme_bw() +
  theme(text = element_text(size = 12)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 1, size = 10), 
        axis.text.y = element_text(size = 10)) +
    guides(fill = F) 

  # Combine all plots
  sp.plot <- arrangeGrob(sp,fp,p, nrow = 2, ncol = 2)
  grid.arrange(sp.plot)
}
```


















### Log-linear expression modelling to explain variance - model for each TF - only WT - without condition
```{r log-linear-model-4, out.width= "80%", fig.align= "center", echo=FALSE, warning= FALSE, message = FALSE}
### For each TF separately - only the non-mutated 
cDNA_df_TF <- cDNA_df[cDNA_df$native_enhancer == "No" &
                        cDNA_df$hPGK == "No" & 
                        cDNA_df$neg_ctrls == "No" &
                        cDNA_df$TF != "Random1"&
                        cDNA_df$TF != "Random2"&
                        cDNA_df$TF != "Random3",] %>% 
  mutate(log_reporter_activity = ave(log_reporter_activity, reporter_id, FUN = function(x) mean(x))) %>%
  dplyr::select(log_reporter_activity, TF, promoter, spacing, distance, background) %>%
  unique()

cDNA_df_rel <- cDNA_df_ctrl[cDNA_df_ctrl$neg_ctrls == "No",] %>% 
  mutate(TF = gsub("(^.*)_[0-9]{1,2}bp_[0-9]{1,2}bp_.*", "\\1", reporter_2),
         promoter = gsub(".*(minP|mCMV|hBGm|Random)_.*", "\\1", reporter_2),
         distance = gsub(".*_([0-9]{1,2}bp)_.*", "\\1", reporter_2),
         spacing = gsub(".*_([0-9]{1,2}bp)_[0-9]{1,2}bp_.*", "\\1", reporter_2),
         background = gsub(".*([0-9]{1}$)", "\\1", reporter_2)) %>%
  dplyr::select(log_reporter_activity_relative, TF, promoter, distance, spacing, background) %>% unique() %>%
  mutate(log_reporter_activity_relative = ave(log_reporter_activity_relative, 
                                              TF, promoter, distance, spacing, background, 
                                              FUN = function(x) mean(x)))

cDNA_df_TF <- merge(cDNA_df_TF, cDNA_df_rel, all = T)



### Calculate expression variance
exp_variance <- data.frame(c("promoter", "spacing", "distance", "background", "Residuals"))
prediction <- data.frame(1:504)
names(prediction) <- "id"
names(exp_variance) <- "feature"
weight <- data.frame(c("(Intercept)", "promotermCMV", "promoterminP", 
                  "spacing5bp", "distance21bp", "background2", "background3"))
names(weight) <- "feature"


cDNA_df_TF$row <- rownames(cDNA_df_TF)

for (i in unique(cDNA_df_TF$TF[cDNA_df_TF$TF != "Homez" & cDNA_df_TF$TF != "Irx3" & cDNA_df_TF$TF != "Nr3c1(GR)"])) {
  cDNA_df_TF_2 <- cDNA_df_TF[cDNA_df_TF$TF != "Homez" & cDNA_df_TF$TF != "Irx3" & cDNA_df_TF$TF != "Nr3c1(GR)",]
  x <- lm(log_reporter_activity ~ promoter + spacing + distance + background, 
                     cDNA_df_TF[cDNA_df_TF$TF == i,])
  y <- data.frame(x$fitted.values) %>% rownames_to_column()
  names(y) <- c("row", i)
  y$id <- 1:nrow(y)
  prediction <- merge(prediction,y, all = T)
  
  w <- data.frame(x$coefficients) %>% rownames_to_column()
  names(w) <- c("feature", i)
  weight <- merge(weight,w, all = T)
  
  x <- anova(x) %>% rownames_to_column("feature") %>%
    setnames("Sum Sq", "sum_sq") %>%
    dplyr::select(feature, sum_sq) %>%
    mutate(sum = sum(sum_sq)) %>%
    mutate(rel_sum_sq = (sum_sq/sum)*100) %>%
    setnames("rel_sum_sq", i)
  x <- x[,c(1,4)]
  
  exp_variance <- merge(exp_variance, x)
}


## Fine-tune dfs
exp_variance <- melt(exp_variance, variable.name = "TF")
exp_variance$group <- "no residual"
exp_variance$group[grep("Residual", exp_variance$feature)] <- "residual"
prediction <- prediction %>% dplyr::select(-id) %>% melt(variable.name = "TF") %>% na.omit() %>%
  dplyr::select(-TF) %>% setnames("value", "activity_predicted")
cDNA_df_TF <- merge(cDNA_df_TF, prediction)
weight <- weight %>% filter(feature != "(Intercept)")
weight <- melt(weight, variable.name = "TF")
categorie <- c("mutated", "condition", "promoter", 
               "spacing", "distance", "background")
weight$features <- gsub(paste(categorie, collapse="|"), "",weight$feature)
feature <- weight$features
weight$condition <- gsub(paste(feature, collapse="|"), "", weight$feature)
weight <- weight %>% dplyr::select(-feature)



# Model parameters
exp_variance$value[exp_variance$feature == "Residuals"] <- 100 - exp_variance$value[exp_variance$feature == "Residuals"]
exp_variance$feature <- revalue(exp_variance$feature, c("Residuals"="total"))

ggplot(exp_variance, 
       aes(x = reorder(feature, -value), y = value, fill = group)) +
  scale_fill_manual(values = c("#1B998B", "#2D3047")) +
  geom_bar(stat = "identity") + 
  ylab("Variance explained (%)") + xlab("") + 
  labs(title = "Linear variance modelling", 
       subtitle = "Expression variance predicted by features") +
  theme_bw() +
  theme(text = element_text(size = 12)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0, size = 12), 
        axis.text.y = element_text(size = 10)) +
  facet_wrap(~TF, ncol = 6)




## Plot weights of parameters in model - this helps to understand the model 
level_order <- c("mCMV", "minP","Random", 
                 "2", "3", "5bp", "21bp")





## Final overview - model fit for each TF
# Feature importance overview for each TF in a single plot (bar-chart including mean activities per feature)


for (i in unique(cDNA_df_TF$TF)) {
  
  # Plot to show prediction accuracy
  sp <- ggscatter(cDNA_df_TF[cDNA_df_TF$TF == i,], x = "log_reporter_activity", y = "activity_predicted",
   add = "reg.line", 
   add.params = list(color = "blue", fill = "lightgray"),
   conf.int = TRUE, ylab = "Predicted log2-activity", 
   title = i,
   xlab = "Average log2-activity") + theme_bw()
  
  
  # Feature importance overview for each TF (including mean activities per feature)
  fp <- ggplot(exp_variance[exp_variance$TF == i,], 
       aes(x = reorder(feature, -value), y = value, fill = group)) +
  scale_fill_manual(values = c("#1B998B", "#2D3047")) +
  geom_bar(stat = "identity") + 
  ylab("Variance explained (%)") + xlab("") +
  theme_bw() +
  theme(text = element_text(size = 12)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0, size = 12), 
        axis.text.y = element_text(size = 10)) +
    guides(fill = F)
  
  # Plot weights of parameters in model - this helps to understand the model 
  p <- ggplot(weight[weight$TF == i,], 
       aes(x = factor(features, level = level_order), y = value, fill = condition)) +
  scale_fill_manual(values = c("#1B998B", "#2D3047", "#FF9B71", "#ECDD7B", "#C2B2B4", "#A54657")) +
  geom_bar(stat = "identity") + 
  ylab("Weight") + xlab("") +
  theme_bw() +
  theme(text = element_text(size = 12)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 1, size = 10), 
        axis.text.y = element_text(size = 10)) +
    guides(fill = F) 

  # Combine all plots
  sp.plot <- arrangeGrob(sp,fp,p, nrow = 2, ncol = 2)
  grid.arrange(sp.plot)
}
```




```{r tf-properties, out.width= "80%", fig.align= "center", echo=FALSE, warning= FALSE, message = FALSE}
# # Import tf properties dataframe
# tf_properties <- read.csv2("/DATA/usr/m.trauernicht/projects/tf_activity_reporter/data/SuRE_TF_1/tf_properties.csv", header = T)
# tf_properties_lamb <- read.csv2("/DATA/usr/m.trauernicht/projects/tf_activity_reporter/data/complete_TF_motif_space/lambert_filt.csv", header = T) %>% setnames("Name", "TF") %>% dplyr::select(TF, DBD)
# 
# tf_properties <- merge(tf_properties, tf_properties_lamb)
# 
# convertletter <- function(x) {
#   substr(x, 2, 20) <- tolower(substr(x, 2, 20))
#   x
# }
# 
# tf_properties$TF <- convertletter(tf_properties$TF)
# 
# tf_properties <- tf_properties[tf_properties$TF %in% cDNA_df$TF,]
# 
# variance <- exp_variance[exp_variance$feature == "total",] %>% dplyr::select(TF, value) %>% unique()
# 
# expression <- cDNA_df_TF %>% dplyr::select(log_reporter_activity_relative, TF) %>% 
#   mutate(log_reporter_activity_relative = ave(log_reporter_activity_relative, TF, FUN = function(x) mean(x))) %>%
#   unique()
# 
# tf_properties <- merge(tf_properties, variance)
# tf_properties <- merge(tf_properties, expression)
# 
# x1 <- ggplot(tf_properties, aes(x = chromatin_regulation_mode, y = value)) +
#   geom_quasirandom(width = 0.2) +
#   theme_bw() +
#   xlab("Chromatin regulation mode") +
#   theme(axis.text.x = element_text(angle = 90, hjust = 1.0, vjust = 0))+
#   ylab("TF reporter quality")
# 
# x2 <- ggplot(tf_properties, aes(x = chromatin_regulation_mode, y = log_reporter_activity_relative)) +
#   geom_quasirandom(width = 0.2) +
#   theme_bw() +
#   xlab("Chromatin regulation mode") +
#   theme(axis.text.x = element_text(angle = 90, hjust = 1.0, vjust = 0))+
#   ylab("TF reporter expression")
# 
# 
# #Grid the plots to one panel
# sp.plot <- arrangeGrob(x1,x2, nrow = 1, ncol = 2)
# grid.arrange(sp.plot)
# 
# 
# y1 <- ggplot(tf_properties, aes(x = TF_superclass, y = value)) +
#   geom_quasirandom(width = 0.2) +
#   theme_bw() +
#   theme(axis.text.x = element_text(angle = 90, hjust = 1.0, vjust = 0))+
#   xlab("TF superclass") +
#   ylab("TF reporter quality")
# 
# y2 <- ggplot(tf_properties, aes(x = TF_superclass, y = log_reporter_activity_relative)) +
#   geom_quasirandom(width = 0.2) +
#   theme_bw() +
#   theme(axis.text.x = element_text(angle = 90, hjust = 1.0, vjust = 0))+
#   xlab("TF superclass") +
#   ylab("TF reporter expression")
# 
# sp.plot <- arrangeGrob(y1,y2, nrow = 1, ncol = 2)
# grid.arrange(sp.plot)
# 
# 
# z1 <- ggplot(tf_properties, aes(x = mode_of_regulation, y = value)) +
#   geom_quasirandom(width = 0.2) +
#   theme_bw() +
#   xlab("Mode of regulation") +
#   ylab("TF reporter quality")
# 
# z2 <- ggplot(tf_properties, aes(x = mode_of_regulation, y = log_reporter_activity_relative)) +
#   geom_quasirandom(width = 0.2) +
#   theme_bw() +
#   xlab("Mode of regulation") +
#   ylab("TF reporter expression")
# 
# sp.plot <- arrangeGrob(z1,z2, nrow = 1, ncol = 2)
# grid.arrange(sp.plot)
# 
# 
# x4 <- ggplot(tf_properties, aes(x = tissue_of_expression, y = value)) +
#   geom_quasirandom(width = 0.2) +
#   theme_bw() +
#   theme(axis.text.x = element_text(angle = 90, hjust = 1.0, vjust = 0))+
#   xlab("Tissue specificity") +
#   ylab("TF reporter quality")
# 
# x5 <- ggplot(tf_properties, aes(x = tissue_of_expression, y = log_reporter_activity_relative)) +
#   geom_quasirandom(width = 0.2) +
#   theme_bw() +
#   theme(axis.text.x = element_text(angle = 90, hjust = 1.0, vjust = 0))+
#   xlab("Tissue specificity") +
#   ylab("TF reporter expression")
# 
# sp.plot <- arrangeGrob(x4,x5, nrow = 1, ncol = 2)
# grid.arrange(sp.plot)
# 
# 
# x6 <- ggplot(tf_properties, aes(x = DBD, y = value)) +
#   geom_quasirandom(width = 0.2) +
#   theme_bw() +
#   theme(axis.text.x = element_text(angle = 90, hjust = 1.0, vjust = 0))+
#   xlab("DNA binding domain") +
#   ylab("TF reporter quality")
# 
# x7 <- ggplot(tf_properties, aes(x = DBD, y = log_reporter_activity_relative)) +
#   geom_quasirandom(width = 0.2) +
#   theme_bw() +
#   theme(axis.text.x = element_text(angle = 90, hjust = 1.0, vjust = 0))+
#   xlab("DNA binding domain") +
#   ylab("TF reporter expression")
# 
# sp.plot <- arrangeGrob(x6,x7, nrow = 1, ncol = 2)
# grid.arrange(sp.plot)


```



### Make the same models as before - but now per TF and per condition
```{r log-linear-model-5, out.width= "80%", fig.align= "center", echo=FALSE, warning= FALSE, message = FALSE}
### For each TF separately - only the non-mutated 
cDNA_df_TF <- cDNA_df[cDNA_df$native_enhancer == "No" &
                        cDNA_df$hPGK == "No" & 
                        cDNA_df$neg_ctrls == "No" &
                        cDNA_df$TF != "Random1"&
                        cDNA_df$TF != "Random2"&
                        cDNA_df$TF != "Random3",] %>% 
  mutate(log_reporter_activity = ave(log_reporter_activity, reporter_id, condition, FUN = function(x) mean(x)),
         TF = paste(TF, condition, sep = "_")) %>%
  dplyr::select(log_reporter_activity, TF, promoter, spacing, distance, background) %>%
  unique()

cDNA_df_rel <- cDNA_df_ctrl[cDNA_df_ctrl$neg_ctrls == "No",] %>% 
  mutate(TF = gsub("(^.*)_[0-9]{1,2}bp_[0-9]{1,2}bp_.*", "\\1", reporter_2),
         tf = paste(TF, condition, sep = "_"),
         promoter = gsub(".*(minP|mCMV|hBGm|Random)_.*", "\\1", reporter_2),
         distance = gsub(".*_([0-9]{1,2}bp)_.*", "\\1", reporter_2),
         spacing = gsub(".*_([0-9]{1,2}bp)_[0-9]{1,2}bp_.*", "\\1", reporter_2),
         background = gsub(".*([0-9]{1}$)", "\\1", reporter_2)) %>%
  dplyr::select(log_reporter_activity_relative, tf, promoter, distance, spacing, background) %>% unique() %>%
  mutate(log_reporter_activity_relative = ave(log_reporter_activity_relative, 
                                              tf, promoter, distance, spacing, background, 
                                              FUN = function(x) mean(x)))%>%
  setnames("tf", "TF")

cDNA_df_TF <- merge(cDNA_df_TF, cDNA_df_rel, all = T)



### Calculate expression variance
exp_variance <- data.frame(c("promoter", "spacing", "distance", "background", "Residuals"))
prediction <- data.frame(1:504)
names(prediction) <- "id"
names(exp_variance) <- "feature"
weight <- data.frame(c("(Intercept)", "promotermCMV", "promoterminP", 
                  "spacing5bp", "distance21bp", "background2", "background3"))
names(weight) <- "feature"


cDNA_df_TF$row <- rownames(cDNA_df_TF)

for (i in unique(cDNA_df_TF$TF[-grep("Irx3|Homez|Nr3c1(GR)", cDNA_df_TF$TF)])) {
  cDNA_df_TF_2 <- cDNA_df_TF[-grep("Irx3|Homez|Nr3c1(GR)", cDNA_df_TF$TF),]
  x <- lm(log_reporter_activity ~ promoter + spacing + distance + background, 
                     cDNA_df_TF[cDNA_df_TF$TF == i,])
  y <- data.frame(x$fitted.values) %>% rownames_to_column()
  names(y) <- c("row", i)
  y$id <- 1:nrow(y)
  prediction <- merge(prediction,y, all = T)
  
  w <- data.frame(x$coefficients) %>% rownames_to_column()
  names(w) <- c("feature", i)
  weight <- merge(weight,w, all = T)
  
  x <- anova(x) %>% rownames_to_column("feature") %>%
    setnames("Sum Sq", "sum_sq") %>%
    dplyr::select(feature, sum_sq) %>%
    mutate(sum = sum(sum_sq)) %>%
    mutate(rel_sum_sq = (sum_sq/sum)*100) %>%
    setnames("rel_sum_sq", i)
  x <- x[,c(1,4)]
  
  exp_variance <- merge(exp_variance, x)
}


## Fine-tune dfs
exp_variance <- melt(exp_variance, variable.name = "TF")
exp_variance$group <- "no residual"
exp_variance$group[grep("Residual", exp_variance$feature)] <- "residual"
prediction <- prediction %>% dplyr::select(-id) %>% melt(variable.name = "TF") %>% na.omit() %>%
  dplyr::select(-TF) %>% setnames("value", "activity_predicted")
cDNA_df_TF <- merge(cDNA_df_TF, prediction)
weight <- weight %>% filter(feature != "(Intercept)")
weight <- melt(weight, variable.name = "TF")
categorie <- c("mutated", "condition", "promoter", 
               "spacing", "distance", "background")
weight$features <- gsub(paste(categorie, collapse="|"), "",weight$feature)
ml <- length(weight$features)
feature <- weight$features
weight$condition <- gsub(paste(head(feature, n = ml/2), collapse = "|"), "", weight$feature)
weight <- weight %>% dplyr::select(-feature)



# Model parameters
exp_variance$value[exp_variance$feature == "Residuals"] <- 100 - exp_variance$value[exp_variance$feature == "Residuals"]
exp_variance$feature <- revalue(exp_variance$feature, c("Residuals"="total"))

ggplot(exp_variance[grep("Trp53", exp_variance$TF),], 
       aes(x = reorder(feature, -value), y = value, fill = group)) +
  scale_fill_manual(values = c("#1B998B", "#2D3047")) +
  geom_bar(stat = "identity") + 
  ylab("Variance explained (%)") + xlab("") + 
  labs(title = "Linear variance modelling", 
       subtitle = "Expression variance predicted by features") +
  theme_bw() +
  theme(text = element_text(size = 12)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0, size = 12), 
        axis.text.y = element_text(size = 10)) +
  facet_wrap(~TF)




## Plot weights of parameters in model - this helps to understand the model 
level_order <- c("mCMV", "minP","Random", 
                 "2", "3", "5bp", "21bp")





## Final overview - model fit for each TF
# Feature importance overview for each TF in a single plot (bar-chart including mean activities per feature)


for (i in unique(cDNA_df_TF$TF[grep("Trp53", cDNA_df_TF$TF)])) {
  
  # Plot to show prediction accuracy
  sp <- ggscatter(cDNA_df_TF[cDNA_df_TF$TF == i,], x = "log_reporter_activity", y = "activity_predicted",
   add = "reg.line", 
   add.params = list(color = "blue", fill = "lightgray"),
   conf.int = TRUE, ylab = "Predicted log2-activity", 
   title = i,
   xlab = "Average log2-activity") + theme_bw()
  
  
  # Feature importance overview for each TF (including mean activities per feature)
  fp <- ggplot(exp_variance[exp_variance$TF == i,], 
       aes(x = reorder(feature, -value), y = value, fill = group)) +
  scale_fill_manual(values = c("#1B998B", "#2D3047")) +
  geom_bar(stat = "identity") + 
  ylab("Variance explained (%)") + xlab("") +
  theme_bw() +
  theme(text = element_text(size = 12)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0, size = 12), 
        axis.text.y = element_text(size = 10)) +
    guides(fill = F)
  
  # Plot weights of parameters in model - this helps to understand the model 
  p <- ggplot(weight[weight$TF == i,], 
       aes(x = factor(features, level = level_order), y = value, fill = condition)) +
  scale_fill_manual(values = c("#1B998B", "#2D3047", "#FF9B71", "#ECDD7B", "#C2B2B4", "#A54657")) +
  geom_bar(stat = "identity") + 
  ylab("Weight") + xlab("") +
  theme_bw() +
  theme(text = element_text(size = 12)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 1, size = 10), 
        axis.text.y = element_text(size = 10)) 

  # Combine all plots
  sp.plot <- arrangeGrob(sp,fp,p, nrow = 2, ncol = 2)
  grid.arrange(sp.plot)
}
```

# Session Info
```{r}
paste("Run time: ",format(Sys.time()-StartTime))
getwd()
date()
sessionInfo()
```

