---
title: "Multiplexed identification of TF-specific reporters"
author: 
  - name: "Max Trauernicht"
    email: "m.trauernicht@nki.nl"
    affiliation: "Netherlands Cancer Institute - van Steensel lab"

date: '`r format(Sys.time(), "%d/%m/%Y")`'
output:
  html_document:
    theme: united
    highlight: pygments
    fig_caption: yes
    code_folding: hide
    df_print: kable
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
---

---

### Aim
~36,000 reporters for 86 TFs were transfected into 9 different cell types + treated with ~90 different siRNAs or various signaling molecules. In this script I will analyze TF reporter activities in detail and review how individual reporters respond to TF concentration variations.


### Description of input data
- library = with which library was this sample transfected (1, 2, or 1+2)
- nchar = barcode length (12 for library 1, 13 for library 2)
- commercial_reporter = positive control reporter sequence from commercial source
- neg_ctrls = reporter with mutated TF binding sites?
- hPGK = positive control, chunk from hPGK promoter (only in lib 1)
- native_enhancer = genomic mES enhancer sequence chunks from Miguels libraries (only lib 1)
- activity = rpm + pseudocount cDNA / rpm + pseudocount pDNA
- mean_activity_sample = mean of activity per unique reporter per sample (individually per replicate)
- reporter_activity = mean activity per unique reporter per condition (mean of replicates)
- minP_activity = activity relative to mean of minimal promoter only reporters of that sample
- mean_activity_sample_minP = mean of minP_activity per unique reporter 
- reporter_activity_minP = mean of mean_activity_sample_minP per condition
- nfya_activity = activity relative to mean of nfya reporters (constitutive positive control) of that sample
- mean_activity_sample_nfya = mean of nfya_activity per unique reporter 
- reporter_activity_nfya = mean of mean_activity_sample_nfya per condition
- mean_activity_sample_neg = mean_activity_sample relative to its paired negative control (identical reporter with mutated binding sites)
- reporter_activity_neg = mean of mean_activity_sample_neg per condition
- pval_minP = p-value of t-test comparing activity distributions of individual reporters with the distribution of minimal promoter only reporters within the same condition
- pval_neg = p-value of t-test comparing activity distributions of individual reporters with the distribution of the activities of its paired negative controls within the same condition

---

## Setup {.tabset}

<!-- little HTML script to do indentation of the table of contents -->
<script>
    $(document).ready(function() {
      $items = $('div#TOC li');
      $items.each(function(idx) {
        num_ul = $(this).parentsUntil('#TOC').length;
        $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
      });

    });
</script>

```{css, echo = FALSE}
div.sourceCode {
  overflow-x: hidden;
}
```

### Libraries 

```{r setup, out.width= "80%", fig.align= "center", warning = FALSE, message= FALSE}
knitr::opts_chunk$set(echo = TRUE)
StartTime <-Sys.time()

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8) 
# libraries:
library(RColorBrewer)
library(ggplot2)
library(dplyr)
library(maditr)
library(tibble)
library(pheatmap)
library(ggpubr)
library(visNetwork)
library(ggbeeswarm)
library(ggforce)
library(viridis)
library(plyr)
library(igraph)
library(ggraph)
library(cowplot)
library(gridExtra)
library(pROC)
library(tidyr)
library(stringr)
library(randomForest)
library(ggrastr)
library(readr)
library(ggbiplot)
library(IHW)
library(biomaRt)
library(ggh4x)
library(ggiraph)
library(plotly)
library(AMR)
```

### Functions

```{r out.width= "80%", fig.align= "center"}
### Custom functions
SetFileName <- function(filename, initials) {
  # Set filename with extension and initials to make filename with date integrated.
  filename <- substitute(filename)
  initials <- substitute(initials)
  filename <- paste0(initials, Date, filename)
  filename
}


# Extract p-value from linear model
lmp <- function (modelobject) {
    if (class(modelobject) != "lm") stop("Not an object of class 'lm' ")
    f <- summary(modelobject)$fstatistic
    p <- pf(f[1],f[2],f[3],lower.tail=F)
    attributes(p) <- NULL
    return(p)
}

# Set custom ggplot2 theme and custom colors
theme_classic_lines <- function() {
  theme_pubr(border = F, legend = "top") +
            theme(panel.grid.major = element_line(colour = "#adb5bd", size = 0.25),
                  strip.background = element_rect(fill = "#ced4da")
            )
}

theme_classic_lines_45 <- function() {
  theme_pubr(border = T, legend = "top", x.text.angle = 45) +
            theme(panel.grid.major = element_line(colour = "#adb5bd", size = 0.25),
                  strip.background = element_rect(fill = "#ced4da")
            )
}

theme_classic_lines_90 <- function() {
  theme_pubr(border = T, legend = "top", x.text.angle = 90) +
            theme(panel.grid.major = element_line(colour = "#adb5bd", size = 0.25),
                  strip.background = element_rect(fill = "#ced4da")
            )
}

theme_barplot <- function() {
  theme_pubr(border = T, legend = "none") +
            theme(panel.grid.major.x = element_line(colour = "black", size = 0.25),
                  strip.background = element_blank(),
                  strip.text = element_text(face="bold", hjust=0)
            )
}

theme_set(theme_classic_lines())

colors_diverse <- c("#264653", "#9AC1AE", "#5D987B", "#f2cc8f", "#e76f51")

feature_colors <- c("hBGm" = "#ECD9B1",  "mCMV" = "#F4A261", "minP" = "#CCCCCC", "Random" = "grey93", 
                    "1" = "#000000", "2" = "#7F7F7F", "3" = "#CCCCCC",
                    "21bp" = "#84A98C", "10bps" = "#CAD2C5",
                    "5bp" = "#BFD0D9", "10bp" = "#7C9EB2")

cell_colors <- c("A549" = "#f2cc8f", "HCT116" = "#ED1C24", "HEK293" = "#00B4D8", 
                               "HepG2" = "#B3B3B3", "K562" = "#A67C52", "MCF7" = "#81B29A", 
                               "U2OS" = "#3D405B", "mES" = "#EAB69F", "NPC" = "#E07A5F")

ggplot_custom <- function(...) ggplot2::ggplot(...) + 
  scale_color_manual(values = colors_diverse) + 
  scale_fill_manual(values = colors_diverse)
```

### Loading data

```{r out.width= "80%", fig.align= "center"}
# Import processed bc counts from the preprocessing step
cDNA_df <- read.csv("/DATA/usr/m.trauernicht/projects/SuRE-TF/data/gcf7124_stimulations/results/mt20230901_reporter_activity_filt_combined.csv", header = T)

cDNA_df$stimulation[is.na(cDNA_df$stimulation)] <- "no"

# final measure of reporter activity: combination of reporter_activity_minP and reporter_activity_neg
## for most reporters I will take reporter_activity_neg, however, this cannot be done for commercial controls for example
cDNA_df <- cDNA_df %>%
  mutate(reporter_activity_neg = ifelse(tf %in% c("POU5F1::SOX2", "NR1H4::RXRA", "CEBPB", "STAT1::STAT2", "THRA") | 
                                          commercial_reporter == "Yes" | is.na(reporter_activity_neg), 
                                        reporter_activity_minP, reporter_activity_neg)) %>%
  mutate(tf = gsub("^SOX9", "SOXE", tf)) %>%
  mutate(reporter_id = gsub("^SOX9(.*)", "SOXE\\1", reporter_id)) %>%
  mutate(tf = gsub("^SOX2", "SOX9", tf)) %>%
  mutate(reporter_id = gsub("^SOX2(.*)", "SOX9\\1", reporter_id)) %>%
  mutate(tf = gsub("^SOXE", "SOX2", tf)) %>%
  mutate(reporter_id = gsub("^SOXE(.*)", "SOX2\\1", reporter_id)) %>%
  mutate(tf = gsub("^NFATc1", "NFATC1", tf)) %>%
  mutate(reporter_id = gsub("^NFATc1(.*)", "NFATC1\\1", reporter_id)) %>%
  mutate(tf = gsub("^Myc", "MYC", tf)) %>%
  mutate(reporter_id = gsub("^Myc(.*)", "MYC\\1", reporter_id))

# Load RNA-seq data
tf_rna <- read_tsv("/DATA/usr/m.trauernicht/data/RNA_seq/rna_tpm_all_tfs.tsv")

## Compute fold-changes upon perturbation
## Prepare data frame
cDNA_df2 <- cDNA_df %>%
  #filter(tf != "RFX1") %>%
  dplyr::filter(neg_ctrls == "No", 
                #commercial_reporter == "No", 
                hPGK == "No", 
                str_detect(tf, "RANDOM", negate = T), 
                native_enhancer == "No") %>%
  distinct(tf, condition, reporter_activity_minP, stimulation, reference_condition, cell, reporter_activity_minP, reporter_id, commercial_reporter) %>%
  mutate(reporter_activity_minP = log2(reporter_activity_minP)) %>%
  distinct()

# Export for shiny app
#write_tsv(cDNA_df2, "/DATA/usr/m.trauernicht/projects/SuRE-TF/data/mt20230302_all_activities.tsv")

## Make separate data frame for reference conditions - this will be used for normalizing the other conditions
reference_condition <- cDNA_df2 %>%
  filter(condition %in% reference_condition) %>%
  dplyr::select(-reference_condition, "reference_condition" = condition, -stimulation, "reference_activity" = reporter_activity_minP, reporter_id) %>%
  mutate(reference_activity = ave(reference_activity, reporter_id, reference_condition, FUN = mean)) %>%
  distinct()

## Normalize KD conditions
cDNA_df3 <- cDNA_df2 %>% 
  left_join(reference_condition) %>%
  filter(stimulation == "KD") %>%
  mutate(condition = gsub("HepG2_|mES_", "", condition)) %>%
  mutate(tf = gsub("_.*", "", tf)) %>%
  filter(tf == condition) %>%
  mutate(reporter_dif =  reference_activity - reporter_activity_minP)

## Normalize overexpression conditions
cDNA_df_OE <- cDNA_df2 %>% 
  left_join(reference_condition) %>%
  filter(stimulation == "overexpression", condition != "mES_NR4A2-OE-cDIM12") %>%
  mutate(condition2 = gsub("mES_", "", condition)) %>%
  mutate(condition2 = gsub("-OE$", "", condition2))

cDNA_df_OE$condition2[cDNA_df_OE$condition2 == "NR4A2"] <- "NR4A2::RXRA"
cDNA_df_OE$condition2[cDNA_df_OE$condition2 == "OTX2"] <- "OTX1"


cDNA_df_OE2 <- cDNA_df_OE %>%
  filter(condition2 == "FOSL1") %>%
  mutate(condition2 = "FOS::JUN")

cDNA_df_OE <- rbind(cDNA_df_OE, cDNA_df_OE2)

cDNA_df_OE <- cDNA_df_OE %>%
  mutate(tf = gsub("_.*", "", tf)) %>%
  filter(tf == condition2) %>%
  mutate(reporter_dif = reporter_activity_minP - reference_activity) %>%
  dplyr::select(-condition2)

## Normalize pathway stimulations
stimulations <- data.frame(condition = c("HepG2_CDCA", "HepG2_Rif", "HepG2_T09", "HepG2_Wy", "mES_2i", "mES_BX", "mES_HQ", "mES_IWP2", 
                                         "mES_LIF_CH", "mES_LIF_PD", "mES_LIF_PD", "mES_vitC", "A549_Dex_100", "A549_Nutlin", 
                                         "mES_Bmp4", "mES_Dex", "mES_FK", "mES_LPS", "mES_N2B27_Bmp4", "mES_N2B27_RA", "mES_N2B27_Shh", 
                                         "mES_Nutlin", "mES_RA", "mES_Shh", "mES_vitA", "U2OS_FK", "U2OS_LPS","HCT116_cDIM12", "U2OS_Calcitriol",
                                         "U2OS_PMA", "HCT116_GSK4112", "MCF7_Hexestrol", "MCF7_DIMc", "HCT116_Calcimycin", "HCT116_Calcimycin",
                                         "HCT116_SR8278", "HCT116_SRI011381", "HCT116_ITE", "U2OS_Aldosterone", "U2OS_SRI1078", "U2OS_Cytosporone",
                                         "U2OS_AICAR", "U2OS_Wortmannin", "U2OS_Heat", "A549_5aDHT", "A549_TCPOBOP", "K562_T3", "K562_T3",
                                         "mES_Progesterone", "mES_GSK4716", "mES_FK", "mES_NR4A2-OE-cDIM12"),
                           tf = c("NR1H4::RXRA", "NR1I2", "NR1H2", "PPARA::RXRA", "STAT3", "IRF3", "NFE2L2", "OTX1", "EGR1", 
                                  "TCF7", "TCF7L2", "NFE2L2", "NR3C1", "TP53", "SMAD4", "NR3C1", "CREB1", "IRF3", "SMAD4",
                                  "RARA", "GLI1", "TP53", "RARA", "GLI1", "RARA", "CREB1", "IRF3", "NR4A2::RXRA", "VDR::RXRA", 
                                  "NFKB1", "NR1D1", "ESR1", "NR4A1", "NFAT5", "NFATC1", "NR1D1", "SMAD2::SMAD3::SMAD4", 
                                  "AHR::ARNT", "NR3C2", "RORA", "NR4A1", "TEAD1", "FOXO1", "HSF1", "AR", "NR1I3", "THRA", 
                                  "THRB", "PGR", "ESRRB", "CREB1", "NR4A2::RXRA"),
                           effect_size = c(1,1,1,1,0,0,1,1,1,0,0,0,1,1,1,1,1,1,1,1,1,1,1,1,0,1,1,1,1,1,0,1,1,1,1,1,1,1,1,1,1,0,1,1,1,1,1,1,1,1,1,1))

stimulations_effective <- data.frame(condition = c("HepG2_CDCA", "HepG2_Rif", "mES_2i", "mES_BX", "mES_HQ", "mES_IWP2", 
                                                   "mES_LIF_CH", "mES_LIF_PD", "mES_LIF_PD", "A549_Nutlin", 
                                                   "mES_N2B27","U2OS_Calcitriol", "U2OS_PMA", "MCF7_Hexestrol", 
                                                   "HCT116_Calcimycin", "HCT116_Calcimycin", "HCT116_ITE",
                                                   "U2OS_AICAR", "U2OS_Wortmannin", "U2OS_Heat", "K562_T3", "K562_T3",
                                                   "mES_FK", "mES_NR4A2-OE-cDIM12", "A549_TCPOBOP", "A549_5aDHT", "U2OS_Cytosporone",
                                                   "U2OS_SRI1078", "U2OS_Aldosterone", "HCT116_GSK4112", "mES_N2B27", "A549_Dex_100",
                                                   "HepG2_T09", "U2OS_PMA", "mES_LIF_CH", "U2OS_Calcitriol", "mES_N2B27",
                                                   "mES_serum_2i_LIF", "mES_serum_2i_LIF", "mES_2i", "U2OS_Wortmannin", "mES_LIF_PD", "MCF7_DIMc"),
                           tf = c("NR1H4::RXRA", "NR1I2", "STAT3", "IRF3", "NFE2L2", "OTX1", "EGR1", "TCF7", "TCF7L2", "TP53", 
                                  "RARA", "VDR::RXRA", "NFKB1", "ESR1","NFAT5", "NFATC1", "AHR::ARNT", "TEAD1", "FOXO1", "HSF1", 
                                  "THRA", "THRB", "CREB1", "NR4A2::RXRA", "NR1I3", "AR", "NR4A1", "RORA", "NR3C2", "NR1D1", 
                                  "RARA:RXRA", "NR3C1", "NR1H2", "REL", "ELK1", "E2F1", "RXRA", "SMAD4",
                                  "SMAD2::SMAD3::SMAD4", "NR5A2", "WT1", "TEAD1", "NR4A1"),
                           effect_size = c(1,1,0,0,1,1,1,0,0,1,1,1,1,1,1,1,1,0,1,1,1,1,1,0,1,1,1,1,1,1,1,1,1,1,0,1,1,1,1,0,0,1,1))

stimulations_interesting <- c("HepG2_Wy", "mES_vitC")

cDNA_df4 <- cDNA_df2 %>%
  left_join(reference_condition) %>%
  mutate(tf = gsub("_.*", "", tf)) %>%
  filter(stimulation == "pathway" | condition == "mES_NR4A2-OE-cDIM12") %>%
  mutate(reporter_dif =  reference_activity - reporter_activity_minP)

## Check out which TFs respond to which stimulus
ggplot(cDNA_df4 %>% 
         filter(tf == "STAT1::STAT2") %>% 
         mutate(dif_mean = ave(reporter_dif, condition, FUN = function(x) mean(x, na.rm = T))),
       aes(x = reorder(condition, dif_mean), y = -reporter_dif)) +
  geom_quasirandom() +
  theme_pubr(x.text.angle = 90)

cDNA_df4 <- merge(cDNA_df4, stimulations_effective) ## Only keep relevant conditions

cDNA_df4 <- cDNA_df4 %>%
  mutate(reporter_dif = ifelse(effect_size == 1, -reporter_dif, reporter_dif)) %>% ## Change sign of difference, based on whether the stimulation should cause an increase or decrease (effect_size)
  dplyr::select(-effect_size)

## Normalize degron conditions
cDNA_df_degron <- cDNA_df2 %>% 
  mutate(tf = gsub("_.*", "", tf)) %>%
  left_join(reference_condition) %>%
  filter(stimulation == "degron") %>%
  filter(condition != "mES_Nanog_AID") %>% ## Doesn't have a direct target (I don't have Nanog reporters)
  mutate(target_tf = toupper(gsub("mES_(.*)_AID", "\\1", condition)))

cDNA_df_degron2 <- cDNA_df_degron
cDNA_df_degron2$target_tf[cDNA_df_degron2$condition == "mES_Sox2_AID"] <- "SOX2"
cDNA_df_degron3 <- cDNA_df_degron
cDNA_df_degron3$target_tf[cDNA_df_degron3$condition == "mES_Pou5f1_AID"] <- "POU5F1::SOX2"
cDNA_df_degron4 <- cDNA_df_degron
cDNA_df_degron4$target_tf[cDNA_df_degron4$condition == "mES_Pou5f1_AID"] <- "POU5F1"
cDNA_df_degron5 <- cDNA_df_degron
cDNA_df_degron5$target_tf[cDNA_df_degron5$condition == "mES_Sox2_AID"] <- "POU5F1::SOX2"
cDNA_df_degron <- rbind(cDNA_df_degron, cDNA_df_degron2, cDNA_df_degron3, cDNA_df_degron4, cDNA_df_degron5)
cDNA_df_degron <- cDNA_df_degron %>% 
  filter(tf == target_tf) %>%
  mutate(reporter_dif =  reference_activity - reporter_activity_minP) %>%
  dplyr::select(-target_tf)

## Combine stimulations and KDs 
perturbation_df <- rbind(cDNA_df3, cDNA_df_OE, cDNA_df4, cDNA_df_degron) %>% ## Combine all perturbation data
  mutate(condition_2 = paste(condition, reference_condition)) %>%
  mutate(mean_reporter_dif = ave(reporter_dif, condition_2, tf, FUN = function(x) mean(x, na.rm = T))) %>%
  group_by(tf) %>%
  arrange(desc(mean_reporter_dif)) %>% 
  top_n(1, wt = mean_reporter_dif) %>% ## Some TFs have multiple perturbations: for those I only keep the strongest perturbations (on average across all reporters)
  ungroup() %>%
  mutate(reporter_dif = ifelse(tf == "CEBPB", -reporter_dif, reporter_dif)) %>%
  dplyr::select(reporter_id, tf, commercial_reporter, "perturbation_condition" = condition, reporter_dif) %>%
  distinct()

## Define optimal candidate conditions for each TF: either the highest expressing cell line, or the stimulated condition
### All stimulation conditions
ref_conditions1 <- cDNA_df4 %>%
  right_join(stimulations_effective) %>%
  filter(effect_size == 1) %>%
  distinct(tf, condition, cell, reporter_id, commercial_reporter, reporter_activity_minP) %>%
  group_by(tf, condition) %>%
  dplyr::arrange(desc(reporter_activity_minP)) %>%
  top_n(10) %>%
  ungroup() %>%
  mutate(median_reporter_activity_minP = ave(reporter_activity_minP, tf, condition, FUN = median)) %>%
  group_by(tf) %>%
  arrange(desc(median_reporter_activity_minP)) %>%
  top_n(1) %>%
  ungroup() %>%
  distinct(tf, condition)

ref_conditions_oe <- cDNA_df_OE %>%
  distinct(tf, condition, cell, reporter_id, commercial_reporter, reporter_activity_minP) %>%
  group_by(tf, condition) %>%
  dplyr::arrange(desc(reporter_activity_minP)) %>%
  top_n(10) %>%
  ungroup() %>%
  mutate(median_reporter_activity_minP = ave(reporter_activity_minP, tf, condition, FUN = median)) %>%
  group_by(tf) %>%
  arrange(desc(median_reporter_activity_minP)) %>%
  top_n(1) %>%
  ungroup() %>%
  distinct(tf, condition)

### Top 2 conditions per TF based on highest TPM and highest median reporter activity
cDNA_df2 <- cDNA_df %>%
  #filter(tf != "RFX1") %>%
  dplyr::filter(neg_ctrls == "No", 
                #commercial_reporter == "No", 
                hPGK == "No", 
                str_detect(tf, "RANDOM", negate = T), 
                native_enhancer == "No") %>%
  distinct(tf, condition, reporter_activity_minP, stimulation, reference_condition, cell, reporter_activity_minP, reporter_id, commercial_reporter) %>%
  mutate(reporter_activity_minP = log2(reporter_activity_minP)) %>%
  distinct()

ref_conditions2 <- merge(cDNA_df2 %>% filter(stimulation == "no", condition != "HCT116_SR8278") %>%
                           mutate(tf = gsub("_.*", "", tf)) %>% 
                           mutate(tf = toupper(tf)), 
                         tf_rna %>%
                           mutate(tf = toupper(tf)), by = c("cell", "tf")) %>%
  mutate(mean_reporter_activity_minP = ave(reporter_activity_minP, tf, condition, FUN = median)) %>%
  distinct(mean_reporter_activity_minP, tf, condition, TPM_norm) %>%
  filter(!condition %in% c("mES_Nanog_ctrl", "mES_Pou5f1_ctrl", "mES_Sox2_AID_UMI", "mES_Sox2_ctrl", "mES_NT", "HepG2_NT")) %>%
  group_by(tf) %>%
  dplyr::arrange(desc(TPM_norm)) %>%
  top_n(n = 3, wt = TPM_norm) %>%
  ungroup() %>%
  group_by(tf) %>%
  dplyr::arrange(desc(mean_reporter_activity_minP)) %>%
  top_n(n = 2, wt = mean_reporter_activity_minP) %>%
  ungroup() %>%
  distinct(tf, condition)

## Now I combine both to have 2-3 top conditions per TF
ref_conditions <- rbind(ref_conditions1, ref_conditions2, ref_conditions_oe) %>%
  rbind(data.frame(tf = c("PAX6", "ESRRB", "SMAD4", "GATA1", "RUNX2", "NR5A2", "SMAD2::SMAD3::SMAD4", "RFX1", 
                          "FOXO1", "FOSL1", "ETS2", "ONECUT1"), 
                   condition = c("HepG2_NT", "mES_serum_2i_LIF", "HepG2_NT", "K562_Hemin", "U2OS", "mES_2i_LIF", "HepG2_NT", "mES_2i_LIF",
                                 "mES_FOXA1-OE", "mES_2i_LIF", "HepG2_NT", "HepG2_NT")))

## Include reporter specificities based on KD of related TFs: I want to punish reporters that reduce activity upon KD of related TFs
related_tfs <- data.frame(tf = c("CREB1", "FOS::JUN", "CEBPB", "XBP1", "CEBPB", "FOS::JUN", "CREB1", "ELK1", "ETS2", "ELK1", "ETS2", 
                                 "ELK1", "FOXA1", "FOXA1", "GATA1", "GATA4", "GATA1", "HNF4A", "IRF3", "TP53", "IRX3",
                                 "KLF4", "SP1", "MAF::NFE2", "NFE2L2", "NFKB1", "NFYA", "NR1H2", "PPARA::RXRA", "NFKB1", 
                                 "REL", "IRF3", "STAT3", "TCF7", "TCF7L2", "TFCP2L1", "TFCP2L1", "GRHL1", "THRB", "THRA", "TP53", 
                                 "TFCP2L1", "E2F1", "E2F1", "SP1", "EGR1", "EGR1", "KLF4", "KLF4", "SP1", "SP1", "NFIA", "HSF1", 
                                 "KLF4", "MEF2A", "PPARG::RXRA", "ESRRB", "POU2F1", "POU5F1", "POU5F1::SOX2", "PPARG::RXRA",
                                 "PPARA::RXRA", "NR1H2", "NR1H2", "SOX2", "SOX9", "SOX2", "SOX9", "TCF7", "TCF7L2", "TEAD1", 
                                 "STAT1::STAT2", "E2F1", "FOSL1", "IRF1", "STAT1::STAT2", "IRF3", "IRF1", "GATA3",
                                 "HIF1A", "TFEB", "STAT6", "ATF4", "ATF6", "CEBPA", "TFEB", "STAT4", "ATF4"),
                          related_tf = c("ATF2", "ATF2", "ATF4", "ATF6", "CEBPG", "CREB1", "FOS::JUN", "ETS1", "ELK1", "ETS2", 
                                         "ETS1", "ETV4", "FOXA2", "FOXP1", "GATA2", "GATA2", "GATA4", "HNF4G", "IRF2", "IRX3", "TP53",
                                         "SP1", "KLF4", "NFE2L2", "MAF::NFE2", "NFKB2", "NFYB", "PPARA", "NR1H2", "REL", "NFKB1", 
                                         "STAT1::STAT2", "STAT1::STAT2", "TCF7L2", "TCF7", "TFCP2", "GRHL1", "TFCP2L1", "THRA", "THRB", 
                                         "TP73", "UBP1", "E2F2", "E2F3", "EGR1", "SP1", "KLF4", "SP1", "EGR1", "KLF4", "EGR1", "HIC2", 
                                         "HSF2", "KLF9", "MEF2B", "NR1H2", "NR5A2", "POU5F1", "POU2F1", "POU2F1", "PPARA::RXRA",
                                         "PPARG::RXRA", "PPARA::RXRA", "PPARG::RXRA", "SOX13", "SOX13", "SOX9", "SOX2", "TCF7L1", 
                                         "TCF7L1", "TEAD2", "STAT3", "SP1", "ATF2", "IRF3", "IRF3", "STAT1::STAT2", "STAT1::STAT2",
                                         "GATA4", "XBP1", "XBP1", "STAT3", "CEBPG", "XBP1", "CEBPG", "ATF4", "STAT3", "ATF6"))

## Off-targets based on KD of related TFs
off_target_activities <- cDNA_df2 %>% 
  left_join(reference_condition) %>%
  filter(stimulation == "KD") %>%
  mutate(condition = gsub("HepG2_|mES_", "", condition)) %>%
  mutate(tf = gsub("_.*", "", tf)) %>%
  mutate(reporter_dif =  reference_activity - reporter_activity_minP) %>%
  left_join(related_tfs) %>%
  na.omit() %>%
  filter(condition == related_tf) %>%
  mutate(reporter_dif_mean = ave(reporter_dif, tf, cell, related_tf, FUN = mean)) %>% 
  ## Some TFs have multiple related TFs for which I have KD data: keep strongest effect only
  group_by(tf) %>%
  arrange(desc(reporter_dif_mean)) %>% 
  top_n(1, wt = reporter_dif_mean) %>%
  ungroup() %>%
  dplyr::select(reporter_id, tf, commercial_reporter, "off_target_activity" = reporter_dif, "off_target" = related_tf) %>%
  distinct() %>%
  mutate(off_target = gsub("(.*)", "\\1_KD", off_target))

# ## Also: punish reporters that are active in cells where the TF is not expressed
# off_target_activities2 <- cDNA_df2 %>%
#   mutate(tf = gsub("_.*", "", tf)) %>%
#   left_join(tf_rna) %>%
#   filter(TPM_norm == 0, stimulation == "no") %>%
#   mutate(off_target_activity = ave(reporter_activity_minP, reporter_id, FUN = mean)) %>%
#   distinct(reporter_id, off_target_activity, tf, commercial_reporter) %>%
#   mutate(off_target = "Expression_0TPM")

## Off-targets based on degrons
related_tfs2 <- data.frame(condition = c("mES_Sox2_AID", "mES_Pou5f1_AID"),
                          related_tf = c("SOX9", "POU2F1"))

off_target_activities3 <- cDNA_df2 %>% 
  left_join(reference_condition) %>%
  filter(stimulation == "degron") %>%
  mutate(tf = gsub("_.*", "", tf)) %>%
  mutate(reporter_dif =  reference_activity - reporter_activity_minP) %>%
  left_join(related_tfs2) %>%
  na.omit() %>%
  filter(tf == related_tf) %>%
  dplyr::select(reporter_id, tf, commercial_reporter, "off_target_activity" = reporter_dif, "off_target" = related_tf) %>%
  mutate(off_target = ifelse(off_target == "SOX9", "SOX2_AID", "POU5F1_AID")) %>%
  distinct()

## Off-targets based on response to wrong stimulus
off_target_activities_stimulus <- data.frame(off_target_tf = c("IRX3", "NR1I2", "WT1", "NR4A2::RXRA", "FOS::JUN", "NR1I3", "VDR::RXRA", "NR1I2",
                                                               "ESR1", "NR5A2", "NR4A1", "NR3C2", "NR3C1", "NR3C1", "NR1H4::RXRA", "NR1H2", "RBPJ",
                                                               "ATF2", "ATF6", "CLOCK", "Myc"),
                                             condition = c("A549_Nutlin", "U2OS_Calcitriol", "mES_LIF_CH", "HepG2_CDCA", "mES_FK", "U2OS_Calcitriol",
                                                           "A549_TCPOBOP", "A549_TCPOBOP", "MCF7_DIMc", "MCF7_DIMc", "NR4A2-OE-cDIM12", "A549_5aDHT",
                                                           "U2OS_Aldosterone", "A549_5aDHT", "HepG2_T09", "HepG2_CDCA", "HCT116_Calcimycin",
                                                           "mES_FOSL1-OE", "mES_FOSL1-OE", "mES_FK", "mES_FK"))

off_target_activities4 <- cDNA_df2 %>% 
  left_join(reference_condition) %>%
  filter(stimulation == "pathway") %>%
  mutate(tf = gsub("_.*", "", tf)) %>%
  mutate(reporter_dif = reporter_activity_minP - reference_activity) %>%
  left_join(off_target_activities_stimulus) %>%
  na.omit() %>%
  filter(tf == off_target_tf) %>%
  dplyr::select(reporter_id, tf, commercial_reporter, "off_target_activity" = reporter_dif, "off_target" = condition) %>%
  distinct()

## Off-targets based on response to overexpression
off_target_activities_stimulus <- data.frame(off_target_tf = c("FOXO1"),
                                             condition = c("mES_FOXA1-OE"))

off_target_activities5 <- cDNA_df2 %>% 
  left_join(reference_condition) %>%
  filter(stimulation == "overexpression") %>%
  mutate(tf = gsub("_.*", "", tf)) %>%
  mutate(reporter_dif = reporter_activity_minP - reference_activity) %>%
  left_join(off_target_activities_stimulus) %>%
  na.omit() %>%
  filter(tf == off_target_tf) %>%
  dplyr::select(reporter_id, tf, commercial_reporter, "off_target_activity" = reporter_dif, "off_target" = condition) %>%
  distinct()

off_target_activities <- rbind(off_target_activities,
                               #off_target_activities2,
                               off_target_activities3,
                               off_target_activities4,
                               off_target_activities5)

off_target_activities <- off_target_activities %>%
  dplyr::select(reporter_id, off_target_activity, tf, off_target) %>%
  mutate(mean_off_target_activity = ave(off_target_activity, tf, off_target, FUN = mean)) %>%
  group_by(tf) %>%
  arrange(desc(mean_off_target_activity)) %>% 
  top_n(1, wt = mean_off_target_activity) %>% ## Some TFs have multiple perturbations: for those I only keep the strongest off-target effect (on average across all reporters)
  ungroup() %>%
  group_by(reporter_id) %>%
  arrange(desc(off_target_activity)) %>% 
  top_n(1, wt = off_target_activity) %>%
  distinct(reporter_id, off_target_activity, off_target) %>%
  mutate(off_target_activity = ifelse(off_target_activity < 0, 0, off_target_activity))
```

---

## Investigate individual reporter activities - 1

Aim: Show that there are differences in activities between reporter variants

```{r}
## Figure 1E: Activities per cell type - density
cDNA_df2 <- cDNA_df %>%
  filter(stimulation == "no",
         commercial_reporter == "No", 
         hPGK == "No", 
         str_detect(tf, "RANDOM", negate = T), 
         native_enhancer == "No") %>%
  distinct(tf, reporter_activity_minP, cell, reporter_id, neg_ctrls) %>%
  mutate(reporter_activity_minP = ave(reporter_activity_minP, cell, reporter_id, FUN = mean)) %>%
  mutate(reporter_activity_minP = log2(reporter_activity_minP)) %>%
  distinct() 

ggplot(cDNA_df2 %>%
         mutate(cell = factor(cell, levels = c("A549", "K562", "HCT116", "MCF7", "HepG2", 
                                               "U2OS", "HEK293", "mES", "NPC"))),
       aes(x = reporter_activity_minP, fill = neg_ctrls)) +
  geom_density(alpha = .4) +
  facet_wrap(~cell, ncol = 2) +
  scale_fill_manual(values = c("Yes" = "grey70", "No" = "#DD6B48")) +
  theme_pubr(border = T)

## Explore PCA
reporter_mat <- cDNA_df %>%
  filter(stimulation == "no",
         neg_ctrls == "No",
         commercial_reporter == "No", 
         hPGK == "No", 
         str_detect(tf, "RANDOM", negate = T), 
         native_enhancer == "No") %>%
  distinct(tf, reporter_activity_minP, cell, neg_ctrls) %>%
  mutate(reporter_activity_minP = ave(reporter_activity_minP, cell, tf, FUN = mean)) %>%
  mutate(reporter_activity_minP = log2(reporter_activity_minP)) %>%
  distinct(reporter_activity_minP, tf, cell) %>%
  spread(key = "cell", value = "reporter_activity_minP") %>%
  column_to_rownames("tf") %>%
  replace(is.na(.), 0) %>%
  as.matrix()

meta_pca <- data.frame(row.names = c("A549", "HCT116", "HEK293", "HepG2", "K562", "MCF7", "mES", "NPC", "U2OS"),
                       cell = c("A549", "HCT116", "HEK293", "HepG2", "K562", "MCF7", "mES", "NPC", "U2OS"),
                       organism = c("human", "human", "human", "human", "human", "human", "mouse", "mouse", "human"))

reporter_pca2 <- prcomp(t(reporter_mat))

ggplot_pca(reporter_pca2, arrows = F, groups = meta_pca$cell) +
  theme_pubr(border = T)

pcaloadings <- reporter_pca2$rotation


plot(pcaloadings[,1:2],   # x and y data
     pch=21,              # point shape
     bg="black",          # point color
     cex=1,               # point size
     main="Loadings"      # title of plot
)

text(pcaloadings[,1:2],             # sets position of labels
     labels=rownames(pcaloadings)   # print labels
)

## UMAP instead of PCA
set.seed(3) #Get consistent umap
tf_umap <- umap(reporter_mat) #umap algorithm {umap package}
colnames(tf_umap$layout) <- c("A","B") #Change colnames

### Extract clustering data
hc <- hclust(dist(reporter_mat), method = "complete")
cluster_tfs <- data.frame("cluster" = cutree(hc, k = 5)) %>%
  rownames_to_column("tf")
tf_umap_plot <- tf_umap$layout %>% as.data.frame() %>% rownames_to_column(var = "tf") %>% left_join(cluster_tfs) #Import data from previous line

ggplot() +
     #geom_polygon(data = all_cliques, aes(x = A, y = B, group = clique, fill = type), alpha = 0.4) +
     stat_density2d(data = tf_umap_plot, geom = "polygon", aes(A,B, fill = cluster, alpha = ..level..), bins = 8) +
     theme_pubr() + geom_point(data = tf_umap_plot, aes(A,B), size = 4) + 
     theme(panel.grid = element_blank()) + scale_alpha(range = c(0.15,0.3)) +
     geom_text_repel(data = tf_umap_plot, aes(A,B,label = tf), force = 40) + 
     ylab("UMAP2") + xlab("UMAP1")

## Figure S1A: Compare TF reporters to native enhancer controls
native_activities <- cDNA_df %>%
  filter(condition == "mES_2i_LIF", neg_ctrls == "No") %>%
  distinct(reporter_id, tf, reporter_activity_minP, condition) %>%
  mutate(tf = gsub("_.*", "", tf)) %>%
  mutate(native_enhancer = ifelse(tf %in% c("e97", "e6", "e19", "e11", "e93"), tf, "tf_reporter")) %>%
  mutate(native_enhancer2 = ifelse(native_enhancer == "tf_reporter", native_enhancer, "native_enhancer"))
  
ggplot(native_activities,
       aes(x = native_enhancer2, y = log2(reporter_activity_minP))) +
  geom_quasirandom_rast(raster.dpi = 600, alpha = .2) +
  geom_violin(fill = NA) +
  stat_summary(fun = median, fun.min = median, fun.max = median, geom = "crossbar", color = "red", width = 0.25, lwd = 0.4) +
  theme_pubr()
```


---


## Correlate with RNA-seq data

Aim: Does the reporter activity correlate with transcript abundance across conditions?
```{r, warning = F}
## Prepare tf reporter data frame
cDNA_df2 <- cDNA_df %>%
  #filter(tf != "RFX1") %>%
  dplyr::filter(neg_ctrls == "No",
                #commercial_reporter == "No",
                stimulation == "no",
                hPGK == "No",
                str_detect(tf, "RANDOM", negate = T),
                native_enhancer == "No") %>%
  distinct(tf, reporter_activity_minP, cell, reporter_id, commercial_reporter) %>%
  mutate(reporter_activity_minP = ave(reporter_activity_minP, cell, reporter_id, FUN = mean)) %>%
  mutate(reporter_activity_minP = log2(reporter_activity_minP)) %>%
  distinct() %>%
  mutate(tf = gsub("_.*", "", tf))

## Combine TF reporter data with RNA-seq data
cDNA_df3 <- merge(cDNA_df2, tf_rna, all = T) %>%
  mutate(TPM = ifelse(is.na(TPM), 0, TPM)) %>%
  mutate(TPM_norm = ifelse(is.na(TPM_norm), 0, TPM_norm)) %>%
  na.omit() %>%
  mutate(log_TPM_norm = log2(TPM_norm + 0.01)) %>%
  mutate(range_TPM = quantile(log_TPM_norm, probs = 0.95) - quantile(log_TPM_norm, probs = 0.05)) %>%
  mutate(range_activity = quantile(reporter_activity_minP, probs = 0.95) - quantile(reporter_activity_minP, probs = 0.05)) %>%
  mutate(dif_range = range_TPM / range_activity) %>%
  mutate(log_TPM_norm = log_TPM_norm / dif_range) %>%
  dplyr::select(-range_TPM, -range_activity, -dif_range)

## Compute pearson correlations for each individual reporter of reporter activity vs. TPM_norm counts across the nine cell types
for (i in unique(cDNA_df3$reporter_id)) {
  if (nrow(cDNA_df3[cDNA_df3$reporter_id == i,]) > 4) {
    
    #print(i)
    
    cDNA_df3$cor_pval[cDNA_df3$reporter_id == i] <- cor.test(2^cDNA_df3$reporter_activity_minP[cDNA_df3$reporter_id == i],
                                                                    cDNA_df3$TPM_norm[cDNA_df3$reporter_id == i], method = "pearson", exact = F)$p.value
    
    cDNA_df3$cor[cDNA_df3$reporter_id == i] <- cor(2^cDNA_df3$reporter_activity_minP[cDNA_df3$reporter_id == i], 
                                                 cDNA_df3$TPM_norm[cDNA_df3$reporter_id == i], method = "pearson")
  }
  else {
    cDNA_df3$cor_pval[cDNA_df3$reporter_id == i] <- 1
    cDNA_df3$cor[cDNA_df3$reporter_id == i] <- 0
  }
}

## I am only interested in positive correlations, so set all negative correlations to zero
cDNA_df3 <- cDNA_df3 %>%
  mutate(cor_pval = ifelse(cor_pval == cor, 1, cor_pval)) %>%
  group_by(tf) %>%
  mutate(cor_pval_fdr = p.adjust(cor_pval, method = "fdr")) %>%
  ungroup() %>%
  mutate(cor_pval = ifelse(cor < 0, 1, cor_pval)) %>%
  mutate(cor_pval_fdr = ifelse(cor < 0, 1, cor_pval_fdr)) %>%
  mutate(cor_pval = -log10(cor_pval)) %>%
  mutate(cor_pval_fdr = -log10(cor_pval_fdr)) %>%
  mutate(cor_pval = ifelse(is.na(cor_pval), 1, cor_pval)) %>%
  mutate(cor_pval_fdr = ifelse(is.na(cor_pval_fdr), 1, cor_pval_fdr)) %>%
  mutate(cor = ifelse(is.na(cor), 1, cor))

#
# ## Plot the correlations including the p-values
# for (i in unique(cDNA_df3$tf)) {
#
#   x <- cDNA_df3 %>%filter(tf == i)
#
#   label <- data.frame(reporter_id = x$reporter_id, label = round(x$cor_pval, 2))
#
#   p <- ggplot(x,
#               aes(x = TPM_norm, y = reporter_activity_minP)) +
#     geom_point() +
#     geom_abline(lty = 2) +
#     geom_smooth(method = "lm", se = F) +
#     facet_wrap(~reporter_id) +
#     ggtitle(i) +
#     ylab("Reporter Activity") +
#     xlab("TPM_norm") +
#     geom_label(data = label, aes(label = label, color = ifelse(label > 1, "red", "black")),
#                x = Inf, y = -Inf, hjust=1, vjust=0,
#                inherit.aes = FALSE) +
#     scale_color_manual(values = c("black", "red"))
#
#   print(p)
# }

## Filter for TFs where correlations actually make sense -> minimum TPM > 0.5, maximum TPM < 0.5, dif_TPM > 5
rna_correlations_df <- cDNA_df3 %>%
  mutate(min_TPM = ave(TPM_norm, tf, FUN = min),
         max_TPM = ave(TPM_norm, tf, FUN = max)) %>%
  mutate(dif_TPM = max_TPM - min_TPM) %>%
  mutate(mean_tf_activity = ave(reporter_activity_minP, tf, cell, FUN = mean)) %>%
  mutate(max_mean_tf_activity = ave(mean_tf_activity, tf, FUN = max)) %>%
  mutate(reasonable = ifelse(max_TPM > .5 & 
                               dif_TPM > 5 & 
                               min_TPM < 6 & 
                               max_mean_tf_activity > 0.9 | 
                               tf %in% c("TEAD1", "GATA1", "NFATC1", "GLI1", "STAT3", "SP1", "FOXA1", "NFYA"), "Yes", "No"))

reasonable_corr <- rna_correlations_df %>%
  filter(reasonable == "No")

## Export correlations
rna_correlations <- cDNA_df3 %>%
  distinct(reporter_id, cor_pval, cor) %>%
  mutate(cor_pval = ifelse(reporter_id %in% reasonable_corr$reporter_id, 0, cor_pval)) %>%
  mutate(cor = ifelse(reporter_id %in% reasonable_corr$reporter_id, 0, cor))


## Plot correlation p-values per TF
ggplot(rna_correlations_df %>%
                filter(commercial_reporter == "No", reasonable == "Yes") %>%
                distinct(reporter_id, cor_pval, tf) %>%
           mutate(sign = ifelse(cor_pval > 1, "Yes", "No")) %>%
             mutate(mean_cor_pval = ave(cor_pval, tf, FUN = mean)),
       aes(x = reorder(tf, -mean_cor_pval), y = cor_pval, color = sign)) +
  geom_quasirandom() +
  geom_hline(yintercept = 1, lty = 2) +
  theme_pubr(border = T, x.text.angle = 90) +
  scale_color_manual(values = c("Yes" = colors_diverse[2], "No" = colors_diverse[1]))

## Show how reporter activity correlates with RNA-seq correlation for a couple of TFs
ggplot(rna_correlations_df %>%
         filter(commercial_reporter == "No", tf == "GRHL1", cell == "K562") %>%
         distinct(reporter_id, cor_pval, reporter_activity_minP) %>%
         mutate(spacer_length = gsub("GRHL1_(.*bp)_.*bp.*", "\\1", reporter_id)) %>%
         mutate(background = gsub(".*_([1-3]$)", "\\1", reporter_id)),
       aes(x = reporter_activity_minP, y = cor_pval, color = spacer_length)) +
  geom_abline(lty = 2) +
  geom_point(size = 2) +
  theme_pubr() +
  theme(legend.position = "none") +
  scale_color_manual(values = c("10bp" = "#7c9eb2", "5bp" = "grey80"))

ggplot(rna_correlations_df %>%
         filter(commercial_reporter == "No", tf == "GATA4", cell == "HepG2") %>%
         distinct(reporter_id, cor_pval, reporter_activity_minP) %>%
         mutate(spacer_length = gsub("GATA4_(.*bp)_.*bp.*", "\\1", reporter_id)) %>%
         mutate(background = gsub(".*_([1-3]$)", "\\1", reporter_id)),
       aes(x = reporter_activity_minP, y = cor_pval, color = spacer_length)) +
  geom_abline(lty = 2) +
  geom_point(size = 2) +
  theme_pubr() +
  theme(legend.position = "none") +
  scale_color_manual(values = c("10bp" = "#7c9eb2", "5bp" = "grey80"))

ggplot(rna_correlations_df %>%
         filter(commercial_reporter == "No", tf == "HNF4A", cell == "HepG2") %>%
         distinct(reporter_id, cor_pval, reporter_activity_minP) %>%
         mutate(spacer_length = gsub("HNF4A_(.*bp)_.*bp.*", "\\1", reporter_id)) %>%
         mutate(background = gsub(".*_([1-3]$)", "\\1", reporter_id)),
       aes(x = reporter_activity_minP, y = cor_pval, color = spacer_length)) +
  geom_abline(lty = 2) +
  geom_point(size = 2) +
  theme_pubr() +
  theme(legend.position = "none") +
  scale_color_manual(values = c("10bp" = "#7c9eb2", "5bp" = "grey80"))

## Examplify for a couple of TFs
ggplot_custom(cDNA_df3 %>%
                filter(commercial_reporter == "No") %>%
                distinct(reporter_id, cor_pval, tf, commercial_reporter) %>%
                mutate(mean_cor_pval = ave(cor_pval, tf, FUN = mean)) %>%
                mutate(sign = ifelse(cor_pval < 1.30103, "No", "Yes")) %>%
                filter(tf %in% c("HNF4A", "GATA4", "ESR1", "THRB", "GATA1")),
       aes(x = reorder(tf, -mean_cor_pval), y = cor_pval, color = sign)) +
  geom_quasirandom() +
  geom_hline(yintercept = 1.30103, lty = 2) +
  theme_pubr(border = T)

ggplot(cDNA_df3 %>%
           filter(reporter_id %in% c("HNF1A_10bp_10bp_minP_1", "HNF1A_5bp_21bp_mCMV_2")) %>%
           distinct(cell, reporter_id, reporter_activity_minP, TPM_norm),
       aes(x = TPM_norm, y = 2^reporter_activity_minP)) +
    geom_abline(lty = 2) +
    geom_smooth(method = "lm", color = "black") +
    geom_point(size = 2) +
    facet_wrap(~reporter_id, ncol = 1) +
    theme_pubr(border = T)

ggplot(cDNA_df3 %>%
           filter(tf == "SP1", commercial_reporter == "No")%>%
           mutate(promoter = gsub(".*(Random|minP|hBGm|mCMV).*", "\\1", reporter_id)) %>%
           distinct(cell, reporter_id, reporter_activity_minP, TPM_norm, promoter, cor_pval),
       aes(x = TPM_norm, y = 2^reporter_activity_minP)) +
  geom_point(alpha = .2, aes(color = cor_pval)) +
  geom_smooth(method = "lm", se = F, aes(group = reporter_id, color = cor_pval)) +
  theme_pubr(border = T) +
  facet_wrap(~promoter) +
  scale_color_gradient(low = "#264653", high = "#9AC1AE")

ggplot(cDNA_df3 %>%
           filter(tf == "NFYA", commercial_reporter == "No")%>%
           mutate(promoter = gsub(".*(Random|minP|hBGm|mCMV).*", "\\1", reporter_id)) %>%
           distinct(cell, reporter_id, reporter_activity_minP, TPM_norm, promoter, cor_pval),
       aes(x = TPM_norm, y = 2^reporter_activity_minP)) +
  geom_point(alpha = .2, aes(color = cor_pval)) +
  geom_smooth(method = "lm", se = F, aes(group = reporter_id, color = cor_pval)) +
  theme_pubr(border = T) +
  facet_wrap(~promoter) +
  scale_color_gradient(low = "#264653", high = "#9AC1AE")

tfs_reasonable <- rna_correlations_df %>%
  filter(commercial_reporter == "No") %>%
  distinct(tf, reasonable) %>%
  filter(reasonable == "Yes")

ggplot(cDNA_df3 %>%
         filter(commercial_reporter == "No", tf %in% tfs_reasonable$tf) %>%
         mutate(promoter = gsub(".*(Random|minP|hBGm|mCMV).*", "\\1", reporter_id)) %>%
         left_join(tf_properties) %>%
         distinct(reporter_id, promoter, cor_pval, chromatin_regulation_mode),
       aes(x = promoter, y = cor_pval)) +
  geom_quasirandom() +
  theme_pubr(border = T) +
  facet_wrap(~chromatin_regulation_mode) +
  scale_color_gradient(low = "#264653", high = "#9AC1AE")



ggplot(cDNA_df3 %>%
           filter(reporter_id %in% c("POU5F1::SOX2_10bp_21bp_minP_1", "POU5F1::SOX2_5bp_10bp_mCMV_1")) %>%
           distinct(cell, reporter_id, reporter_activity_minP, TPM_norm),
       aes(x = TPM_norm, y = reporter_activity_minP)) +
    geom_abline(lty = 2) +
    geom_smooth(method = "lm", color = "black") +
    geom_point(size = 2) +
    facet_wrap(~reporter_id, ncol = 1) +
    theme_pubr(border = T)

ggplot(cDNA_df3 %>%
         filter(reporter_id %in% c("HNF4A_10bp_10bp_mCMV_2", "HNF4A_5bp_10bp_minP_1")) %>%
         distinct(cell, reporter_id, reporter_activity_minP, TPM_norm),
       aes(x = TPM_norm, y = 2^reporter_activity_minP)) +
  geom_smooth(method = "lm", color = "black") +
  geom_point(size = 3, aes(color = cell)) +
  scale_color_manual(values = cell_colors) +
  facet_wrap(~reporter_id, ncol = 1) +
  theme_pubr(border = T)

ggplot(cDNA_df3 %>%
         filter(reporter_id %in% c("GATA4_10bp_10bp_minP_3", "GATA4_5bp_10bp_minP_2")) %>%
         distinct(cell, reporter_id, reporter_activity_minP, TPM_norm),
       aes(x = TPM_norm, y = 2^reporter_activity_minP)) +
  geom_abline(lty = 2) +
  geom_smooth(method = "lm", color = "black") +
  geom_point(size = 2) +
  facet_wrap(~reporter_id, ncol = 1) +
  theme_pubr(border = T)

ggplot(cDNA_df3 %>%
         filter(reporter_id %in% c("ESR1_5bp_10bp_mCMV_2", "ESR1_10bp_21bp_mCMV_3")) %>%
         distinct(cell, reporter_id, reporter_activity_minP, TPM_norm),
       aes(x = TPM_norm, y = 2^reporter_activity_minP)) +
  geom_smooth(method = "lm", color = "black") +
  geom_point(size = 3, aes(color = cell)) +
  scale_color_manual(values = cell_colors) +
  facet_wrap(~reporter_id, ncol = 1) +
  theme_pubr(border = T)


ggplot(cDNA_df3 %>%
         filter(reporter_id %in% c("THRB_5bp_10bp_mCMV_3", "THRB_10bp_10bp_mCMV_3")) %>%
         distinct(cell, reporter_id, reporter_activity_minP, TPM_norm),
       aes(x = TPM_norm, y = 2^reporter_activity_minP)) +
  geom_abline(lty = 2) +
  geom_smooth(method = "lm", color = "black") +
  geom_point(size = 2) +
  facet_wrap(~reporter_id, ncol = 1) +
  theme_pubr(border = T)

ggplot(cDNA_df3 %>%
         filter(reporter_id %in% c("FOXO1_10bp_21bp_minP_2", "FOXO1_5bp_21bp_minP_2")) %>%
         distinct(cell, reporter_id, reporter_activity_minP, TPM_norm),
       aes(x = TPM_norm, y = 2^reporter_activity_minP)) +
  geom_abline(lty = 2) +
  geom_smooth(method = "lm", color = "black") +
  geom_point(size = 2) +
  facet_wrap(~reporter_id, ncol = 1) +
  theme_pubr(border = T)

ggplot(cDNA_df3 %>%
         filter(reporter_id %in% c("TFCP2L1_5bp_10bp_mCMV_1", "TFCP2L1_5bp_10bp_minP_2")) %>%
         distinct(cell, reporter_id, reporter_activity_minP, TPM_norm),
       aes(x = TPM_norm, y = 2^reporter_activity_minP)) +
  geom_abline(lty = 2) +
  geom_smooth(method = "lm", color = "black") +
  geom_point(size = 2) +
  facet_wrap(~reporter_id, ncol = 1) +
  theme_pubr(border = T)


ggplot(cDNA_df3 %>%
         filter(reporter_id %in% c("GATA1_5bp_10bp_mCMV_2", "GATA1_5bp_21bp_minP_1")) %>%
         distinct(cell, reporter_id, reporter_activity_minP, TPM_norm),
       aes(x = TPM_norm, y = 2^reporter_activity_minP)) +
  geom_abline(lty = 2) +
  geom_smooth(method = "lm", color = "black") +
  geom_point(size = 2) +
  facet_wrap(~reporter_id, ncol = 1) +
  theme_pubr(border = T)




# ## Tissue specificity of top TFs
# chosen_tfs <- cDNA_df6 %>%
#   distinct(tf, quality_score) %>%
#   mutate(quality_score = ave(quality_score, tf, FUN = mean)) %>%
#   distinct() %>%
#   arrange(desc(quality_score)) %>%
#   top_n(n = 50)
# 
# tf_expression_pattern <- read_csv("/DATA/usr/m.trauernicht/projects/SuRE-TF/data/Fig4A.csv") %>%
#   pivot_longer(cols = -Sample, names_to = "tf", values_to = "expression") %>%
#   filter(tf %in% chosen_tfs$tf)
# 
# 
# ggplot(tf_expression_pattern %>%
#          mutate(expression_mean = ave(expression, Sample, FUN = mean)),
#        aes(x = reorder(Sample, -expression_mean),  y = expression)) +
#   geom_quasirandom() +
#   theme_classic_lines_90()
# 
# 
# cancer_models <- read_csv("/DATA/usr/m.trauernicht/projects/SuRE-TF/data/CRISPRGeneEffect.csv") %>%
#   pivot_longer(cols = -`...1`, names_to = "tf", values_to = "expression") %>%
#   mutate(tf = gsub(" (.*)", "", tf)) %>%
#   filter(tf %in% chosen_tfs$tf)
# 
# 
# cancer_models <- cancer_models %>%
#   mutate(mean_effect = ave(expression, `...1`, FUN = mean))
# 
# 
# ggplot(cancer_models,
#        aes(x = reorder(`...1`, -mean_effect),  y = expression)) +
#   geom_quasirandom() +
#   theme_classic_lines_90()
```

---

## Add proteomics data

Aim: Similar to before: does reporter activity correlate with protein abundance across cell types?
```{r, warning=F}
proteomics_df_long <- read_tsv("/DATA/usr/m.trauernicht/data/RNA_seq/protein_abundance_all_tfs.tsv")

## Combine TF reporter data with RNA-seq data
cDNA_df3 <- merge(cDNA_df2, proteomics_df_long, all = T) %>%
  na.omit()

## Compute spearman correlations for each individual reporter of reporter activity vs. TPM counts across the nine cell types
for (i in unique(cDNA_df3$reporter_id)) {
  if (nrow(cDNA_df3[cDNA_df3$reporter_id == i,]) > 4) {
    
    cDNA_df3$cor_pval[cDNA_df3$reporter_id == i] <- cor.test(cDNA_df3$reporter_activity_minP[cDNA_df3$reporter_id == i],
                                                             cDNA_df3$protein_abundance[cDNA_df3$reporter_id == i], method = "pearson", exact = F)$p.value
    
    cDNA_df3$cor[cDNA_df3$reporter_id == i] <- cor(cDNA_df3$reporter_activity_minP[cDNA_df3$reporter_id == i],
                                                   cDNA_df3$protein_abundance[cDNA_df3$reporter_id == i], method = "pearson")
  }
  else {
    cDNA_df3$cor_pval[cDNA_df3$reporter_id == i] <- 1
    cDNA_df3$cor[cDNA_df3$reporter_id == i] <- 0
  }
}

## I am only interested in positive correlations, so set all negative correlations to zero
cDNA_df3 <- cDNA_df3 %>%
  mutate(cor_pval = ifelse(cor_pval == cor, 1, cor_pval)) %>%
  group_by(tf) %>%
  mutate(cor_pval_fdr = p.adjust(cor_pval, method = "fdr")) %>%
  ungroup() %>%
  mutate(cor_pval = ifelse(cor < 0, 1, cor_pval)) %>%
  mutate(cor_pval_fdr = ifelse(cor < 0, 1, cor_pval_fdr)) %>%
  mutate(cor_pval = -log10(cor_pval)) %>%
  mutate(cor_pval_fdr = -log10(cor_pval_fdr)) %>%
  mutate(cor_pval = ifelse(is.na(cor_pval), 1, cor_pval)) %>%
  mutate(cor_pval_fdr = ifelse(is.na(cor_pval_fdr), 1, cor_pval_fdr)) %>%
  mutate(cor = ifelse(is.na(cor), 1, cor))

## Export correlations
protein_correlations <- cDNA_df3 %>%
  distinct(reporter_id, cor_pval, cor)

## Correlate RNA and protein correlations
rna_protein_cor <- merge(protein_correlations %>%
                           dplyr::select(reporter_id, "prot_cor" = cor_pval),
                         rna_correlations %>%
                           dplyr::select(reporter_id, "rna_cor" = cor_pval), all = T) %>%
  replace(is.na(.), 0)

ggplot(rna_protein_cor %>%
         mutate(tf = gsub("(.*)_.*_.*_.*_.*", "\\1", reporter_id)) %>%
         mutate(source = gsub(".*_(.*)", "\\1", tf)) %>%
         mutate(source = ifelse(source == "TF-romanov", "romanov", source)) %>%
         mutate(source = ifelse(!source %in% c("romanov", "TF-seq", "promega"), "designed", source)) %>%
         mutate(tf = gsub("_.*", "", tf)),
       aes(x = prot_cor, y = rna_cor)) +
  geom_point(aes(color = source)) +
  geom_smooth(method = "lm") +
  theme_pubr() +
  facet_wrap(~tf)

protein_correlations2 <- protein_correlations %>%
  dplyr::select(reporter_id, "protein_cor" = cor)

ggplot(rna_correlations_df %>%
         left_join(protein_correlations2) %>%
         filter(commercial_reporter == "No") %>%
         distinct(reporter_id, cor_pval, tf, protein_cor) %>%
         mutate(sign_rna = ifelse(cor_pval < 1, "No", "Yes")) %>%
         mutate(sign_protein = ifelse(protein_cor < 1, "No", "Yes")) %>%
         mutate(mean_cor_pval = ave(cor_pval, tf, FUN = mean)),
       aes(x = reorder(tf, -mean_cor_pval), y = cor_pval)) +
  geom_quasirandom(aes(color = protein_cor)) +
  geom_hline(yintercept = 1, lty = 2) +
  scale_color_gradient2(low = "#264653", mid = "white", high = "#e76f51") +
  theme_pubr(border = T, x.text.angle = 90)

ggplot_custom(cDNA_df3 %>%
                filter(commercial_reporter == "No") %>%
                distinct(reporter_id, cor_pval, tf, commercial_reporter) %>%
                mutate(mean_cor_pval = ave(cor_pval, tf, FUN = mean)) %>%
                mutate(sign = ifelse(cor_pval < 1.30103, "No", "Yes")) %>%
                left_join(rna_correlations_df %>% dplyr::select(reporter_id, "cor_rna" = cor_pval) %>% distinct()),
       aes(x = reorder(tf, -mean_cor_pval), y = cor_pval)) +
  geom_quasirandom(aes(color = cor_rna)) +
  geom_hline(yintercept = 1.30103, lty = 2) +
  scale_color_gradient(low = "grey90",high = "#e76f51") +
  theme_pubr(border = T, x.text.angle = 90)

ggplot(rna_correlations_df %>%
         left_join(protein_correlations2) %>%
         filter(commercial_reporter == "No") %>%
         distinct(reporter_id, cor, tf, protein_cor) %>%
         mutate(sign_rna = ifelse(cor < 1, "No", "Yes")) %>%
         mutate(sign_protein = ifelse(protein_cor < 1, "No", "Yes")) %>%
         mutate(mean_cor_pval = ave(cor, tf, FUN = mean)),
       aes(x = reorder(tf, -mean_cor_pval), y = cor)) +
  geom_quasirandom(aes(color = protein_cor)) +
  geom_hline(yintercept = 1, lty = 2) +
  scale_color_gradient2(low = "#264653", mid = "white", high = "#e76f51") +
  theme_pubr(border = T, x.text.angle = 90)
  

ggplot_custom(cDNA_df3 %>%
                filter(commercial_reporter == "No") %>%
                distinct(reporter_id, cor_pval, tf, commercial_reporter) %>%
                mutate(mean_cor_pval = ave(cor_pval, tf, FUN = mean)) %>%
                mutate(sign = ifelse(cor_pval < 1.30103, "No", "Yes")) %>%
                filter(tf %in% c("HNF4A", "GATA4", "ESR1", "THRB", "GATA1")),
       aes(x = reorder(tf, -mean_cor_pval), y = cor_pval, color = sign)) +
  geom_quasirandom() +
  geom_hline(yintercept = 1.30103, lty = 2) +
  theme_pubr(border = T)

ggplot(cDNA_df3 %>%
         filter(commercial_reporter == "No", tf == "NFKB1", cell == "HepG2") %>%
         distinct(reporter_id, cor_pval, reporter_activity_minP) %>%
         mutate(spacer_length = gsub("NFKB1_(.*bp)_.*bp.*", "\\1", reporter_id)) %>%
         mutate(background = gsub(".*_([1-3]$)", "\\1", reporter_id)),
       aes(x = reporter_activity_minP, y = cor_pval, color = spacer_length)) +
  geom_abline(lty = 2) +
  geom_point(size = 2) +
  theme_pubr() +
  theme(legend.position = "none") +
  scale_color_manual(values = c("10bp" = "#7c9eb2", "5bp" = "grey80"))

ggplot(cDNA_df3 %>%
           filter(reporter_id %in% c("GATA1_5bp_10bp_minP_3", "GATA1_10bp_10bp_minP_3")) %>%
           distinct(cell, reporter_id, reporter_activity_minP, protein_abundance),
       aes(x = protein_abundance, y = reporter_activity_minP)) +
    geom_abline(lty = 2) +
    geom_smooth(method = "lm", color = "black") +
    geom_point(size = 2) +
    facet_wrap(~reporter_id, ncol = 1) +
    theme_pubr(border = T)

ggplot(cDNA_df3 %>%
           filter(reporter_id %in% c("HOMEZ_10bp_10bp_minP_1", "HOMEZ_5bp_10bp_minP_1")) %>%
           distinct(cell, reporter_id, reporter_activity_minP, protein_abundance),
       aes(x = protein_abundance, y = reporter_activity_minP)) +
    geom_abline(lty = 2) +
    geom_smooth(method = "lm", color = "black") +
    geom_point(size = 2) +
    facet_wrap(~reporter_id, ncol = 1) +
    theme_pubr(border = T)

ggplot(cDNA_df3 %>%
         filter(reporter_id %in% c("NR3C1_10bp_21bp_hBGm_3", "NR3C1_10bp_10bp_minP_2")) %>%
         distinct(cell, reporter_id, reporter_activity_minP, protein_abundance),
       aes(x = protein_abundance, y = reporter_activity_minP)) +
  geom_abline(lty = 2) +
  geom_smooth(method = "lm", color = "black") +
  geom_point() +
  facet_wrap(~reporter_id, ncol = 1) +
  theme_pubr(border = T)

ggplot(cDNA_df3 %>%
         filter(reporter_id %in% c("STAT3_10bp_10bp_minP_3", "STAT3_10bp_10bp_minP_1")) %>%
         distinct(cell, reporter_id, reporter_activity_minP, protein_abundance),
       aes(x = protein_abundance, y = reporter_activity_minP)) +
  geom_abline(lty = 2) +
  geom_smooth(method = "lm", color = "black") +
  geom_point(size = 2) +
  facet_wrap(~reporter_id, ncol = 1) +
  theme_pubr(border = T)

ggplot(cDNA_df3 %>%
         filter(reporter_id %in% c("WT1_10bp_21bp_mCMV_1", "WT1_5bp_21bp_mCMV_1")) %>%
         distinct(cell, reporter_id, reporter_activity_minP, protein_abundance),
       aes(x = protein_abundance, y = reporter_activity_minP)) +
  geom_abline(lty = 2) +
  geom_smooth(method = "lm", color = "black") +
  geom_point() +
  facet_wrap(~reporter_id, ncol = 1) +
  theme_pubr(border = T)

ggplot(cDNA_df3 %>%
         filter(reporter_id %in% c("SMAD4_10bp_21bp_minP_2", "SMAD4_5bp_21bp_minP_2")) %>%
         distinct(cell, reporter_id, reporter_activity_minP, protein_abundance),
       aes(x = protein_abundance, y = reporter_activity_minP)) +
  geom_abline(lty = 2) +
  geom_smooth(method = "lm", color = "black") +
  geom_point(size = 2) +
  facet_wrap(~reporter_id, ncol = 1) +
  theme_pubr(border = T)

ggplot(cDNA_df3 %>%
         filter(reporter_id %in% c("SRF_10bp_21bp_mCMV_3", "SRF_5bp_21bp_mCMV_3")) %>%
         distinct(cell, reporter_id, reporter_activity_minP, protein_abundance),
       aes(x = protein_abundance, y = reporter_activity_minP)) +
  geom_abline(lty = 2) +
  geom_smooth(method = "lm", color = "black") +
  geom_point(size = 2) +
  facet_wrap(~reporter_id, ncol = 1) +
  theme_pubr(border = T)

ggplot(cDNA_df3 %>%
         filter(reporter_id %in% c("SOX2_5bp_10bp_minP_2", "SOX2_10bp_10bp_minP_2")) %>%
         distinct(cell, reporter_id, reporter_activity_minP, protein_abundance),
       aes(x = protein_abundance, y = reporter_activity_minP)) +
  geom_abline(lty = 2) +
  geom_smooth(method = "lm", color = "black") +
  geom_point(size = 2) +
  facet_wrap(~reporter_id, ncol = 1) +
  theme_pubr(border = T)

ggplot(cDNA_df3 %>%
         filter(reporter_id %in% c("RXRA_10bp_21bp_mCMV_3", "RXRA_5bp_21bp_mCMV_3")) %>%
         distinct(cell, reporter_id, reporter_activity_minP, protein_abundance),
       aes(x = protein_abundance, y = reporter_activity_minP)) +
  geom_abline(lty = 2) +
  geom_smooth(method = "lm", color = "black") +
  geom_point(size = 2) +
  facet_wrap(~reporter_id, ncol = 1) +
  theme_pubr(border = T)

ggplot(cDNA_df3 %>%
         filter(reporter_id %in% c("OTX1_5bp_10bp_minP_2", "OTX1_10bp_10bp_minP_2")) %>%
         distinct(cell, reporter_id, reporter_activity_minP, protein_abundance),
       aes(x = protein_abundance, y = reporter_activity_minP)) +
  geom_abline(lty = 2) +
  geom_smooth(method = "lm", color = "black") +
  geom_point(size = 2) +
  facet_wrap(~reporter_id, ncol = 1) +
  theme_pubr(border = T)
```
Some reporter activities correlate surprisingly well with protein abundance.

---


## Linear model - TF reporter grammar

Aim: If the measured activities can be fitted by a linear model, it is likely that the activity is true. If not, we might be looking at noise instead.

```{r log-linear-model-5, out.width= "80%", fig.align= "center", echo=FALSE, warning= FALSE, message = FALSE}
## Only include 2-3 most interesting conditions per TF (as determined before)
ref_conditions_sel <- ref_conditions %>%
  mutate(tf_condition = paste(tf, condition, sep = "_"))

cDNA_df_TF <- cDNA_df %>%
  filter(neg_ctrls == "No", commercial_reporter == "No", hPGK == "No", native_enhancer == "No", stimulation %in% c("no", "pathway", "overexpression"), promoter != "Random") %>%
  distinct(reporter_activity_minP, promoter, spacing, distance, background, tf, condition, reporter_id, stimulation, reference_condition) %>%
  mutate(tf_condition = paste(tf, condition, sep = "_")) %>%
  filter(tf_condition %in% ref_conditions_sel$tf_condition) %>%
  # mutate(max_activity = ave(reporter_activity_minP, tf_condition, FUN = max)) %>%
  # mutate(min_activity = ave(reporter_activity_minP, tf_condition, FUN = min)) %>%
  # mutate(dif = max_activity - min_activity) %>%
  # filter(dif > 2.5) %>%
  mutate(reporter_activity_minP = log2(reporter_activity_minP)) %>%
  left_join(reference_condition %>% dplyr::select(-tf), by = c("reporter_id", "reference_condition")) %>%
  mutate(reporter_dif =  reference_activity - reporter_activity_minP) %>%
  mutate(background = as.character(background)) %>%
  mutate(promoter = as.factor(promoter)) %>%
  mutate(spacing = ifelse(spacing == "10bp", "10bps", "5bp")) %>%
  mutate(spacing = as.factor(spacing)) %>%
  #mutate(reporter_activity_minP = ifelse(stimulation != "no", reporter_dif, reporter_activity_minP)) %>%
  distinct(reporter_activity_minP, promoter, spacing, distance, background, condition, tf, tf_condition, reporter_id)


## Fit linear model with design features for actual data
exp_variance <- data.frame("feature" = c("promoter", "spacing", "distance", "background", "Residuals"))
weight <- data.frame("feature" = c("(Intercept)", "promoterminP", "promoterhBGm", "promotermCMV",
                  "spacing10bps","spacing5bp", "distance10bp","distance21bp", "background1","background2", "background3"))

cDNA_df_TF$promoter <- relevel(cDNA_df_TF$promoter, ref = "minP")
cDNA_df_TF$spacing <- relevel(cDNA_df_TF$spacing, ref = "5bp")
cDNA_df_TF$row <- rownames(cDNA_df_TF)
cDNA_df_TF3 <- data.frame()

lm_list <- list()
for (i in unique(cDNA_df_TF$tf_condition)) {
  if (nrow(cDNA_df_TF[cDNA_df_TF$tf_condition == i,]) > 9) {
  print(i)
  lm_list[[i]] <- lm(reporter_activity_minP ~
                       distance + promoter + spacing + background,
                cDNA_df_TF[cDNA_df_TF$tf_condition == i,])
  }
}

f <- lapply(names(lm_list), function(x) summary(lm_list[[x]])$fstatistic)
names(f) <- names(lm_list)
p <- lapply(names(f), function(x) pf(f[[x]][1],f[[x]][2],f[[x]][3],lower.tail=F))
p <- data.frame(bind_rows(p)) %>% setnames("value", "pval")
p$tf_condition <- names(lm_list)

y <- lapply(names(lm_list), function(x) data.frame("activity_predicted" = lm_list[[x]]$fitted.values) %>% rownames_to_column("row") %>% mutate(tf_condition = x))
y <- data.frame(bind_rows(y)) %>% left_join(p, by = "tf_condition")

cDNA_df_TF2 <- y %>%
  left_join(cDNA_df_TF)

cDNA_df_TF3 <- rbind(cDNA_df_TF3, cDNA_df_TF2)

mean(cDNA_df_TF3$pval)

w <- lapply(names(lm_list), function(x) 
  if (length(lm_list[[x]]$coefficients) == length(summary(lm_list[[x]])$coefficients[,4])) {
  data.frame('weight' = lm_list[[x]]$coefficients, 'pval' = summary(lm_list[[x]])$coefficients[,4]) %>% mutate(tf = x) %>% rownames_to_column('feature')
  }
)

w <- data.frame(bind_rows(w))
weight <- weight %>%
  left_join(w)

x <- lapply(names(lm_list), function(x) anova(lm_list[[x]]) %>% rownames_to_column("feature") %>% mutate(tf = x))
x <- data.frame(bind_rows(x)) %>% 
  setnames(c("Sum.Sq", "Pr..F."), c("sum_sq", "pval")) %>%
  distinct(feature, sum_sq, pval, tf) %>%
  mutate(sum = ave(sum_sq, tf, FUN = sum)) %>%
  mutate(rel_sum_sq = (sum_sq/sum)*100) %>%
  distinct(tf, rel_sum_sq, pval, feature)

exp_variance <- exp_variance %>%
  left_join(x)

## Fine-tune dfs
exp_variance$group <- "no residual"
exp_variance$group[grep("Residual", exp_variance$feature)] <- "residual"
exp_variance$rel_sum_sq[exp_variance$feature == "Residuals"] <- 100 - exp_variance$rel_sum_sq[exp_variance$feature == "Residuals"]
exp_variance$feature <- revalue(exp_variance$feature, c("Residuals"="total"))
prediction <- cDNA_df_TF3 %>%
  dplyr::select(-row) %>%
  na.omit()
weight <- weight %>%
  filter(feature != "(Intercept)")
categorie <- c("promoter", "spacing", "distance", "background")
weight$features <- gsub(paste(categorie, collapse="|"), "",weight$feature)
ml <- length(unique(weight$features))
feature <- unique(weight$features)
weight$condition <- gsub(paste(feature, collapse = "|"), "", weight$feature)
weight <- weight %>% dplyr::select(-feature)


cDNA_df_TF3 <- cDNA_df_TF3 %>%
  mutate(pval_adj = p.adjust(pval, method = "fdr"))
```


## Identify best condition per TF based on linear models and other data

```{r log-linear-model-5, out.width= "80%", fig.align= "center", echo=FALSE, warning= FALSE, message = FALSE}
## Then integrate the linear models to select the least noisy of the top 2-3 conditions per TF
cDNA_df_TF4 <- cDNA_df_TF3 %>%
  distinct(tf, condition, pval_adj)

chosen_conditions <- ref_conditions %>%
  left_join(cDNA_df_TF4) %>%
  replace(is.na(.), 1) %>%
  group_by(tf) %>%
  slice_min(n = 1, order_by = pval_adj, with_ties = T) %>%
  ungroup() %>%
  mutate(cell = gsub("_.*", "", condition)) %>%
  left_join(tf_rna) %>%
  group_by(tf) %>%
  slice_max(n = 1, order_by = TPM_norm, with_ties = F) %>%
  ungroup() %>%
  distinct(tf, condition) %>%
  mutate(tf_condition = paste(tf, condition, sep = "_"))

## Change some conditions manually
### NFAT5: Even though the synthetic reporters don't seem to respond to Calcimycin, the published reporters do
### NFAT5: since the published reporters are not taken along in the linear modeling, I should select this condition manually
chosen_conditions$tf_condition[chosen_conditions$tf_condition == "NFAT5_mES_2i_LIF"] <- "NFAT5_HCT116_Calcimycin"
```


## RNA correlation analysis
```{r}
## Repeat RNA correlation plot - highlight predicted activities
cDNA_df_TF5 <- cDNA_df_TF3 %>%
  distinct(tf, condition, reporter_id, activity_predicted) %>%
  mutate(tf_condition = paste(tf, condition, sep = "_")) %>%
  filter(tf_condition %in% chosen_conditions$tf_condition) %>%
  mutate(activity_predicted = 2^activity_predicted) %>%
  mutate(activity_predicted_rel = ave(-activity_predicted, tf, FUN = function(x) rank(x)))

ggplot(rna_correlations_df %>%
         left_join(cDNA_df_TF5) %>%
         filter(commercial_reporter == "No", str_detect(reporter_id, "andom", negate = T)) %>%
         distinct(reporter_id, cor_pval, tf, activity_predicted_rel) %>%
         mutate(sign_rna = ifelse(cor_pval < 1, "No", "Yes")) %>%
         mutate(mean_cor_pval = ave(cor_pval, tf, FUN = mean)),
       aes(x = reorder(tf, -mean_cor_pval), y = cor_pval)) +
  geom_quasirandom(aes(color = activity_predicted_rel)) +
  geom_hline(yintercept = 1, lty = 2) +
  scale_color_gradient(low = "#B2182B", high = "#FDEDEF", na.value = "grey70", limits = c(1,24), oob = squish) +
  theme_pubr(border = T, x.text.angle = 90)

protein_correlations3 <- cDNA_df3 %>%
  distinct(reporter_id, cor_pval, cor, commercial_reporter)

ggplot(protein_correlations3 %>%
         left_join(cDNA_df_TF5) %>%
         filter(commercial_reporter == "No", str_detect(reporter_id, "andom", negate = T)) %>%
         distinct(reporter_id, cor_pval, tf, activity_predicted_rel) %>%
         mutate(sign_rna = ifelse(cor_pval < 1, "No", "Yes")) %>%
         mutate(mean_cor_pval = ave(cor_pval, tf, FUN = mean)),
       aes(x = reorder(tf, -mean_cor_pval), y = cor_pval)) +
  geom_quasirandom(aes(color = activity_predicted_rel)) +
  geom_hline(yintercept = 1, lty = 2) +
  scale_color_gradient(low = "#FDEDEF", high = "#B2182B", na.value = "grey70") +
  theme_pubr(border = T, x.text.angle = 90)
```


## Explore linear models for all TFs in the chosen conditions

```{r log-linear-model-5, out.width= "80%", fig.align= "center", echo=FALSE, warning= FALSE, message = FALSE}
## Example: Predicted vs. measured for one TF
ggplot(cDNA_df_TF3 %>%
         filter(tf == "STAT3", condition == "mES_2i_LIF"),
       aes(x = activity_predicted, y = reporter_activity_minP, color = background)) +
  geom_abline(lty = 2) +
  geom_point(size = 2) +
  scale_color_manual(values = c("black", "grey50", "grey80")) +
  facet_wrap(~spacing) +
  scale_x_continuous(breaks = c(0,2,4))+
  scale_y_continuous(breaks = c(0,2,4))

ggplot(cDNA_df_TF3 %>%
         filter(tf == "STAT3", condition == "mES_2i_LIF"),
       aes(x = activity_predicted, y = reporter_activity_minP, color = spacing)) +
  geom_abline(lty = 2) +
  geom_point(size = 2) +
  scale_color_manual(values = c("10bps" = "#7C9EB2", "5bp" = "#BFD0D9")) +
  scale_x_continuous(breaks = c(0,2,4))+
  scale_y_continuous(breaks = c(0,2,4))


weight_ex <- weight %>%
  filter(tf == "STAT3_mES_2i_LIF")

zero_weights <- weight %>%
  filter(is.na(tf)) %>%
  mutate(weight = 0, pval = 1) %>%
  mutate(tf = "STAT3_mES_2i_LIF")
                   
weight_ex <- rbind(weight_ex, zero_weights)
                   
feature_names <- data.frame("features" = c("minP", "mCMV", "hBGm", "1", "2", "3", "5bp", "10bps", "10bp", "21bp"),
                            "feature_names" = c("minP", "minCMV", "minHBG", "Spacer 1", "Spacer 2", "Spacer 3", "5bp Spacer", 
                                                "10bp Spacer", "10bp Distance", "21bp Distance"))
                     
weight_ex <- weight_ex %>%
  left_join(feature_names)
                   

ggplot(weight_ex %>% 
         mutate(sign = ifelse(pval < 0.05, "Yes", "No")) %>%
         mutate(feature_names = factor(feature_names, levels = c("minP", "minCMV", "minHBG", "Spacer 1", "Spacer 2", "Spacer 3", "5bp Spacer", 
                                                                                                        "10bp Spacer", "10bp Distance", "21bp Distance"))),
       aes(x = feature_names, y = weight, alpha = sign)) +
  geom_hline(yintercept = 0, lty = 2) +
  geom_bar_interactive(stat = "identity", color = "black") +
  ylab("Weight") + 
  xlab("Reporter design feature") +
  theme_classic_lines_45() +
  scale_alpha_discrete(range = c(0.3,1))

## Explore weights for all TFs
cDNA_df_TF5 <- cDNA_df_TF4 %>%
  mutate(tf = paste(tf, condition, sep = "_")) %>%
  distinct(tf, pval_adj)

### Add total variances as bar plot
exp_variance2 <- exp_variance %>%
  filter(tf %in% chosen_conditions$tf_condition) %>%
  filter(feature == "total") %>%
  dplyr::select(tf, "total_variance" = rel_sum_sq)

weight2 <- weight %>%
  filter(tf %in% chosen_conditions$tf_condition) %>%
  left_join(cDNA_df_TF5) %>%
  #filter(pval_adj < .005) %>%
  left_join(exp_variance2) %>%
  filter(total_variance > 52 & pval_adj < 0.05) %>%
  mutate(tf = gsub("_.*", "", tf)) %>%
  mutate(weight = ifelse(pval > 0.1, 0, weight)) %>%
  distinct(tf, features, weight) %>%
  spread(key = "features", value = "weight") %>%
  column_to_rownames("tf") %>%
  filter(rowSums(., na.rm = T) != 0)

myBreaks <- c(seq(-2, 0, length.out=50+1), 
              seq(0.000001, 2, length.out=50))

pheatmap(weight2 %>%
                  t(),
         clustering_method = "complete",
         cellwidth = 10,
         cellheight = 10,
         border_color = F,
         color = colorRampPalette(rev(brewer.pal(n = 7, name = "RdBu")))(100),
         angle_col = 90,
         cluster_cols = T,
         cutree_rows = 4,
         breaks = myBreaks)

hc <- hclust(dist(weight2), method = "complete")
order_dendogram <- data.frame("order" = 1:nrow(weight2),
                              "tf" = "")

for(i in unique(order_dendogram$order)) {
  order_dendogram$tf[order_dendogram$order == i] <- rownames(weight2[hc$order[i],])
}

exp_variance_2 <- exp_variance %>%
  filter(tf %in% chosen_conditions$tf_condition) %>%
  mutate(tf = gsub("_.*", "", tf)) %>%
  filter(tf %in% rownames(weight2)) %>%
  filter(feature == "total") %>%
  dplyr::select(tf, "total_variance" = rel_sum_sq) %>%
  left_join(order_dendogram)

ggplot(exp_variance_2,
       aes(x = reorder(tf, order), y = "total_variance", fill = total_variance)) +
  geom_tile() +
  coord_fixed() +
  scale_fill_gradient(low = "white", high = "black", limits = c(33, 100)) +
  theme_pubr(border = T, x.text.angle = 90)

weight3 <- weight2 %>%
  rownames_to_column("tf") %>%
  pivot_longer(cols = -tf, names_to = "features", values_to = "weight") %>%
  distinct(tf, features, weight) %>%
  left_join(feature_names) %>%
  left_join(exp_variance_2) %>%
  left_join(order_dendogram)

ggplot(weight3 %>%
         mutate(feature_names = factor(feature_names, levels = rev(c("minP", "minCMV", "minHBG", "Spacer 1", "Spacer 2", "Spacer 3", "5bp Spacer", 
                                                                 "10bp Spacer", "10bp Distance", "21bp Distance")))),
       aes(x = reorder(tf, order), y = feature_names, fill = weight)) +
  geom_tile() +
  coord_fixed() +
  scale_fill_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", na.value = "grey80", limits = c(-2, 2), oob = squish) +
  theme_pubr(border = T, x.text.angle = 90)
  


## Investigate hits with strong impact on spacer length
weight4 <- weight %>%
  filter(tf %in% chosen_conditions$tf_condition) %>%
  left_join(cDNA_df_TF5) %>%
  #filter(pval_adj < .001) %>%
  mutate(tf = gsub("_.*", "", tf)) %>%
  filter(tf %in% rownames(weight2)) %>%
  filter(condition == "spacing", pval < 0.05, abs(weight) >= 0.2)

actual_spacer_lengths <- data.frame(tf = c("CEBPB", "CREB1", "ETS2", "FOS::JUN", "GRHL1", "HNF1A", "HNF4A", "KLF4", "MAF::NFE2", "MTF1", "NFE2L2", 
                                            "NFKB1", "NR1H4::RXRA", "NR1I2", "REL", "SRF", "STAT3", "TCF7L2", "TFCP2L1"),
                                    center_center_dist = c(10, 8, 20, 10, 10, 15, 16, 8, 15, 17, 10, 11, 15, 17, 10, 12, 9, 10, 14),
                                    monomer_center_dist = c(5, 4, 8, 6, 5, 8, 9, 8, 10, 17, 5, 6, 8, 8, 5, 6, 4, 10, 6),
                                    mode = c("homodimer", "homodimer", "homodimer", "heterodimer", "homodimer", "homodimer", "homodimer", "monomer", "heterodimer", 
                                             "monomer", "heterodimer", "homodimer", "heterodimer", "heterodimer", 
                                             "homodimer", "homodimer", "homodimer", "monomer", "tetramer"))

binding_modes <- data.frame(tf = c("CEBPB", "CREB1", "E2F1", "ESR1", "ETS2", "FOS::JUN", "GRHL1", "HNF1A", "HNF4A", "IRF3", "KLF4", "MAF::NFE2", "MTF1",
                                                            "NFAT5", "NFE2L2", "NFIA",
                                            "NFKB1", "NFYA", "NR1H4::RXRA", "NR1I2",  "POU2F1", "POU5F1::SOX2", "RARA", "RBPJ", "REL", "RXRA", 
                                             "SOX2", "SP1", "SRF", "STAT1::STAT2", "STAT3", "TCF7L2", "TFCP2L1", "TP53", "VDR::RXRA"),
                                    mode = c("homodimer", "homodimer", "homodimer", "heterodimer", "homodimer", "heterodimer", "homodimer", "homodimer", "homodimer", "homodimer", 
                                             "monomer", "heterodimer", 
                                             "monomer", "homodimer", "heterodimer", "homodimer", "homodimer", "heterodimer", "heterodimer", "heterodimer", "homodimer", "heterodimer", 
                                             "heterodimer", "monomer", 
                                             "homodimer", "homodimer", "monomer", "monomer", "homodimer", "heterodimer", "homodimer", "monomer", "tetramer", "tetramer", "heterodimer"))

tf_properties <- read_csv2("/DATA/usr/m.trauernicht/projects/SuRE-TF/data/library_design/lambert_2018/garcia_tf-properties.csv") %>%
  dplyr::select("tf" = TF, DNA_binding_mode, chromatin_regulation_mode, TF_superclass, TF_class)

weight_spacer <- weight4 %>%
  left_join(actual_spacer_lengths) %>%
  mutate(center_center_dist = center_center_dist + 5, monomer_center_dist = monomer_center_dist + 5)

ggplot(weight_spacer,
       aes(x = monomer_center_dist, y = weight)) +
  geom_hline(yintercept = 0, lty = 2) +
  geom_smooth(method = "lm", se = F) +
  geom_point(size = 2) +
  theme_pubr(border = T)

ggplot(weight_spacer,
       aes(x = center_center_dist, y = weight)) +
  geom_hline(yintercept = 0, lty = 2) +
  geom_smooth(method = "lm", se = F, color = "#E07A5F") +
  geom_point(size = 2, aes(shape = mode)) +
  theme_pubr(border = T)

## Additional plot: center-center dist of all TFs vs. only spacer length-dependent TFs
lib1_BSs <- read.csv2("/DATA/usr/m.trauernicht/projects/SuRE-TF/data/library_design/parameter_files/TF_motifs.csv", header = T) %>%
  dplyr::select("tf" = TF, "seq" = Motif) %>%
  filter(str_detect(tf, "neg", negate = T), str_detect(tf, "andom", negate = T)) %>%
  mutate(tf = toupper(tf))

lib1_BSs$tf[lib1_BSs$tf == "TCFCP2L1"] <- "TFCP2L1"
lib1_BSs$tf[lib1_BSs$tf == "NR3C1(GR)"] <- "NR3C1"
lib1_BSs$tf[lib1_BSs$tf == "TRP53"] <- "TP53"

lib2_BSs <- read.csv("/DATA/usr/m.trauernicht/projects/SuRE-TF/data/library_design/output/gen2_motifs.csv")  %>%
  dplyr::select("tf" = TF, "seq" = motif) 

BSs_all <- rbind(lib1_BSs, lib2_BSs) %>%
  mutate(BS_length = nchar(seq))

weight_spacing_length <- weight %>%
  filter(tf %in% chosen_conditions$tf_condition) %>%
  left_join(cDNA_df_TF5) %>%
  #filter(pval_adj < .001) %>%
  filter(features == "10bps") %>%
  mutate(tf = gsub("_.*", "", tf)) %>%
  filter(tf %in% rownames(weight2)) %>%
  mutate(spacer_sign = ifelse(tf %in% weight4$tf, "Yes", "No")) %>%
  distinct(tf, spacer_sign, weight, pval_adj, pval) %>%
  left_join(BSs_all)

ggplot() +
    geom_density(data = weight_spacing_length,
       mapping = aes(x = BS_length, fill = spacer_sign),
       alpha = .4) +
    geom_density(data = BSs_all,
       mapping = aes(x = BS_length),
       lty = 4) +
  scale_x_continuous(limits = c(5, 20))

ggplot(weight_spacing_length,
       aes(x = BS_length, y = weight, color = spacer_sign)) +
  geom_point(size = 2) +
  scale_color_manual(values = c("Yes" = "#7c9eb2", "No" = "grey80"))

ggplot(weight_spacing_length,
       aes(x = BS_length, y = -log10(pval))) +
  geom_smooth(method = "loess") +
  geom_point(aes(color = spacer_sign)) +
  scale_color_manual(values = c("Yes" = "#7c9eb2", "No" = "grey80"))

## Transform to helical turns
weight_spacing_length2 <- weight_spacing_length %>%
  mutate(BS_spacer1 = BS_length + 5) %>%
  mutate(BS_spacer2 = BS_length + 10) %>%
  mutate(spacing_rotation1 = (BS_spacer1) / 10.5) %>%
  mutate(spacing_rotation2 = (BS_spacer2) / 10.5) %>%
  mutate(spacing_degree1 = 2*pi * spacing_rotation1) %>%
  mutate(spacing_degree2 = 2*pi * spacing_rotation2) %>%
  #mutate(spacing_degree = ifelse(is.infinite(spacing_degree), 0, spacing_degree)) %>%
  mutate(spacing_degree_transf1 = cos(spacing_degree1)) %>%
  mutate(spacing_degree_transf2 = cos(spacing_degree2)) %>%
  mutate(dif_spacing = -spacing_degree_transf1 + spacing_degree_transf2) %>%
  left_join(binding_modes)

ggplot(weight_spacing_length2,
       aes(x = dif_spacing, y = weight)) +
  geom_point(aes(color = spacer_sign, shape = mode), size = 2) +
  geom_smooth(method = "lm", se = F) 


ggplot(weight_spacing_length2,
       aes(x = abs(dif_spacing), y = abs(weight))) +
  geom_point(aes(color = spacer_sign, shape = mode), size = 2) +
  geom_smooth(method = "lm", se = F) 


## Overview figures: which features matter how much and for which TF?
exp_variance_plot <- exp_variance %>%
  filter(tf %in% chosen_conditions$tf_condition) %>%
  mutate(tf = gsub("_.*", "", tf)) %>%
  mutate(text = paste("TF =", tf)) %>%
  filter(tf %in% unique(weight3$tf))

total_variance <- exp_variance_plot %>%
  filter(feature == "total") %>%
  dplyr::select(tf, "total_variance" = rel_sum_sq)

exp_variance_plot <- exp_variance_plot %>%
  filter(feature != "total") %>%
  left_join(total_variance) %>%
  #filter(total_variance > 50) %>%
  mutate(mean_rel_sum_sq = ave(rel_sum_sq, feature, FUN = mean))

ggplot(exp_variance_plot,
         aes(x = reorder(feature, -mean_rel_sum_sq), y = rel_sum_sq, fill = total_variance)) +
  geom_quasirandom(shape = 21, color = "white") +
  stat_summary(fun = median, fun.min = median, fun.max = median, geom = "crossbar", color = "red", width = 0.25, lwd = 0.4) +
  ylab("Variance explained (%)") + xlab("") +
  scale_fill_gradient(low = "grey90", high = "grey10") +
  theme_pubr(x.text.angle = 90)

exp_variance2 <- exp_variance %>%
  filter(tf %in% chosen_conditions$tf_condition) %>%
  distinct(tf, feature, rel_sum_sq) %>%
  spread(key = "feature", value = "rel_sum_sq") %>%
  column_to_rownames("tf")

pheatmap(exp_variance2 %>%
                  t(),
         cellwidth = 10,
         cellheight = 10,
         border_color = F,
         color = colorRampPalette(brewer.pal(n = 7, name = "Greys"))(100),
         angle_col = 90,
         cluster_cols = T,
         cutree_rows = 2)

# write_tsv(cDNA_df_TF3, "/DATA/usr/m.trauernicht/projects/SuRE-TF/data/mt20230302_predicted_activities.tsv")
# write_tsv(weight, "/DATA/usr/m.trauernicht/projects/SuRE-TF/data/mt20230302_lm_weights.tsv")
```


## Reporter quality score computation

Aim: We want to find for each TF the optimal reporters. Here I combine all data to compute reporter quality scores to find the best reporters for each TF.
```{r}
## Combine all data into one data table
reference_activities <- cDNA_df %>%
  filter(neg_ctrls == "No", 
         hPGK == "No", 
         str_detect(tf, "RANDOM", negate = T), 
         native_enhancer == "No") %>%
  distinct(tf, condition, reporter_id, commercial_reporter, reporter_activity_minP) %>%
  mutate(reporter_activity_minP = log2(reporter_activity_minP)) %>%
  mutate(tf = gsub("_.*", "", tf)) %>%
  mutate(tf_condition = paste(tf, condition, sep = "_")) %>%
  filter(tf_condition %in% chosen_conditions$tf_condition) %>%
  dplyr::select(-tf_condition)

cDNA_df6 <- reference_activities %>%
  mutate(reporter_activity_minP = ifelse(tf == "NR5A2", -reporter_activity_minP, reporter_activity_minP)) %>%
  left_join(perturbation_df) %>%
  left_join(rna_protein_cor) %>%
  #left_join(viper_correlations) %>%
  left_join(off_target_activities) %>%
  mutate(reporter_dif = ifelse(is.na(reporter_dif), 0, reporter_dif)) %>%
  mutate(cor_pval = base::pmax(rna_cor, prot_cor)) %>%
  #mutate(viper_cor_pval = ifelse(is.na(viper_cor_pval), 0, viper_cor_pval)) %>%
  mutate(off_target_activity = ifelse(is.na(off_target_activity), 0, off_target_activity)) %>%
  mutate(off_target_activity = -off_target_activity) %>%
  mutate(Sensitivity = reporter_dif + reporter_activity_minP + cor_pval + off_target_activity) %>%
  mutate(Specificity = reporter_dif + off_target_activity + cor_pval) %>%
  pivot_longer(cols = c("Sensitivity", "Specificity"), names_to = "quality", values_to = "quality_score") %>%
  mutate(quality_score = ifelse(quality_score < 0, 0, quality_score))

# write_tsv(cDNA_df6, "/DATA/usr/m.trauernicht/projects/SuRE-TF/data/mt20230302_reporter_scores.tsv")

lm_pval <- cDNA_df_TF4 %>%
  group_by(tf) %>%
  arrange(pval_adj, .by_group = T) %>%
  dplyr::slice(n = 1, .preserve = T) %>%
  distinct(tf, pval_adj)

## Compute condifence levels based on thresholds
cDNA_df_confidence <- cDNA_df6 %>%
  left_join(lm_pval) %>%
  mutate(conf_level = ifelse((commercial_reporter == "No" & pval_adj < 0.05 & reporter_activity_minP > 1) |
                               (commercial_reporter == "Yes" & reporter_activity_minP > 1), 1, 0)) %>% 
  ## Confidence level 1 = reporter activities can be predicted by log-linear model
  mutate(conf_level = ifelse(conf_level == 1 & cor_pval > 1, 2, conf_level)) %>% 
  ## Confidence level 2 = reporter activity follows transcript/protein levels
  mutate(conf_level = ifelse(conf_level == 1 & reporter_dif > 1, 3, conf_level)) %>% 
  ## Confidence level 3 = TF perturbation leads to change in activity
  mutate(conf_level = ifelse(conf_level >= 1 & cor_pval > 1 & reporter_dif > 1, 4, conf_level)) %>% 
  ## Confidence level 4 = All of the above together
  mutate(conf_level = ifelse(conf_level >= 2 & off_target_activity < -.5, conf_level - 1, conf_level)) ## Penalize if off-target activity

## Alternative: simple additions instead of decision tree
cDNA_df_confidence2 <- cDNA_df6 %>%
  left_join(lm_pval) %>%
  ## Add a level if reporter activities can be predicted by log-linear model
  mutate(conf_level = ifelse((commercial_reporter == "No" & pval_adj < 0.05 & reporter_activity_minP > 1) |
                               (commercial_reporter == "Yes" & reporter_activity_minP > 1), 1, 0)) %>% 
  ## Add a level if reporter activity follows transcript/protein levels
  mutate(conf_level = ifelse(cor_pval > 1, conf_level + 1, conf_level)) %>% 
  ## Add a level if TF perturbation leads to change in activity
  mutate(conf_level = ifelse(reporter_dif > 1, conf_level + 2, conf_level)) %>% 
  ## Subtract a level if substantial off-target response
  mutate(conf_level = ifelse(conf_level >= 1 & off_target_activity < -1, conf_level - 1, conf_level))

reporter_features <- cDNA_df %>%
  distinct(promoter, spacing, distance, background, tf, reporter_id)

for (i in unique(cDNA_df_confidence2$tf)){
  cDNA_df_confidence_example <- cDNA_df_confidence2 %>%
    filter(quality == "Sensitivity") %>%
    filter(tf == i) %>%
    distinct(reporter_id, cor_pval, reporter_dif, reporter_activity_minP, off_target_activity, quality_score, conf_level, pval_adj, commercial_reporter)
  
  x_order <- c("off_target_activity", "reporter_dif", "cor_pval", "reporter_activity_minP")
  y_order <- cDNA_df_confidence_example %>%
    ## sort by confidence level and then by quality score
    arrange(desc(conf_level), desc(quality_score)) %>%
    distinct(reporter_id)
  
  plot_activities <- ggplot(cDNA_df_confidence_example %>%
         dplyr::select(-conf_level, -quality_score) %>%
         pivot_longer(cols = c("reporter_activity_minP", "cor_pval", "reporter_dif", "off_target_activity", "pval_adj"), 
                      names_to = "feature", values_to = "score") %>%
         mutate(sign = ifelse((feature == "reporter_activity_minP" & unique(cDNA_df_confidence_example$pval_adj) < 0.05) & score > 1 |
                                feature == "reporter_activity_minP" & commercial_reporter == "Yes" & score > 1 |
                                (feature == "cor_pval" & score > 1) |
                                (feature == "reporter_dif" & score > 1) |
                                (feature == "off_target_activity" & score < -1),
                              "Yes", "No")) %>%
         filter(feature != "pval_adj") %>%
         mutate(feature = factor(feature, levels = x_order)) %>%
         mutate(reporter_id = factor(reporter_id, levels = y_order$reporter_id)),
       aes(x = reporter_id, y = feature, fill = score, color = sign)) +
  geom_tile(aes(size = sign)) +
  coord_fixed() +
  scale_fill_gradient2(low = "#2a9d8f", mid = "white", high = "#264653", na.value = "grey80", limits = c(-3, 6), oob = squish) +
  scale_color_manual(values = c("Yes" = "black", "No" = "white")) +
  theme_pubr(border = T, x.text.angle = 90, legend = "none") +
  scale_size_manual(values = c("Yes" = .35, "No" = 0)) +
    theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

  x_order2 <- c("promoter", "spacing", "background", "distance")
  reporter_features$distance[reporter_features$distance == "10bp"] <- "10bps"
  
  plot_features <- ggplot(cDNA_df_confidence_example %>%
         left_join(reporter_features) %>%
         mutate(background = as.character(background)) %>%
         pivot_longer(cols = c("promoter", "spacing", "background", "distance"), names_to = "feature", values_to = "feature_name") %>%
         mutate(feature = factor(feature, levels = x_order2)) %>%
         mutate(reporter_id = factor(reporter_id, levels = y_order$reporter_id)),
       aes(x = reporter_id, y = feature, fill = feature_name)) +
  geom_tile() +
  coord_fixed() +
  scale_fill_manual(values = feature_colors, na.value = "white") +
  theme_pubr(border = T, x.text.angle = 90, legend = "none") +
    theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
  
  plot_scores <- ggplot(cDNA_df_confidence_example %>%
                          distinct(reporter_id, conf_level) %>%
                          mutate(conf_level = as.character(conf_level)) %>%
                          mutate(reporter_id = factor(reporter_id, levels = y_order$reporter_id)),
                        aes(y = "Confidence level", x = reporter_id, fill = conf_level)) +
    geom_tile() +
    coord_fixed() +
    scale_fill_manual(values = c("0" = "#CCCCCB", "1" = "#EFCABE", "2" = "#E7A08B", "3" = "#DD6840", "4" = "#BF4A27")) +
  theme_pubr(border = T, x.text.angle = 90, legend = "none") +
    theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
  
  combined_plot <- ggarrange(plot_scores, NULL, plot_activities, NULL, plot_features,
                             ncol = 1, nrow = 5, heights = c(1,-.5,1,-.5,1))
  
  combined_plot_ann <- annotate_figure(combined_plot, text_grob(i))
  
  print(combined_plot_ann)
  
}

ggplot(cDNA_df_confidence2 %>%
         distinct(tf, reporter_id, reporter_dif, commercial_reporter) %>%
         mutate(n_reporters = as.numeric(ave(reporter_id, tf, FUN = length))) %>%
         filter(n_reporters > 10) %>%
         mutate(dif_mean = ave(reporter_dif, tf, FUN = mean)),
       aes(x = reorder(tf, -dif_mean), y = reporter_dif, shape = commercial_reporter)) +
  geom_hline(yintercept = 1, lty = 2) +
  geom_quasirandom(aes(color = ifelse(reporter_dif > 1, "Yes", "No"))) +
  scale_color_manual(values = c("Yes" = "black", "No" = "grey")) +
  theme_pubr(x.text.angle = 90)


ggplot(cDNA_df_confidence2 %>%
         distinct(tf, reporter_id, conf_level) %>%
         mutate(conf_level_sum = ave(conf_level, tf, FUN = sum)) %>% 
         mutate(n_reporters = as.numeric(ave(reporter_id, tf, FUN = length))) %>%
         filter(n_reporters > 10) %>%
         mutate(average_level = conf_level_sum / n_reporters) %>%
         mutate(conf_level = factor(conf_level, levels = c("4", "3", "2", "1", "0"))) %>% 
         na.omit(),
       aes(x = reorder(tf, average_level), fill = conf_level)) +
  geom_bar(stat = "count", position = "stack") +
  theme_pubr() + 
  scale_fill_manual(values = c("0" = "#CCCCCB", "1" = "#EFCABE", "2" = "#E7A08B", "3" = "#DD6840", "4" = "#BF4A27")) +
  coord_flip()

ggplot(cDNA_df_confidence2 %>%
         filter(quality == "Sensitivity") %>%
         distinct(tf, reporter_id, conf_level, quality_score, commercial_reporter) %>%
         mutate(conf_level_sum = ave(conf_level, tf, FUN = sum)) %>% 
         mutate(n_reporters = as.numeric(ave(reporter_id, tf, FUN = length))) %>%
         filter(n_reporters > 10) %>%
         mutate(average_level = conf_level_sum / n_reporters) %>%
         mutate(conf_level = factor(conf_level, levels = c("4", "3", "2", "1", "0"))) %>% 
         na.omit(),
       aes(x = reorder(tf, -average_level), y = quality_score, color = conf_level, shape = commercial_reporter)) +
  geom_quasirandom() +
  theme_pubr(x.text.angle = 90) + 
  scale_color_manual(values = c("0" = "#CCCCCB", "1" = "#EFCABE", "2" = "#E7A08B", "3" = "#DD6840", "4" = "#BF4A27"))
```


## Fit linear model to predict which features explain reporter ranking per TF
```{r}
## Fit linear model to predict which features explain reporter ranking per TF
cDNA_df_confidence_lm <- cDNA_df_confidence2 %>%
  filter(quality == "Sensitivity", commercial_reporter == "No") %>%
  distinct(tf, reporter_id, conf_level, quality_score) %>%
  group_by(tf) %>%
  arrange(desc(conf_level), desc(quality_score)) %>%
  mutate(reporter_rank = 1:n()) %>%
  left_join(reporter_features) %>%
  mutate(promoter = factor(promoter)) %>%
  mutate(background = factor(background)) %>%
  mutate(spacing = gsub("10bp", "10bps", spacing)) %>%
  mutate(spacing = factor(spacing)) %>%
  arrange(reporter_id)
  
## Fit linear model with design features for actual data
exp_variance <- data.frame("feature" = c("promoter", "spacing", "distance", "background", "Residuals"))
weight <- data.frame("feature" = c("(Intercept)", "promoterminP", "promoterhBGm", "promotermCMV",
                  "spacing10bps","spacing5bp", "distance10bp","distance21bp", "background1","background2", "background3"))

cDNA_df_confidence_lm$promoter <- relevel(cDNA_df_confidence_lm$promoter, ref = "minP")
cDNA_df_confidence_lm$spacing <- relevel(cDNA_df_confidence_lm$spacing, ref = "5bp")
cDNA_df_confidence_lm$row <- rownames(cDNA_df_confidence_lm)
cDNA_df_confidence_lm3 <- data.frame()

lm_list <- list()
for (i in unique(cDNA_df_confidence_lm$tf)) {
  if (nrow(cDNA_df_confidence_lm[cDNA_df_confidence_lm$tf == i,]) > 9) {
  print(i)
  lm_list[[i]] <- lm(reporter_rank ~
                       distance + promoter + spacing + background,
                cDNA_df_confidence_lm[cDNA_df_confidence_lm$tf == i,])
  }
}

f <- lapply(names(lm_list), function(x) summary(lm_list[[x]])$fstatistic)
names(f) <- names(lm_list)
p <- lapply(names(f), function(x) pf(f[[x]][1],f[[x]][2],f[[x]][3],lower.tail=F))
p <- data.frame(bind_rows(p)) %>% setnames("value", "pval")
p$tf <- names(lm_list)

y <- lapply(names(lm_list), function(x) data.frame("rank_predicted" = lm_list[[x]]$fitted.values) %>% mutate(tf = x))
y <- data.frame(bind_rows(y)) %>% mutate(row = as.character(1:n())) %>% left_join(p, by = "tf")

cDNA_df_confidence_lm2 <- y %>%
  left_join(cDNA_df_confidence_lm)

cDNA_df_confidence_lm3 <- rbind(cDNA_df_confidence_lm3, cDNA_df_confidence_lm2)

mean(cDNA_df_confidence_lm3$pval)

w <- lapply(names(lm_list), function(x) 
  if (length(lm_list[[x]]$coefficients) == length(summary(lm_list[[x]])$coefficients[,4])) {
  data.frame('weight' = lm_list[[x]]$coefficients, 'pval' = summary(lm_list[[x]])$coefficients[,4]) %>% mutate(tf = x) %>% rownames_to_column('feature')
  }
)

w <- data.frame(bind_rows(w))
weight <- weight %>%
  left_join(w)

x <- lapply(names(lm_list), function(x) anova(lm_list[[x]]) %>% rownames_to_column("feature") %>% mutate(tf = x))
x <- data.frame(bind_rows(x)) %>% 
  setnames(c("Sum.Sq", "Pr..F."), c("sum_sq", "pval")) %>%
  distinct(feature, sum_sq, pval, tf) %>%
  mutate(sum = ave(sum_sq, tf, FUN = sum)) %>%
  mutate(rel_sum_sq = (sum_sq/sum)*100) %>%
  distinct(tf, rel_sum_sq, pval, feature)

exp_variance <- exp_variance %>%
  left_join(x)

## Fine-tune dfs
exp_variance$group <- "no residual"
exp_variance$group[grep("Residual", exp_variance$feature)] <- "residual"
exp_variance$rel_sum_sq[exp_variance$feature == "Residuals"] <- 100 - exp_variance$rel_sum_sq[exp_variance$feature == "Residuals"]
exp_variance$feature <- revalue(exp_variance$feature, c("Residuals"="total"))
prediction <- cDNA_df_confidence_lm3 %>%
  dplyr::select(-row) %>%
  na.omit()
weight <- weight %>%
  filter(feature != "(Intercept)")
categorie <- c("promoter", "spacing", "distance", "background")
weight$features <- gsub(paste(categorie, collapse="|"), "",weight$feature)
ml <- length(unique(weight$features))
feature <- unique(weight$features)
weight$condition <- gsub(paste(feature, collapse = "|"), "", weight$feature)
weight <- weight %>% dplyr::select(-feature)


cDNA_df_confidence_lm3 <- cDNA_df_confidence_lm3 %>%
  mutate(pval_adj = p.adjust(pval, method = "fdr"))

## Example: Predicted vs. measured for one TF
ggplot(cDNA_df_confidence_lm3 %>%
         filter(tf == "STAT3"),
       aes(x = rank_predicted, y = reporter_rank, color = background)) +
  geom_abline(lty = 2) +
  geom_point(size = 2) +
  scale_color_manual(values = c("black", "grey50", "grey80")) +
  theme_pubr()

weight_ex <- weight %>%
  filter(tf == "NR1H4::RXRA")

zero_weights <- weight %>%
  filter(is.na(tf)) %>%
  mutate(weight = 0, pval = 1) %>%
  mutate(tf = "NR1H4::RXRA")
                   
weight_ex <- rbind(weight_ex, zero_weights)
                   
feature_names <- data.frame("features" = c("minP", "mCMV", "hBGm", "1", "2", "3", "5bp", "10bps", "10bp", "21bp"),
                            "feature_names" = c("minP", "minCMV", "minHBG", "Spacer 1", "Spacer 2", "Spacer 3", "5bp Spacer", 
                                                "10bp Spacer", "10bp Distance", "21bp Distance"))
                     
weight_ex <- weight_ex %>%
  left_join(feature_names)
                   

ggplot(weight_ex %>% 
         mutate(sign = ifelse(pval < 0.05, "Yes", "No")) %>%
         mutate(feature_names = factor(feature_names, levels = c("minP", "minCMV", "minHBG", "Spacer 1", "Spacer 2", "Spacer 3", "5bp Spacer", 
                                                                                                        "10bp Spacer", "10bp Distance", "21bp Distance"))),
       aes(x = feature_names, y = -weight, alpha = sign)) +
  geom_hline(yintercept = 0, lty = 2) +
  geom_bar_interactive(stat = "identity", color = "black") +
  ylab("Weight") + 
  xlab("Reporter design feature") +
  theme_classic_lines_45() +
  scale_alpha_discrete(range = c(0.3,1))

```


## Figure 2: Plot activities in chosen conditions

Aim: Get a good overview of reporter activities
```{r}
cell_colors <- c("A549" = "#f2cc8f", "HCT116" = "#babf95", "HEK293" = "#f4f1de", 
                               "HepG2" = "#5f797b", "K562" = "#8f5d5d", "MCF7" = "#81b29a", 
                               "U2OS" = "#3d405b", "mES" = "#eab69f", "NPC" = "#e07a5f") %>%
  as.data.frame() %>%
  rownames_to_column("cell")

## Quasirandom plots of reporter activities -- combined plot with best cell line per TF
cDNA_df_plot <- cDNA_df6 %>%
  distinct(tf, reporter_id, reporter_activity_minP, condition) %>%
  mutate(cell = gsub("_.*", "", condition)) %>%
  mutate(mean_tf_activity = ave(reporter_activity_minP, tf, FUN = mean)) %>%
  mutate(cell = factor(cell, levels = c("A549", "HCT116", "HEK293", "HepG2", "K562",
                                        "MCF7", "U2OS", "mES", "NPC"))) %>%
  mutate(n_tf = as.numeric(ave(tf, tf, FUN = length))) %>%
  filter(n_tf > 10)

tfs_colors <- cDNA_df_plot %>%
  distinct(tf, cell, mean_tf_activity) %>%
  arrange(desc(mean_tf_activity)) %>%
  left_join(cell_colors)
  
  
tfs_colors_vec <- tfs_colors %>%
  column_to_rownames("tf") %>%
  dplyr::select(-cell, -mean_tf_activity) %>%
  t() %>%
  base::as.vector()

names(tfs_colors_vec) <- tfs_colors$tf


ggplot(cDNA_df_plot,
       aes(x = reorder(tf, -mean_tf_activity), y = reporter_activity_minP)) +
  geom_hline(yintercept = 1, lty = 3) + 
  geom_hline(yintercept = -1, lty = 3) + 
  geom_quasirandom(width = .2, aes(color = ifelse(abs(reporter_activity_minP) > 1, "black", "grey70")),
                   alpha = .6) +
  stat_summary(fun = median, fun.min = median, fun.max = median, geom = "crossbar", color = "red", width = 0.5, lwd = 0.4) +
  theme_pubr(x.text.angle = 90) +
  scale_color_manual(values = c("black", "grey70")) +
  theme(axis.text.x = element_text(colour=tfs_colors_vec))

## Same but just show two cell lines
cDNA_df_plot <- cDNA_df %>%  
  filter(neg_ctrls == "No",
         commercial_reporter == "No",
         hPGK == "No", 
         str_detect(tf, "RANDOM", negate = T), 
         native_enhancer == "No") %>%
  distinct(tf, reporter_id, reporter_activity_minP, condition) %>%
  mutate(tf = gsub("_.*", "", tf)) %>%
  mutate(cell = gsub("_.*", "", condition)) %>%
  mutate(reporter_activity_minP = log2(ave(reporter_activity_minP, reporter_id, cell, FUN = mean))) %>%
  distinct(tf, reporter_id, reporter_activity_minP, cell) %>%
  mutate(n_tf = as.numeric(ave(tf, tf, cell, FUN = length))) %>%
  filter(n_tf >= 20) %>%
  filter(cell %in% c("NPC", "mES"))

cDNA_df_plot2 <- cDNA_df_plot %>%
  filter(cell == "NPC") %>%
  mutate(mean_tf_activity = ave(reporter_activity_minP, tf, cell, FUN = mean)) %>%
  distinct(tf, mean_tf_activity)

cDNA_df_plot <- cDNA_df_plot %>%
  left_join(cDNA_df_plot2)

ggplot(cDNA_df_plot %>%
           mutate(cell = factor(cell, levels = c("NPC", "mES"))),
       aes(x = reorder(tf, -mean_tf_activity), y = reporter_activity_minP)) +
    geom_hline(yintercept = 1, lty = 3) + 
    geom_hline(yintercept = -1, lty = 3) + 
    geom_quasirandom_rast(width = .2, shape = 21, stroke = NA, aes(fill = ifelse(abs(reporter_activity_minP) > 1, "black", "grey70")),
                          alpha = .6, size = 2, raster.dpi = 600) +
    stat_summary(fun = median, fun.min = median, fun.max = median, geom = "crossbar", color = "red", width = 0.5, lwd = 0.4) +
    theme_classic_lines_90() +
    scale_fill_manual(values = c("black", "grey70")) +
    facet_wrap(~cell, nrow = 2)

ggplot(cDNA_df_plot %>%
         mutate(mean_tf_activity_cell = ave(reporter_activity_minP, tf, cell, FUN = mean)) %>%
           mutate(cell = factor(cell, levels = c("NPC", "mES"))) %>%
         distinct(mean_tf_activity, tf, cell, mean_tf_activity_cell),
       aes(x = reorder(tf, -mean_tf_activity), y = mean_tf_activity_cell)) +
    geom_hline(yintercept = 1, lty = 3) + 
    geom_hline(yintercept = -1, lty = 3) + 
    geom_bar(stat = "identity", aes(fill = ifelse(abs(mean_tf_activity_cell) > 1, "black", "grey70"))) +
    theme_classic_lines_90() +
    scale_fill_manual(values = c("black", "grey70")) +
    facet_wrap(~cell, nrow = 2)


## Investigate neg ctrls better
cDNA_df_ctrl <- cDNA_df %>%  
  filter(hPGK == "No", 
         str_detect(tf, "RANDOM", negate = T), 
         native_enhancer == "No") %>%
  distinct(tf, reporter_id, reporter_activity_minP, condition, neg_ctrls) %>%
  mutate(tf = gsub("_.*", "", tf)) %>%
  mutate(cell = gsub("_.*", "", condition)) %>%
  mutate(reporter_activity_minP = log2(ave(reporter_activity_minP, reporter_id, cell, FUN = mean))) %>%
  distinct(tf, reporter_id, reporter_activity_minP, cell, neg_ctrls) %>%
  mutate(n_tf = as.numeric(ave(tf, tf, cell, FUN = length))) %>%
  filter(n_tf >= 20) %>%
  filter(cell %in% c("mES")) %>%
  mutate(reporter_id = gsub("_neg", "", reporter_id)) %>%
  pivot_wider(names_from = neg_ctrls, values_from = reporter_activity_minP)

ggplot(cDNA_df_ctrl,
       aes(x = Yes, y = No)) +
  geom_abline() +
  geom_point(aes(color = ifelse(Yes > 2 | (Yes - No) > 1, "Yes", "No"))) +
  scale_color_manual(values = c("Yes"= "red", "No" = "black")) +
  facet_wrap(~tf)

cDNA_df_ctrl2 <- cDNA_df %>%  
  filter(condition %in% c("HepG2_CEBPB", "HepG2_NT"),
         hPGK == "No", 
         str_detect(tf, "RANDOM", negate = T), 
         native_enhancer == "No") %>%
  distinct(tf, reporter_id, reporter_activity_minP, condition, neg_ctrls) %>%
  mutate(tf = gsub("_.*", "", tf)) %>%
  filter(tf == "CEBPB") %>%
  mutate(reporter_activity_minP = log2(ave(reporter_activity_minP, reporter_id, condition, FUN = mean))) %>%
  distinct(tf, reporter_id, reporter_activity_minP, condition, neg_ctrls) %>%
  mutate(n_tf = as.numeric(ave(tf, tf, condition, FUN = length))) %>%
  filter(n_tf >= 20) %>%
  mutate(reporter_id = gsub("_neg", "", reporter_id)) %>%
  pivot_wider(names_from = neg_ctrls, values_from = reporter_activity_minP) %>%
  mutate(dif = No - Yes) %>%
  dplyr::select(-Yes, -No) %>%
  distinct() %>%
  pivot_wider(names_from = condition, values_from = dif)

ggplot(cDNA_df_ctrl2,
       aes(x = Yes, y = No)) +
  geom_abline() +
  geom_point(aes(color = ifelse(Yes > 2 | (Yes - No) > 1, "Yes", "No"))) +
  scale_color_manual(values = c("Yes"= "red", "No" = "black")) +
  facet_wrap(~condition)
```

## Figure 4: Showcase how reporter activities change in different conditions
```{r}
## Showcase example
ggplot(cDNA_df %>% 
         filter(str_detect(tf, "VDR"), condition %in% c("U2OS", "U2OS_Calcitriol"), neg_ctrls == "No", promoter != "Random") %>%
         distinct(reporter_id, commercial_reporter, condition, reporter_activity_minP, promoter) %>%
         spread(key = "condition", value = "reporter_activity_minP"),
       aes(x = log2(U2OS),
           y = log2(U2OS_Calcitriol), 
           shape = commercial_reporter, 
           color = promoter)) +
  geom_hline(yintercept = 0, lty = 2) + 
  geom_vline(xintercept = 0, lty = 2) +
  geom_abline(lty = "twodash") +
  geom_point(size = 2.5) +
  xlim(-0.5, 8) +
  ylim(-0.5, 8) +
  xlab("Activity (log2)") +
  ylab(paste("Perturbation effect (log2)")) +
  theme_pubr(border = T) +
  scale_color_manual(values = c("mCMV" = "black", "hBGm" = "grey50", "minP" = "grey80"))

ggplot(cDNA_df %>% 
         filter(str_detect(tf, "NFE2L2"), condition %in% c("HepG2_NT", "HepG2_NFE2L2"), neg_ctrls == "No") %>%
         distinct(reporter_id, commercial_reporter, spacing, condition, reporter_activity_minP) %>%
         spread(key = "condition", value = "reporter_activity_minP"),
       aes(x = log2(HepG2_NT),
           y = log2(HepG2_NFE2L2), 
           shape = commercial_reporter,
           color = spacing)) +
  geom_hline(yintercept = 0, lty = 2) +
  geom_vline(xintercept = 0, lty = 2) +
  geom_abline(lty = "twodash") +
  geom_point(size = 2.5) +
  #xlim(-0.5, 3.5) +
  #ylim(-0.5, 3.5) +
  xlab("Activity (log2)") +
  ylab(paste("Perturbation effect (log2)")) +
  theme_pubr(border = T) +
  scale_color_manual(values = c("#7C9EB2", "#BFD0D9"))

ggplot(cDNA_df %>% 
         filter(str_detect(tf, "GRHL1"), condition %in% c("mES_NT", "mES_TFCP2L1"), neg_ctrls == "No") %>%
         distinct(reporter_id, commercial_reporter, spacing, condition, reporter_activity_minP) %>%
         spread(key = "condition", value = "reporter_activity_minP"),
       aes(x = log2(mES_NT),
           y = log2(mES_TFCP2L1), 
           shape = commercial_reporter,
           color = spacing)) +
  geom_abline(lty = 2) +
  geom_point(size = 2.5) +
  #xlim(-0.5, 3.5) +
  #ylim(-0.5, 3.5) +
  xlab("Activity (log2)") +
  ylab(paste("Perturbation effect (log2)")) +
  theme_pubr(border = T) +
  scale_color_manual(values = c("#7C9EB2", "#BFD0D9"))

ggplot(cDNA_df %>% 
         filter(str_detect(tf, "FOXA1"), condition %in% c("mES_Dox", "mES_FOXA1-OE"), neg_ctrls == "No") %>%
         distinct(reporter_id, commercial_reporter, spacing, condition, reporter_activity_minP) %>%
         spread(key = "condition", value = "reporter_activity_minP"),
       aes(x = log2(mES_Dox),
           y = log2(`mES_FOXA1-OE`), 
           shape = commercial_reporter,
           color = spacing)) +
  geom_abline(lty = 2) +
  geom_point(size = 2.5) +
  #xlim(-0.5, 3.5) +
  #ylim(-0.5, 3.5) +
  xlab("Activity (log2)") +
  ylab(paste("Perturbation effect (log2)")) +
  theme_pubr(border = T) +
  scale_color_manual(values = c("#7C9EB2", "#BFD0D9"))


ggplot(cDNA_df %>% 
         filter(str_detect(tf, "TP53"), condition %in% c("HepG2", "HepG2_TP73"), neg_ctrls == "No", promoter != "Random") %>%
         distinct(reporter_id, commercial_reporter, spacing, condition, promoter, reporter_activity_minP) %>%
         spread(key = "condition", value = "reporter_activity_minP"),
       aes(x = log2(HepG2),
           y = log2(HepG2_TP73), 
           shape = commercial_reporter)) +
  geom_hline(yintercept = 0, lty = 2) +
  geom_vline(xintercept = 0, lty = 2) +
  geom_abline(lty = "twodash") +
  geom_point(size = 2.5, aes(color = promoter)) +
  xlab("Activity (log2)") +
  ylab(paste("Off-target effect (log2)")) +
  theme_pubr(border = T) +
  scale_color_manual(values = c("mCMV" = "black", "hBGm" = "grey50", "minP" = "grey80"))

ggplot(cDNA_df %>% 
         filter(str_detect(tf, "STAT3"), condition %in% c("mES_2i_LIF", "mES_2i"), neg_ctrls == "No") %>%
         distinct(reporter_id, commercial_reporter, spacing, condition, background, reporter_activity_minP) %>%
         spread(key = "condition", value = "reporter_activity_minP"),
       aes(x = log2(mES_2i),
           y = log2(mES_2i_LIF), 
           shape = commercial_reporter,
           color = as.factor(spacing))) +
  geom_hline(yintercept = 0, lty = 2) +
  geom_vline(xintercept = 0, lty = 2) +
  geom_abline(lty = "twodash") +
  geom_point(size = 2.5) +
  xlab("Activity (log2)") +
  ylab(paste("Perturbation effect (log2)")) +
  theme_pubr(border = T) +
  scale_color_manual(values = c("#7C9EB2", "#BFD0D9")) +
  facet_wrap(~background)
```


---


## Rate performance of synthetic vs. published reporters
```{r, fig.width=9, fig.height=9}
## First I want to show how well commercial reporters perform compared to synthetic ones
bad_tfs <- cDNA_df_confidence2 %>%
  distinct(tf, conf_level) %>%
  filter(!is.na(conf_level)) %>%
  group_by(tf) %>%
  mutate(removed = conf_level > 0) %>%
  mutate(removed2 = TRUE %in% removed) %>%
  filter(removed2 == TRUE) %>%
  distinct(tf)

commercial_ranking <- cDNA_df_confidence2 %>%
  filter(quality == "Sensitivity", tf %in% bad_tfs$tf) %>%
  group_by(tf) %>%
  arrange(conf_level, quality_score) %>%
  mutate(rank = row_number()) %>%
  distinct(tf, reporter_id, rank, commercial_reporter, conf_level) %>%
  ## remove tfs for which I don't have commercial reporters
  group_by(tf) %>%
  mutate(commercial = "Yes" %in% commercial_reporter) %>%
  ungroup() %>%
  filter(commercial == TRUE) %>%
  ## calculate relative rank
  mutate(max_rank = ave(rank, tf, FUN = max)) %>%
  filter(max_rank > 10) %>%
  mutate(rel_rank = rank/max_rank)

ggplot(commercial_ranking %>%
         filter(commercial_reporter == "Yes") %>%
         mutate(max_rank2 = ave(rel_rank, tf, FUN = max)),
       aes(x = reorder(tf, -max_rank2), y = rel_rank)) +
  geom_point(aes(color = as.factor(conf_level))) +
  theme_pubr(x.text.angle = 90) +
  scale_color_manual(values = c("0" = "#CCCCCB", "1" = "#EFCABE", "2" = "#E7A08B", "3" = "#DD6840", "4" = "#BF4A27")) 

## Plot overview figure of commercial vs. synthetic reporters (only best reporter per TF)
commercial_ranking <- cDNA_df_confidence2 %>%
  filter(quality == "Sensitivity") %>%
  group_by(tf) %>%
  arrange(conf_level, quality_score) %>%
  mutate(rank = row_number()) %>%
  distinct(tf, reporter_id, rank, commercial_reporter, conf_level) %>%
  ## remove tfs for which I don't have commercial reporters
  group_by(tf) %>%
  mutate(commercial = "Yes" %in% commercial_reporter) %>%
  ungroup() %>%
  #filter(commercial == TRUE) %>%
  ## calculate relative rank
  mutate(max_rank = ave(rank, tf, FUN = max)) %>%
  filter(max_rank > 10) %>%
  mutate(rel_rank = rank/max_rank)

commercial_ranking_best <- commercial_ranking %>%
  mutate(max_rank2 = ave(rank, tf, commercial_reporter, FUN = max)) %>%
  distinct(tf, max_rank2, conf_level, commercial_reporter, rank, commercial) %>%
  filter(rank == max_rank2)

conf_level_synth <- commercial_ranking_best %>%
  filter(commercial_reporter == "No") %>%
  mutate(max_conf_level = ave(conf_level, tf, FUN = max)) %>%
  distinct(tf, max_conf_level)

commercial_ranking_best <- commercial_ranking_best %>%
  left_join(conf_level_synth)

better_synth <- commercial_ranking_best %>%
  filter(commercial == T) %>%
  distinct(max_rank2, tf, commercial_reporter) %>%
  pivot_wider(values_from = "max_rank2", names_from = "commercial_reporter") %>%
  mutate(better_synth = ifelse(No > Yes, "0_Yes", "1_No")) %>%
  distinct(tf, better_synth)

commercial_ranking_best <- commercial_ranking_best %>%
  left_join(better_synth)

commercial_ranking_best_comm <- commercial_ranking_best %>%
  filter(commercial == T) %>%
  mutate(conf_level_sum = ave(conf_level, tf, FUN = sum)) %>%
  arrange(max_conf_level, better_synth, conf_level_sum) %>%
  mutate(tf = factor(tf, levels = unique(tf))) %>%
  mutate(commercial_reporter = factor(commercial_reporter, levels = c("No", "Yes")))

### TFs where comparison synthtic vs. published is possible
ggplot(commercial_ranking_best_comm,
       aes(y = tf, x = commercial_reporter, fill = as.factor(conf_level))) +
  geom_tile() +
  coord_fixed() +
  theme_pubr(x.text.angle = 90) +
  scale_fill_manual(values = c("0" = "#CCCCCB", "1" = "#EFCABE", "2" = "#E7A08B", "3" = "#DD6840", "4" = "#BF4A27")) +
  scale_size_manual(values = c("Yes" = .35, "No" = 0))

### TFs with only synthetic
ggplot(commercial_ranking_best %>%
         filter(commercial == F),
       aes(y = reorder(tf, conf_level), x = "x", fill = as.factor(conf_level))) +
  geom_tile() +
  coord_fixed() +
  theme_pubr(x.text.angle = 90) +
  scale_fill_manual(values = c("0" = "#CCCCCB", "1" = "#EFCABE", "2" = "#E7A08B", "3" = "#DD6840", "4" = "#BF4A27")) 

### TFs with only published
commercial_ranking2 <- cDNA_df_confidence2 %>%
  filter(quality == "Sensitivity") %>%
  group_by(tf) %>%
  arrange(conf_level, quality_score) %>%
  mutate(rank = row_number()) %>%
  distinct(tf, reporter_id, rank, commercial_reporter, conf_level) %>%
  ## remove tfs for which I don't have commercial reporters
  group_by(tf) %>%
  mutate(commercial = "Yes" %in% commercial_reporter) %>%
  ungroup() %>%
  #filter(commercial == TRUE) %>%
  ## calculate relative rank
  mutate(max_rank = ave(rank, tf, FUN = max)) %>%
  filter(max_rank < 12) %>%
  mutate(rel_rank = rank/max_rank)

commercial_ranking_best_only_comm <- commercial_ranking2 %>%
  mutate(max_rank2 = ave(rank, tf, commercial_reporter, FUN = max)) %>%
  distinct(tf, max_rank2, conf_level, commercial_reporter, rank, commercial) %>%
  filter(rank == max_rank2)

ggplot(commercial_ranking_best_only_comm,
       aes(y = reorder(tf, conf_level), x = "x", fill = as.factor(conf_level))) +
  geom_tile() +
  coord_fixed() +
  theme_pubr(x.text.angle = 90) +
  scale_fill_manual(values = c("0" = "#CCCCCB", "1" = "#EFCABE", "2" = "#E7A08B", "3" = "#DD6840", "4" = "#BF4A27")) 

```

---

### Explore activities of best reporters in all cell types
```{r}
## Select set of best reporters
optimal_reporters <- cDNA_df_confidence2 %>%
  group_by(tf) %>%
  arrange(desc(conf_level), desc(quality_score)) %>%
  mutate(rank = row_number()) %>%
  filter(rank <= 3 & conf_level >= 3 & tf != "RFX1")


### Activity across the 9 cell types
heatmap_cell_activities <- cDNA_df %>%
  filter(reporter_id %in% optimal_reporters$reporter_id) %>%
  dplyr::filter(neg_ctrls == "No", 
                stimulation == "no",
                hPGK == "No", 
                str_detect(tf, "RANDOM", negate = T), 
                native_enhancer == "No") %>%
  distinct(tf, reporter_activity_minP, cell) %>%
  mutate(tf = gsub("_.*", "", tf)) %>%
  mutate(reporter_activity_minP = log2(ave(reporter_activity_minP, tf, cell, FUN = mean))) %>%
  distinct() %>%
  spread(key = "tf", value = "reporter_activity_minP") %>% 
  replace(is.na(.), 0) %>%
  column_to_rownames("cell")

myBreaks <- c(seq(-1, 0, length.out=50+1), 
              seq(0.000001, 5, length.out=50))

pheatmap(heatmap_cell_activities %>% 
           as.matrix(),
         color = colorRampPalette(c("#F2CC8F", "white", "#3D405B"))(100),
         clustering_method = "ward.D2",
         border_color = F, 
         cellwidth = 8, 
         cellheight = 8,
         angle_col = 90,
         cutree_cols = 10,
         #cutree_rows = 3,
         breaks = myBreaks)
  
rna_prot_cor2 <- rna_protein_cor %>%
  filter(reporter_id %in% optimal_reporters$reporter_id) %>%
  mutate(max_cor = pmax(prot_cor, rna_cor)) %>%
  mutate(tf = gsub("(.*)_.*_.*_.*_.*", "\\1", reporter_id)) %>%
  mutate(tf = gsub("_.*", "", tf)) %>%
  distinct(tf, max_cor) %>%
  mutate(max_cor_mean = ave(max_cor, tf, FUN = mean)) %>%
  distinct(tf, max_cor_mean)


hc <- hclust(dist(heatmap_cell_activities %>% t()), method = "ward.D2")
order_dendogram <- data.frame("order" = 1:ncol(heatmap_cell_activities),
                              "tf" = "")

for(i in unique(order_dendogram$order)) {
  order_dendogram$tf[order_dendogram$order == i] <- names(heatmap_cell_activities[hc$order[i]])
}

ggplot(rna_prot_cor2 %>%
         mutate(cor_bin = ifelse(max_cor_mean > 1, 1, 0)) %>%
         left_join(order_dendogram),
       aes(x = reorder(tf, order), y = "", fill = max_cor_mean)) +
  geom_tile() +
  coord_fixed() +
  scale_fill_gradient(low = "white", high = "#2a9d8f", na.value = "grey80", limits = c(0, 3), oob = squish) +
  theme_pubr(border = T, x.text.angle = 90, legend = "none") 


### UMAP analysis

```

---

### Explore activities of optimal reporters upon KDs  
```{r}
### Activities upon KD
cDNA_df2 <- cDNA_df %>%
  filter(reporter_id %in% optimal_reporters$reporter_id) %>%
  dplyr::filter(neg_ctrls == "No", 
                #commercial_reporter == "No", 
                hPGK == "No", 
                str_detect(tf, "RANDOM", negate = T), 
                native_enhancer == "No") %>%
  distinct(tf, condition, reporter_activity_minP, stimulation, reference_condition, cell) %>%
  mutate(tf = gsub("_.*", "", tf)) %>%
  mutate(reporter_activity_minP = log2(ave(reporter_activity_minP, tf, condition, FUN = mean))) %>%
  distinct()

reference_condition <- cDNA_df2 %>%
  filter(condition %in% reference_condition) %>%
  dplyr::select(-reference_condition, "reference_condition" = condition, -stimulation, "reference_activity" = reporter_activity_minP) #%>%
  #filter(abs(reference_activity) > 1)

cDNA_df3 <- cDNA_df2 %>%
  filter(cell %in% c("HepG2", "mES")) %>%
  filter(stimulation == "KD") %>%
  left_join(reference_condition) %>%
  mutate(reporter_FC = reporter_activity_minP - reference_activity) %>%
  mutate(TF_dif = ave(reporter_FC, tf, condition, FUN = mean)) %>%
  mutate(TF_dif = ifelse(abs(TF_dif) > 0.2, TF_dif, 0))

cDNA_df3[is.na(cDNA_df3)] <- 0

cDNA_df4 <- cDNA_df3 %>%
  distinct(tf, condition, TF_dif, cell) %>%
  mutate(condition = gsub("HepG2_", "", condition)) %>%
  mutate(condition = gsub("mES_", "", condition)) %>%
  filter(condition %in% tf) %>%
  filter(tf %in% condition) %>%
  mutate(diag = ifelse(tf == condition, T, F)) 

KD_doubles <- cDNA_df4 %>%
  filter(diag == T) %>%
  mutate(n_kd = ave(tf, tf, FUN = length)) %>%
  filter(n_kd == 2) %>%
  group_by(tf) %>%
  dplyr::arrange(TF_dif, .by_group = T) %>%
  dplyr::slice(n = 1, .preserve = T) %>%
  ungroup() %>%
  mutate(condition2 = paste(condition, cell, sep = "_"))

KD_singles <- cDNA_df4 %>%
  filter(diag == T) %>%
  mutate(n_kd = ave(tf, tf, FUN = length)) %>%
  filter(n_kd == 1) %>%
  mutate(condition2 = paste(condition, cell, sep = "_"))

cDNA_df5 <- cDNA_df4 %>%
  mutate(condition2 = paste(condition, cell, sep = "_")) %>%
  filter(condition2 %in% KD_doubles$condition2 | condition2 %in% KD_singles$condition2) %>%
  filter(!condition %in% c("TP53"), !tf %in% c("TP53")) %>%
  arrange(diag)
 
ggplot(cDNA_df5,
       aes(x = tf, y = condition, fill = TF_dif)) +
  geom_tile(color = cDNA_df5$diag, size = .5) +
  scale_color_manual(values = c("black", "white")) +
  scale_fill_gradient2(low = "#99B2DD",
                        mid = "white",
                        high = "#dd6b48",
                       limits = c(-2,2),
                       oob = squish) +
  coord_fixed() +
  xlab("TF reporter") +
  ylab("KD condition") +
  theme_pubr(border = T, x.text.angle = 90) 

## All of the reporters should be sensitive to perturbation - so if the KD doesn't show any signal, the KD did not work
### I will therefore now remove all KDs where I don't see any direct changes
cDNA_df5_eff_conditions <- cDNA_df5 %>%
  filter(diag == T & abs(TF_dif) > 1)

cDNA_df5_eff <- cDNA_df5 %>%
  filter(condition %in% cDNA_df5_eff_conditions$condition) %>%
  filter(tf %in% cDNA_df5_eff_conditions$condition)

ggplot(cDNA_df5_eff,
       aes(x = tf, y = condition, fill = TF_dif)) +
  geom_tile(color = cDNA_df5_eff$diag, size = .5) +
  scale_color_manual(values = c("black", "white")) +
  scale_fill_gradient2(low = "#99B2DD",
                        mid = "white",
                        high = "#dd6b48",
                       limits = c(-2,2),
                       oob = squish) +
  coord_fixed() +
  xlab("TF reporter") +
  ylab("KD condition") +
  theme_pubr(border = T, x.text.angle = 90) 


### Interaction analysis
threshold <- 1

interaction_df <- cDNA_df5_eff %>%
  dplyr::select("receiver" = tf, "signal" = condition, TF_dif) %>%
  distinct() %>%
  mutate(TF_dif_binary = "0") %>%
  mutate(TF_dif_binary = ifelse(TF_dif > threshold, "-1", TF_dif_binary)) %>%
  mutate(TF_dif_binary = ifelse(TF_dif < -threshold, "1", TF_dif_binary)) %>%
  mutate(TF_dif_binary = ifelse(TF_dif > -threshold & TF_dif < threshold, "0", TF_dif_binary)) %>%
  mutate(TF_dif_binary = as.numeric(TF_dif_binary)) %>%
  mutate(TF_dif_binary = (ifelse(receiver == signal, 0, TF_dif_binary))) %>%
  mutate(sum_receiver = ave(TF_dif_binary, receiver, FUN = function(x) sum(x))) %>%
  mutate(sum_signal = ave(TF_dif_binary, signal, FUN = function(x) sum(x))) %>%
  mutate(sum_receiver_size = ave(TF_dif_binary, receiver, FUN = function(x) sum(abs(x)))) %>%
  mutate(sum_signal_size = ave(TF_dif_binary, signal, FUN = function(x) sum(abs(x))))

interaction_matrix <- interaction_df %>%
  distinct(TF_dif_binary, receiver, signal) %>%
  spread(key = "receiver", value = "TF_dif_binary") %>%
  column_to_rownames("signal")

interaction_nodes <- interaction_df %>%
  filter(signal == receiver) %>%
  distinct(receiver, signal, sum_signal, sum_receiver, sum_signal_size, sum_receiver_size) %>%
  mutate(sum_total = sum_receiver_size + sum_signal_size) %>%
  mutate(class = ifelse(sum_signal <= 0, "repressor", "activator")) %>%
  distinct(receiver, sum_total, class)

interaction_edges <- interaction_df %>%
  left_join(target_tf %>% dplyr::select("signal" = condition, target)) %>%
  filter(signal != receiver) %>%
  distinct(receiver, signal, TF_dif_binary, TF_dif) %>%
  mutate(weight = 0) %>%
  mutate(weight = ifelse(TF_dif_binary == -1, 1, weight)) %>%
  mutate(weight = ifelse(TF_dif_binary == 1, 2, weight)) %>%
  mutate(class = ifelse(weight == 1, "repressor", "activator")) %>%
  dplyr::select(signal, receiver, weight, class, TF_dif) %>%
  filter(weight != 0) %>%
  mutate(TF_dif = abs(TF_dif))

interaction_nodes <- interaction_nodes %>%
  filter(receiver %in% c(unique(interaction_edges$receiver),unique(interaction_edges$signal)))

routes_igraph <- graph_from_data_frame(d = interaction_edges, vertices = interaction_nodes, directed = TRUE)

visIgraph(routes_igraph) %>%
    visNodes(size = 25, shape = "circle") %>%
    visOptions(highlightNearest = list(enabled = TRUE, hover = TRUE),
               nodesIdSelection = TRUE) %>%
    visInteraction(keyboard = TRUE)


ggraph(routes_igraph, layout = "sugiyama") +
  geom_node_point(aes(size = sum_total, color = class)) +
  geom_edge_link(aes(color = class, width = TF_dif),
                 alpha = 0.8,
                 arrow = arrow(length = unit(3, 'mm')),
                 start_cap = circle(2, 'mm'),
                 end_cap = circle(3, 'mm')) +
  scale_edge_width(range = c(0.2, 2)) +
  scale_radius(limits = c(1, NA), range = c(1, 6)) +
  geom_node_text(aes(label = name), repel = TRUE) +
  scale_edge_color_manual(values = c("activator" = colors_diverse[2], "repressor" = colors_diverse[1])) +
  scale_color_manual(values = c("activator" = colors_diverse[2], "repressor" = colors_diverse[1])) +
  theme_graph()

```


---

### Explore overexpression data
``` {r}
### Activities upon overexpression
cDNA_df2 <- cDNA_df %>%
  #filter(reporter_id %in% optimal_reporters$reporter_id) %>%
  dplyr::filter(neg_ctrls == "No", 
                #commercial_reporter == "No", 
                hPGK == "No", 
                str_detect(tf, "RANDOM", negate = T), 
                native_enhancer == "No") %>%
  distinct(tf, condition, reporter_activity_neg, stimulation, reference_condition, cell) %>%
  mutate(tf = gsub("_.*", "", tf)) %>%
  mutate(reporter_activity_neg = log2(ave(reporter_activity_neg, tf, condition, FUN = mean))) %>%
  distinct()

reference_condition <- cDNA_df2 %>%
  filter(condition %in% reference_condition) %>%
  dplyr::select(-reference_condition, "reference_condition" = condition, -stimulation, "reference_activity" = reporter_activity_neg) #%>%
  #filter(abs(reference_activity) > 1)

cDNA_df3 <- cDNA_df2 %>%
  filter(stimulation == "overexpression") %>%
  left_join(reference_condition) %>%
  mutate(reporter_FC = reporter_activity_neg - reference_activity) %>%
  mutate(TF_dif = ave(reporter_FC, tf, condition, FUN = mean)) %>%
  mutate(TF_dif = ifelse(abs(TF_dif) > 0.2, TF_dif, 0))

cDNA_df4 <- cDNA_df3 %>%
  distinct(tf, condition, TF_dif) %>%
  mutate(condition = gsub("mES_", "", condition)) %>%
  mutate(condition = gsub("-OE", "", condition)) %>%
  #filter(condition %in% tf) %>%
  #filter(tf %in% condition) %>%
  mutate(diag = ifelse(tf == condition, T, F)) %>%
  arrange(diag)

ggplot(cDNA_df4,
       aes(x = tf, y = condition, fill = TF_dif)) +
  geom_tile(color = cDNA_df4$diag, size = .5) +
  scale_color_manual(values = c("black", "white")) +
  scale_fill_gradient2(low = "#99B2DD",
                        mid = "white",
                        high = "#dd6b48",
                       limits = c(-2,2),
                       oob = squish) +
  coord_fixed() +
  xlab("TF reporter") +
  ylab("KD condition") +
  theme_pubr(border = T, x.text.angle = 90) 
```

---

### Explore most effective TF perturbations
```{r}
### TF activity changes upon stimulation 
cDNA_df2 <- cDNA_df %>%
  filter(reporter_id %in% optimal_reporters$reporter_id) %>%
  dplyr::filter(neg_ctrls == "No", 
                #commercial_reporter == "No", 
                hPGK == "No", 
                str_detect(tf, "RANDOM", negate = T), 
                native_enhancer == "No") %>%
  distinct(tf, condition, reporter_activity_neg, stimulation, reference_condition, cell) %>%
  mutate(tf = gsub("_.*", "", tf)) %>%
  mutate(reporter_activity_neg = log2(ave(reporter_activity_neg, tf, condition, FUN = mean))) %>%
  distinct()

reference_condition <- cDNA_df2 %>%
  filter(condition %in% reference_condition) %>%
  dplyr::select(-reference_condition, "reference_condition" = condition, -stimulation, "reference_activity" = reporter_activity_neg) #%>%
  #filter(abs(reference_activity) > 1)

stimulations <- stimulations %>%
  distinct(condition, tf) %>%
  mutate(target = "Yes")

cDNA_df3 <- cDNA_df2 %>%
  filter((stimulation == "pathway" & ## Define most effective pathway stimulations
         condition %in% c("HCT116_Calcymicin", "HepG2_CDCA", "K562_Hemin", "K562_T3", "MCF7_Hexestrol", "mES_2i", "mES_BX", 
                          "mES_HQ", "mES_IWP2", "mES_LIF_CH", "mES_LIF_PD", "mES_N2B27", "mES_serum_2i_LIF", 
                          "mES_vitC", "U2OS_Calcitriol", "U2OS_Heat", "U2OS_PMA", "U2OS_Wortmannin", "mES_FK",
                          "HepG2_Rif")) |
           (stimulation == "overexpression" & 
              condition %in% c("mES_FOSL1-OE", "mES_FOXA1-OE", "mES_GLI1-OE", "mES_NR4A2-OE-cDIM12")) |
           (stimulation == "degron") |
           (stimulation == "KD" &
           condition %in% c("mES_PPARG::RXRA", "mES_PPARA::RXRA",
                            "mES_POU5F1", "mES_SP1", "mES_ATF2", "mES_TFCP2L1",
                            "HepG2_CEBPG", "HepG2_ATF2", "HepG2_SRF", 
                            "HepG2_STAT3",  "HepG2_ETS2", "HepG2_CEBPB", 
                            "HepG2_MTF1", "HepG2_NRF1",  "HepG2_SP1",  "HepG2_ATF4",
                             "HepG2_NFKB1", "HepG2_STAT1", "HepG2_MEF2A",
                            "HepG2_IRF2", "HepG2_IRX3", "HepG2_ATF6", "HepG2_NFYA", "HepG2_MAF::NFE2", 
                             "HepG2_PAX6", "HepG2_NFE2L2"
                            #"HepG2_RARA", "HepG2_FOXA1", "HepG2_GRHL1", "HepG2_TCF7", "HepG2_REL", "HepG2_RXRA", "mES_HIC2", 
                            #"mES_EGR1","mES_NR1H2","HepG2_NR1H2","HepG2_TFCP2","HepG2_ETV4", "HepG2_ETS1", "HepG2_TP73",
                            #"HepG2_HNF1A","HepG2_THRB"
                            ))) %>%
  left_join(stimulations) %>%
  left_join(reference_condition) %>%
  mutate(reporter_FC = reporter_activity_neg - reference_activity) %>%
  mutate(TF_dif = ave(reporter_FC, tf, condition, FUN = mean))

cDNA_df3[is.na(cDNA_df3)] <- 0

cDNA_df3 <- cDNA_df3 %>%
  distinct(tf, condition, TF_dif) %>%
  spread(tf, TF_dif) %>%
  column_to_rownames("condition")

myBreaks <- c(seq(-1.5, 0, length.out=50+1), 
              seq(0.000001, 1.5, length.out=50))

pheatmap(cDNA_df3 %>% 
                 t(),
         color = colorRampPalette(c("#99B2DD", "white", "#E07A5F"))(100),
         clustering_method = "ward.D2",
         border_color = F, 
         cellwidth = 10, 
         cellheight = 10,
         #annotation_col = tf_clusters, 
         angle_col = 90,
         cutree_cols = 10,
         cutree_rows = 18,
         breaks = myBreaks)

## Make another overview focusing on only high-confidence conditions


```

---

## Focus on a set of mES conditions
```{r}
cDNA_df2 <- cDNA_df %>%
  filter(reporter_id %in% optimal_reporters$reporter_id) %>%
  dplyr::filter(neg_ctrls == "No", 
                #commercial_reporter == "No", 
                hPGK == "No", 
                str_detect(tf, "RANDOM", negate = T), 
                native_enhancer == "No") %>%
  distinct(tf, condition, reporter_activity_minP, stimulation, reference_condition, cell) %>%
  mutate(tf = gsub("_.*", "", tf)) %>%
  mutate(reporter_activity_minP = log2(ave(reporter_activity_minP, tf, condition, FUN = mean))) %>%
  distinct()

# All mES perturbations: KDs, overexpressions, stimulations, degron
cDNA_df2_mES <- cDNA_df2 %>%
  filter(str_detect(reference_condition, "mES") | is.na(reference_condition))

reference_condition <- cDNA_df2_mES %>%
  filter(condition %in% reference_condition) %>%
  dplyr::select(-reference_condition, "reference_condition" = condition, -stimulation, "reference_activity" = reporter_activity_minP)

cDNA_df2_mES2_filt <- cDNA_df2_mES %>%
  left_join(reference_condition) %>%
  mutate(reporter_dif = reporter_activity_minP - reference_activity) %>%
  filter(condition %in% stimulations_effective$condition | stimulation %in% c("KD", "degron", "overexpression")) %>%
  dplyr::select(-reference_condition, -stimulation, -cell, -reporter_activity_minP, -reference_activity) %>%
  mutate(cell = "mES") %>%
  left_join(tf_rna) %>%
  filter(TPM_norm > 5 | tf %in% c("EGR1", "PAX6", "ESR1", "FOSL1", "FOXA1", "GATA1", "NFIA", "PPARG::RXRA", "RUNX2", "WT1")) %>%
  dplyr::select(-TPM, - TPM_norm, -cell)

### Remove conditions that don't seem to induce any changes
cDNA_df2_mES2_filt2 <- cDNA_df2_mES2_filt %>%
  mutate(reporter_dif_max = ave(reporter_dif, condition, FUN = function(x) max(x, na.rm = T))) %>%
  mutate(reporter_dif_min = ave(reporter_dif, condition, FUN = function(x) min(x, na.rm = T))) %>%
  filter(reporter_dif_max > .75 | reporter_dif_min < -.75) %>%
  filter(!condition %in% c("mES_KLF9", "mES_GRHL1", "mES_HSF1", "mES_CREB1", "mES_RXRA", "mES_KLF4", "mES_SOX9",
                            "mES_TEAD2", "mES_POU2F1", "mES_MEF2B", "mES_TCF7", "mES_SOX2", "mES_RFX1-OE", "mES_GATA1-OE", 
                            "mES_OTX2-OE", "mES_EGR1", "mES_MEF2A",
                           "mES_NFIA", "mES_SOX13", "mES_E2F1", "mES_HSF2", "mES_TCF7L1", "mES_TEAD1"))

cDNA_df2_mES2_heatmap <- cDNA_df2_mES2_filt %>%
  filter(condition %in% cDNA_df2_mES2_filt2$condition | condition %in% c("mES_MEF2A")) %>%
  filter(!tf %in% c("AR")) %>%
  #mutate(reporter_dif = ifelse(abs(reporter_dif) > 0.5, reporter_dif, 0)) %>%
  pivot_wider(names_from = "condition", values_from = "reporter_dif") %>%
  replace(is.na(.), 0) %>%
  column_to_rownames("tf") 

myBreaks <- c(seq(-1.5, 0, length.out=50+1), 
              seq(0.000001, 1.5, length.out=50))

pheatmap(cDNA_df2_mES2_heatmap,
         color = colorRampPalette(c("#99B2DD", "white", "#E07A5F"))(100),
         clustering_method = "ward.D",
         border_color = F, 
         cellwidth = 10, 
         cellheight = 10,
         #annotation_col = tf_clusters, 
         angle_col = 90,
         cutree_cols = 8,
         cutree_rows = 9,
         breaks = myBreaks)

target_tf <- data.frame("condition" = c("mES_2i", "mES_ATF2", "mES_BX", "mES_FK", "mES_FOSL1-OE", "mES_FOXA1-OE", 
                                        "mES_GRHL1", "mES_HQ", "mES_KLF4", "mES_LIF_CH", "mES_LIF_PD", "mES_MEF2A", 
                                        "mES_NFIA", "mES_NR1H2", "mES_NR4A2-OE-cDIM12", "mES_Pou5f1_AID", "mES_PPARA::RXRA", 
                                        "mES_PPARG::RXRA", "mES_Sox2_AID", "mES_SP1", "mES_TEAD1", "mES_TFCP2L1", "mES_vitC"),
                        "target" = c("STAT3", "ATF2", "IRF3", "CREB1", "FOS::JUN", "FOXA1", "GRHL1", "NFE2L2",
                                     "KLF4", "EGR1", "TCF7", "MEF2A", "NFIA", "NR1H2", "NR4A2", "POU5F1::SOX2", 
                                     "PPARA::RXRA", "PPARG::RXRA", "SOX2", "SP1", "TEAD1", "TFCP2L1", "NFE2L2"),
                        "effect_size" = c(0, 0, 0, 1, 1, 1, 0, 1,  0, 1, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0))

target_tf_all <- data.frame("condition" = c("mES_2i", "mES_ATF2", "mES_BX", "mES_FK", "mES_FOSL1-OE", "mES_FOXA1-OE", 
                                        "mES_GRHL1", "mES_HQ", "mES_KLF4", "mES_LIF_CH", "mES_LIF_PD", "mES_MEF2A", 
                                        "mES_NFIA", "mES_NR1H2", "mES_NR4A2-OE-cDIM12", "mES_Pou5f1_AID", "mES_PPARA::RXRA", 
                                        "mES_PPARG::RXRA", "mES_Sox2_AID", "mES_SP1", "mES_TEAD1", "mES_TFCP2L1",
                                        "mES_GLI1-OE", "mES_HIC2", "mES_Nanog_AID", "mES_vitC"),
                        "target" = c("STAT3", "ATF2", "IRF3", "CREB1", "FOS::JUN", "FOXA1", "GRHL1", "NFE2L2",
                                     "KLF4", "EGR1", "TCF7", "MEF2A", "NFIA", "NR1H2", "NR4A2", "POU5F1::SOX2", 
                                     "PPARA::RXRA", "PPARG::RXRA", "SOX2", "SP1", "TEAD1", "TFCP2L1",
                                     "GLI1", "HIC2", "NANOG", "NFE2L2"),
                        "effect_size" = c(0, 0, 0, 1, 1, 1, 0, 1,  0, 1, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0))

cDNA_df4 <- cDNA_df2_mES2_filt %>%
  left_join(target_tf) %>%
  distinct(tf, condition, reporter_dif, target) %>%
  filter(target %in% tf) %>%
  filter(tf %in% target) %>%
  mutate(diag = ifelse(tf == target, T, F)) %>%
  arrange(diag)

ggplot(cDNA_df4,
       aes(x = tf, y = condition, fill = reporter_dif)) +
  geom_tile(color = cDNA_df4$diag, size = .5) +
  scale_color_manual(values = c("black", "white")) +
  scale_fill_gradient2(low = "#99B2DD",
                        mid = "white",
                        high = "#dd6b48",
                       limits = c(-2,2),
                       oob = squish) +
  coord_fixed() +
  xlab("TF reporter") +
  ylab("perturbation condition") +
  ggtitle("Only TFs that are direct target of KD") +
  theme_pubr(border = T, x.text.angle = 90) +
  labs(fill="FC compared to control - mean of all reporters")

## UMAP visualization
set.seed(124) 
tf_umap <- umap(cDNA_df2_mES2_heatmap) 
colnames(tf_umap$layout) <- c("A","B") 

### Extract clustering data
hc <- hclust(dist(cDNA_df2_mES2_heatmap), method = "complete")
cluster_tfs <- data.frame("cluster" = cutree(hc, k = 9)) %>%
  rownames_to_column("tf")
tf_umap_plot <- tf_umap$layout %>% as.data.frame() %>% rownames_to_column(var = "tf") %>% left_join(cluster_tfs) #Import data from previous line

#Plot with umap with 4-node cliques
ggplot(tf_umap_plot,
       aes(A,B)) +
  stat_density2d(geom = "polygon", aes(alpha = ..level..), bins = 8) +
  theme_pubr() + 
  geom_point(size = 4, aes(color = as.factor(cluster))) + 
  theme(panel.grid = element_blank()) + 
  scale_alpha(range = c(0.15,0.3)) +
  geom_text_repel(aes(label = tf), force = 40) + 
  ylab("UMAP2") + 
  xlab("UMAP1")

## UMAP of all conditions?
reference_condition <- cDNA_df2 %>%
  filter(condition %in% reference_condition) %>%
  dplyr::select(-reference_condition, "reference_condition" = condition, -stimulation, "reference_activity" = reporter_activity_minP)

cDNA_df_heatmap_all <- cDNA_df2 %>%
  filter(cell == "HepG2") %>%
  left_join(reference_condition) %>%
  mutate(reporter_dif = reporter_activity_minP - reference_activity) %>%
  filter(condition %in% stimulations_effective$condition | stimulation %in% c("KD", "degron", "overexpression")) %>%
  distinct(tf, condition, reporter_dif) %>%
  pivot_wider(names_from = "condition", values_from = "reporter_dif") %>%
  replace(is.na(.), 0) %>%
  column_to_rownames("tf") 

pheatmap(cDNA_df_heatmap_all,
         color = colorRampPalette(c("#99B2DD", "white", "#dd6b48"))(100),
         clustering_method = "ward.D",
         border_color = F, 
         cellwidth = 10, 
         cellheight = 10,
         #annotation_col = tf_clusters, 
         angle_col = 90,
         cutree_cols = 8,
         cutree_rows = 9,
         breaks = myBreaks)

set.seed(353) 
tf_umap <- umap(cDNA_df_heatmap_all) 
colnames(tf_umap$layout) <- c("A","B") 

### Extract clustering data
hc <- hclust(dist(cDNA_df_heatmap_all), method = "complete")
cluster_tfs <- data.frame("cluster" = cutree(hc, k = 9)) %>%
  rownames_to_column("tf")
tf_umap_plot <- tf_umap$layout %>% as.data.frame() %>% rownames_to_column(var = "tf") %>% left_join(cluster_tfs) #Import data from previous line

#Plot with umap with 4-node cliques
ggplot(tf_umap_plot,
       aes(A,B)) +
  stat_density2d(geom = "polygon", aes(alpha = ..level..), bins = 8) +
  theme_pubr() + 
  geom_point(size = 4, aes(color = as.factor(cluster))) + 
  theme(panel.grid = element_blank()) + 
  scale_alpha(range = c(0.15,0.3)) +
  geom_text_repel(aes(label = tf), force = 40) + 
  ylab("UMAP2") + 
  xlab("UMAP1")

## Heatmap - UMAP coordinates per cell type - or cluster per cell type


```

---

## Interaction network between TFs
```{r}
threshold <- 1

cDNA_df4 <- cDNA_df2_mES2_filt %>%
  filter(condition %in% cDNA_df2_mES2_filt2$condition | condition %in% c("mES_MEF2A")) %>%
  filter(!tf %in% c("AR")) %>%
  left_join(target_tf_all) %>%
  na.omit() %>%
  distinct(tf, condition, reporter_dif, target)# %>%
  # filter(target %in% tf) %>%
  # filter(tf %in% target) %>%
  # mutate(diag = ifelse(tf == target, T, F)) %>%
  # arrange(diag)

interaction_df <- cDNA_df4 %>%
  dplyr::select("receiver" = tf, "signal" = target, reporter_dif) %>%
  distinct() %>%
  mutate(TF_dif_binary = "0") %>%
  mutate(TF_dif_binary = ifelse(reporter_dif > threshold, "-1", TF_dif_binary)) %>%
  mutate(TF_dif_binary = ifelse(reporter_dif < -threshold, "1", TF_dif_binary)) %>%
  mutate(TF_dif_binary = ifelse(reporter_dif > -threshold & reporter_dif < threshold, "0", TF_dif_binary)) %>%
  mutate(TF_dif_binary = as.numeric(TF_dif_binary)) %>%
  mutate(TF_dif_binary = (ifelse(receiver == signal, 0, TF_dif_binary))) %>%
  mutate(sum_receiver = ave(TF_dif_binary, receiver, FUN = function(x) sum(x))) %>%
  mutate(sum_signal = ave(TF_dif_binary, signal, FUN = function(x) sum(x))) %>%
  mutate(sum_receiver_size = ave(TF_dif_binary, receiver, FUN = function(x) sum(abs(x)))) %>%
  mutate(sum_signal_size = ave(TF_dif_binary, signal, FUN = function(x) sum(abs(x))))


interaction_matrix <- interaction_df %>%
  distinct(TF_dif_binary, receiver, signal) %>%
  spread(key = "receiver", value = "TF_dif_binary") %>%
  column_to_rownames("signal")

pheatmap(interaction_matrix,
         border_color = NA,
         color = colorRampPalette(c("#d5bdaf", 'white', "#84a98c"))(100),
         clustering_method = "ward.D")


###
interaction_nodes1 <- interaction_df %>%
  filter(signal == receiver) %>%
  distinct(receiver, signal, sum_signal, sum_receiver, sum_signal_size, sum_receiver_size) %>%
  mutate(sum_total = sum_receiver_size + sum_signal_size) %>%
  mutate(class = ifelse(sum_signal <= 0, "repressor", "activator")) %>%
  distinct(receiver, sum_total, class)

### Add only-signal
interaction_nodes2 <- interaction_df %>%
  filter(!signal %in% unique(interaction_nodes1$receiver)) %>%
  dplyr::select("receiver" = signal, "sum_total" = sum_signal_size, sum_signal) %>%
  mutate(class = ifelse(sum_signal <= 0, "repressor", "activator")) %>%
  dplyr::select(-sum_signal) %>%
  distinct()

### Add only-receiver
interaction_nodes3 <- interaction_df %>%
  filter(!receiver %in% unique(interaction_nodes1$receiver)) %>%
  dplyr::select(receiver, "sum_total" = sum_receiver_size, sum_receiver) %>%
  mutate(class = ifelse(sum_receiver <= 0, "repressor", "activator")) %>%
  dplyr::select(-sum_receiver) %>%
  distinct()

interaction_nodes <- rbind(interaction_nodes1, interaction_nodes2, interaction_nodes3)

interaction_edges <- interaction_df %>%
  left_join(target_tf %>% dplyr::select("signal" = condition, target)) %>%
  filter(signal != receiver) %>%
  distinct(receiver, signal, TF_dif_binary, reporter_dif) %>%
  mutate(weight = 0) %>%
  mutate(weight = ifelse(TF_dif_binary == -1, 1, weight)) %>%
  mutate(weight = ifelse(TF_dif_binary == 1, 2, weight)) %>%
  mutate(class = ifelse(weight == 1, "repressor", "activator")) %>%
  dplyr::select(signal, receiver, weight, class, reporter_dif) %>%
  filter(weight != 0) %>%
  mutate(reporter_dif = abs(reporter_dif)) #%>%
  #mutate(signal = gsub("POU5F1::SOX2", "POU5F1", signal)) %>%
  #mutate(receiver = gsub("POU5F1::SOX2", "POU5F1", receiver))

interaction_edges <- interaction_edges #%>%
  #filter(!(signal == "POU5F1::SOX2" & receiver == "SOX2")) %>% ## Remove interactions caused by off-target
  #filter(!(signal == "SOX2" & receiver == "POU5F1::SOX2")) 

interaction_nodes <- interaction_nodes %>%
  #mutate(receiver = gsub("POU5F1::SOX2", "POU5F1", receiver)) %>%
  filter(receiver %in% c(unique(interaction_edges$receiver),unique(interaction_edges$signal)))

interaction_nodes <- interaction_nodes %>%
  mutate(color = ifelse(class == "repressor", "grey90", "#9AC1AE"))

interaction_edges <- interaction_edges %>%
  mutate(color = ifelse(class == "repressor", "grey90", "#9AC1AE"))

routes_igraph <- graph_from_data_frame(d = interaction_edges, vertices = interaction_nodes, directed = TRUE)

visIgraph(routes_igraph) %>%
    visNodes(size = 25, shape = "circle") %>%
    visOptions(highlightNearest = list(enabled = TRUE, hover = TRUE),
               nodesIdSelection = TRUE) %>%
    visInteraction(keyboard = TRUE)

ggraph(routes_igraph, layout = "sugiyama") +
  geom_node_point(aes(size = sum_total, color = class)) +
  geom_edge_link(aes(color = class, width = reporter_dif),
                 alpha = 0.8,
                 arrow = arrow(length = unit(3, 'mm')),
                 start_cap = circle(2, 'mm'),
                 end_cap = circle(3, 'mm')) +
  scale_edge_width(range = c(0.2, 2)) +
  scale_radius(limits = c(1, NA), range = c(1, 6)) +
  geom_node_text(aes(label = name), repel = TRUE) +
  scale_edge_color_manual(values = c("activator" = colors_diverse[2], "repressor" = colors_diverse[1])) +
  scale_color_manual(values = c("activator" = colors_diverse[2], "repressor" = colors_diverse[1])) +
  theme_graph()
```


## For individual reporter variants
```{r}
cDNA_df2 <- cDNA_df %>%
  dplyr::filter(neg_ctrls == "No", 
                #commercial_reporter == "No", 
                hPGK == "No", 
                str_detect(tf, "RANDOM", negate = T), 
                native_enhancer == "No") %>%
  distinct(tf, condition, reporter_activity_minP, reporter_id, stimulation, reference_condition, cell) %>%
  mutate(tf = gsub("_.*", "", tf)) %>%
  mutate(reporter_activity_minP = log2(reporter_activity_minP)) %>%
  distinct()

# All mES perturbations: KDs, overexpressions, stimulations, degron
cDNA_df2_mES <- cDNA_df2 %>%
  filter(str_detect(reference_condition, "mES") | is.na(reference_condition))

reference_condition <- cDNA_df2_mES %>%
  filter(condition %in% reference_condition) %>%
  dplyr::select(-reference_condition, "reference_condition" = condition, -stimulation, "reference_activity" = reporter_activity_minP)

cDNA_df2_mES2 <- cDNA_df2_mES %>%
  left_join(reference_condition) %>%
  mutate(reporter_dif = reporter_activity_minP - reference_activity) %>%
  filter(condition %in% stimulations_effective$condition | stimulation %in% c("KD", "degron", "overexpression")) %>%
  dplyr::select(-reference_condition, -stimulation, -cell, -reporter_activity_minP, -reference_activity) %>%
  mutate(cell = "mES") %>%
  left_join(tf_rna) %>%
  filter(TPM_norm > 5) %>%
  dplyr::select(-TPM, - TPM_norm, -cell)

### Remove conditions that don't seem to induce any changes
cDNA_df2_mES2_filt <- cDNA_df2_mES2 %>%
  mutate(reporter_dif_max = ave(reporter_dif, condition, FUN = function(x) max(x, na.rm = T))) %>%
  mutate(reporter_dif_min = ave(reporter_dif, condition, FUN = function(x) min(x, na.rm = T))) %>%
  filter(reporter_dif_max > .75 | reporter_dif_min < -.75) %>%
  filter(!condition %in% c("mES_KLF9", "mES_GRHL1", "mES_HSF1", "mES_CREB1", "mES_RXRA", "mES_KLF4", "mES_SOX9",
                            "mES_TEAD2", "mES_POU2F1", "mES_MEF2B", "mES_TCF7", "mES_SOX2", "mES_RFX1-OE", "mES_GATA1-OE", 
                            "mES_OTX2-OE",
                           "mES_NFIA", "mES_SOX13", "mES_E2F1", "mES_HSF2", "mES_TCF7L1", "mES_TEAD1"))

cDNA_df2_mES2_heatmap <- cDNA_df2_mES2 %>%
  filter(condition %in% cDNA_df2_mES2_filt$condition | condition %in% c("mES_MEF2A")) %>%
  filter(!tf %in% c("AR")) %>%
  #mutate(reporter_dif = ifelse(abs(reporter_dif) > 0.5, reporter_dif, 0)) %>%
  pivot_wider(names_from = "condition", values_from = "reporter_dif") %>%
  replace(is.na(.), 0) %>%
  column_to_rownames("reporter_id") 

myBreaks <- c(seq(-1.5, 0, length.out=50+1), 
              seq(0.000001, 1.5, length.out=50))

for (i in unique(cDNA_df2_mES2_heatmap$tf)) {
  if (nrow(cDNA_df2_mES2_heatmap[cDNA_df2_mES2_heatmap$tf == i,]) > 10) {
      p <- pheatmap(cDNA_df2_mES2_heatmap %>% 
             filter(tf == i) %>%
             dplyr::select(-tf),
         color = colorRampPalette(c("#99B2DD", "white", "#E07A5F"))(100),
         clustering_method = "ward.D",
         border_color = F, 
         cellwidth = 10, 
         cellheight = 10,
         #annotation_col = tf_clusters, 
         angle_col = 90,
         cutree_cols = 8,
         cutree_rows = 9,
         breaks = myBreaks)
      print(p)
  }
}


target_tf <- data.frame("condition" = c("mES_2i", "mES_ATF2", "mES_BX", "mES_FK", "mES_FOSL1_OE", "mES_FOXA1_OE", 
                                        "mES_GRHL1", "mES_HQ", "mES_KLF4", "mES_LIF_CH", "mES_LIF_PD", "mES_MEF2A", 
                                        "mES_NFIA", "mES_NR1H2", "mES_NR4A2-OE-cDIM12", "mES_Pou5f1_AID", "mES_PPARA::RXRA", 
                                        "mES_PPARG::RXRA", "mES_Sox2_AID", "mES_SP1", "mES_TEAD1", "mES_TFCP2L1"),
                        "target" = c("STAT3", "ATF2", "IRF3", "CREB1", "FOS::JUN", "FOXA1", "GRHL1", "NFE2L2",
                                     "KLF4", "EGR1", "TCF7", "MEF2A", "NFIA", "NR1H2", "NR4A2", "POU5F1::SOX2", 
                                     "PPARA::RXRA", "PPARG::RXRA", "SOX2", "SP1", "TEAD1", "TFCP2L1"),
                        "effect_size" = c(0, 0, 0, 1, 1, 1, 0, 1,  0, 1, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0))

cDNA_df4 <- cDNA_df2_mES2 %>%
  left_join(target_tf) %>%
  distinct(tf, condition, reporter_dif, target) %>%
  filter(target %in% tf) %>%
  filter(tf %in% target) %>%
  mutate(diag = ifelse(tf == target, T, F)) %>%
  arrange(diag)

ggplot(cDNA_df4,
       aes(x = tf, y = condition, fill = reporter_dif)) +
  geom_tile(color = cDNA_df4$diag, size = .5) +
  scale_color_manual(values = c("black", "white")) +
  scale_fill_gradient2(low = "#99B2DD",
                        mid = "white",
                        high = "#dd6b48",
                       limits = c(-2,2),
                       oob = squish) +
  coord_fixed() +
  xlab("TF reporter") +
  ylab("perturbation condition") +
  ggtitle("Only TFs that are direct target of KD") +
  theme_pubr(border = T, x.text.angle = 90) +
  labs(fill="FC compared to control - mean of all reporters")
```

## Focus on HepG2 KD-conditions only
```{r}
cDNA_df2 <- cDNA_df %>%
  filter(reporter_id %in% optimal_reporters$reporter_id) %>%
  dplyr::filter(neg_ctrls == "No", 
                #commercial_reporter == "No", 
                hPGK == "No", 
                str_detect(tf, "RANDOM", negate = T), 
                native_enhancer == "No") %>%
  distinct(tf, condition, reporter_activity_minP, stimulation, reference_condition, cell) %>%
  mutate(tf = gsub("_.*", "", tf)) %>%
  mutate(reporter_activity_minP = log2(ave(reporter_activity_minP, tf, condition, FUN = mean))) %>%
  distinct()

reference_condition <- cDNA_df2 %>%
  filter(condition %in% reference_condition) %>%
  dplyr::select(-reference_condition, "reference_condition" = condition, -stimulation, "reference_activity" = reporter_activity_minP) #%>%
  #filter(abs(reference_activity) > 1)

cDNA_df3 <- cDNA_df2 %>%
  filter(cell == "HepG2") %>%
  filter(stimulation == "KD") %>%
  left_join(reference_condition) %>%
  mutate(reporter_FC = reporter_activity_minP - reference_activity) %>%
  mutate(TF_dif = ave(reporter_FC, tf, condition, FUN = mean)) %>%
  #mutate(TF_dif = ifelse(abs(TF_dif) > 0.2, TF_dif, 0)) %>%
  mutate(mean_activity = ave(reporter_activity_minP, tf, condition, FUN = mean)) %>%
  left_join(tf_rna) %>%
  filter(TPM_norm > 5) %>%
  filter(!condition %in% c("HepG2_TP53", "HepG2_IRF3", "HepG2_TCF3", "HepG2_GATA4")) %>%
  mutate(condition = gsub("HepG2_", "", condition))

cDNA_df3[is.na(cDNA_df3)] <- 0

cDNA_df4 <- cDNA_df3 %>%
  distinct(tf, condition, TF_dif) %>%
  spread(condition, TF_dif) %>%
  column_to_rownames("tf") %>%
  replace(is.na(.), 0) %>%
  select_if(colSums(.) != 0)

myBreaks <- c(seq(-1.5, 0, length.out=50+1), 
              seq(0.000001, 1.5, length.out=50))

pheatmap(cDNA_df4 %>% 
                 as.matrix(),
         color = colorRampPalette(c("#99B2DD", "white", "#E07A5F"))(100),
         clustering_method = "ward.D",
         border_color = F, 
         cellwidth = 8, 
         cellheight = 8,
         #annotation_col = tf_clusters, 
         angle_col = 90,
         cutree_cols = 5,
         cutree_rows = 6,
         breaks = myBreaks)

cDNA_df4 <- cDNA_df3 %>%
  distinct(tf, condition, TF_dif) %>%
  filter(condition %in% tf) %>%
  filter(tf %in% condition) %>%
  mutate(diag = ifelse(tf == condition, T, F)) %>%
  arrange(diag)

ggplot(cDNA_df4,
       aes(x = tf, y = condition, fill = TF_dif)) +
  geom_tile(color = cDNA_df4$diag, size = .5) +
  scale_color_manual(values = c("black", "white")) +
  scale_fill_gradient2(low = "#99B2DD",
                        mid = "white",
                        high = "#dd6b48",
                       limits = c(-2,2),
                       oob = squish) +
  coord_fixed() +
  xlab("TF reporter") +
  ylab("KD condition") +
  ggtitle("Only TFs that are direct target of KD") +
  theme_pubr(border = T, x.text.angle = 90) +
  labs(fill="FC compared to control - mean of all reporters")
```


---


## Interaction network between TFs
```{r}
threshold <- 1

interaction_df <- cDNA_df4 %>%
  dplyr::select("receiver" = tf, "signal" = condition, TF_dif) %>%
  distinct() %>%
  mutate(TF_dif_binary = "0") %>%
  mutate(TF_dif_binary = ifelse(TF_dif > threshold, "-1", TF_dif_binary)) %>%
  mutate(TF_dif_binary = ifelse(TF_dif < -threshold, "1", TF_dif_binary)) %>%
  mutate(TF_dif_binary = ifelse(TF_dif > -threshold & TF_dif < threshold, "0", TF_dif_binary)) %>%
  mutate(TF_dif_binary = as.numeric(TF_dif_binary)) %>%
  mutate(TF_dif_binary = (ifelse(receiver == signal, 0, TF_dif_binary))) %>%
  mutate(sum_receiver = ave(TF_dif_binary, receiver, FUN = function(x) sum(x))) %>%
  mutate(sum_signal = ave(TF_dif_binary, signal, FUN = function(x) sum(x))) %>%
  mutate(sum_receiver_size = ave(TF_dif_binary, receiver, FUN = function(x) sum(abs(x)))) %>%
  mutate(sum_signal_size = ave(TF_dif_binary, signal, FUN = function(x) sum(abs(x))))


interaction_matrix <- interaction_df %>%
  distinct(TF_dif_binary, receiver, signal) %>%
  spread(key = "receiver", value = "TF_dif_binary") %>%
  column_to_rownames("signal")

pheatmap(interaction_matrix,
         border_color = NA,
         color = colorRampPalette(c("#d5bdaf", 'white', "#84a98c"))(100),
         clustering_method = "ward.D")


###
interaction_nodes <- interaction_df %>%
  filter(signal == receiver) %>%
  distinct(receiver, signal, sum_signal, sum_receiver, sum_signal_size, sum_receiver_size) %>%
  mutate(sum_total = sum_receiver_size + sum_signal_size) %>%
  mutate(class = ifelse(sum_signal <= 0, "repressor", "activator")) %>%
  distinct(receiver, sum_total, class)


interaction_edges <- interaction_df %>%
  filter(receiver != signal) %>%
  distinct(receiver, signal, TF_dif_binary, TF_dif) %>%
  mutate(weight = 0) %>%
  mutate(weight = ifelse(TF_dif_binary == -1, 1, weight)) %>%
  mutate(weight = ifelse(TF_dif_binary == 1, 2, weight)) %>%
  mutate(class = ifelse(weight == 1, "repressor", "activator")) %>%
  dplyr::select(signal, receiver, weight, class, TF_dif) %>%
  filter(weight != 0) %>%
  mutate(TF_dif = abs(TF_dif))

interaction_nodes <- interaction_nodes %>%
  filter(receiver %in% c(unique(interaction_edges$receiver),unique(interaction_edges$signal)))

routes_igraph <- graph_from_data_frame(d = interaction_edges, vertices = interaction_nodes, directed = TRUE)


visIgraph(routes_igraph) %>%
    visNodes(size = 25, shape = "circle") %>%
    visOptions(highlightNearest = list(enabled = TRUE, hover = TRUE),
               nodesIdSelection = TRUE) %>%
    visInteraction(keyboard = TRUE)

ggraph(routes_igraph, layout = "sugiyama") +
  geom_node_point(aes(size = sum_total, color = class)) +
  geom_edge_link(aes(color = class, width = TF_dif),
                 alpha = 0.8,
                 arrow = arrow(length = unit(3, 'mm')),
                 start_cap = circle(2, 'mm'),
                 end_cap = circle(3, 'mm')) +
  scale_edge_width(range = c(0.2, 2)) +
  scale_radius(limits = c(1, NA), range = c(1, 6)) +
  geom_node_text(aes(label = name), repel = TRUE) +
  scale_edge_color_manual(values = c("activator" = colors_diverse[2], "repressor" = colors_diverse[1])) +
  scale_color_manual(values = c("activator" = colors_diverse[2], "repressor" = colors_diverse[1])) +
  theme_graph()

```




## Cistrome TF targets
```{r}
# bc_files = list.files('/DATA/usr/m.trauernicht/projects/SuRE-TF/data/gcf6952_KD_HepG2/cistrome/',
#                        full.names=T, patter='*.txt')
# bc_list <- lapply(bc_files, fread, header = T)
# names(bc_list)<- gsub('.*/cistrome//(.*?).txt', '\\1', bc_files)
# 
# cistrome_df <- bind_rows(bc_list, .id = "TF") %>%
#   distinct(score, symbol, TF)
# 
# cistrome_df <- cistrome_df %>%
#   filter(TF %in% interaction_df$receiver, symbol %in% interaction_df$receiver) %>%
#   mutate(score = ave(score, TF, symbol, FUN = mean)) %>%
#   distinct() %>%
#   filter(TF != symbol) %>%
#   filter(score > 2) %>%
#   mutate(sum_target = as.numeric(ave(symbol, symbol, FUN = function(x) length(x))))
# 
# cistrome_matrix <- cistrome_df %>%
#   dplyr::select(-sum_target) %>%
#   spread(key = "symbol", value = "score") %>%
#   column_to_rownames("TF")
# 
# cistrome_matrix[is.na(cistrome_matrix)] <- 0
# 
# pheatmap(cistrome_matrix,
#          border_color = NA,
#          color = colorRampPalette(c("#d5bdaf", 'white', "#84a98c"))(100),
#          clustering_method = "ward.D")
# 
# cistrome_nodes <- cistrome_df %>%
#   distinct(symbol, sum_target)
# 
# cistrome_edges <- cistrome_df %>%
#   filter(symbol %in% cistrome_nodes$symbol) %>%
#   distinct(TF, symbol, "weight" = score)
# 
# cistrome_igraph <- graph_from_data_frame(d = cistrome_edges, vertices = cistrome_nodes, directed = TRUE)
# 
# ggraph(cistrome_igraph, layout = "sugiyama") + 
#   geom_node_point(aes(size = sum_target)) +
#   geom_edge_link(aes(width = weight), 
#                  alpha = 0.8,
#                  arrow = arrow(length = unit(3, 'mm')), 
#                  start_cap = circle(2, 'mm'),
#                  end_cap = circle(3, 'mm')) + 
#   scale_edge_width(range = c(0.2, 2)) +
#   geom_node_text(aes(label = name), repel = TRUE) +
#   theme_graph()
```


## Cistrome TF targets
```{r}
# bc_files = list.files('/DATA/usr/m.trauernicht/projects/SuRE-TF/data/gcf6952_KD_HepG2/cistrome/',
#                        full.names=T, patter='*.csv')
# bc_list <- lapply(bc_files, fread, header = T)
# names(bc_list)<- gsub('.*/cistrome//.*M_.*_(.*?)hg38.*.csv', '\\1', bc_files)
# 
# cistrome_df <- bind_rows(bc_list, .id = "target") %>%
#   distinct(RP_score, "TF" = Factor, target)
# 
# cistrome_df <- cistrome_df %>%
#   filter(TF %in% interaction_df$receiver, target %in% interaction_df$receiver) %>%
#   mutate(RP_score = ave(RP_score, TF, target, FUN = mean)) %>%
#   distinct() %>%
#   filter(TF != target) %>%
#   #filter(RP_score > 0.3) %>%
#   mutate(sum_target = as.numeric(ave(target, target, FUN = function(x) length(x))))
# 
# cistrome_matrix <- cistrome_df %>%
#   dplyr::select(-sum_target) %>%
#   spread(key = "target", value = "RP_score") %>%
#   column_to_rownames("TF")
# 
# cistrome_matrix[is.na(cistrome_matrix)] <- 0
# 
# pheatmap(cistrome_matrix,
#          border_color = NA,
#          color = colorRampPalette(c("#d5bdaf", 'white', "#84a98c"))(100),
#          clustering_method = "ward.D")
# 
# cistrome_nodes <- cistrome_df %>%
#   distinct(target, sum_target)
# 
# cistrome_edges <- cistrome_df %>%
#   filter(target %in% cistrome_nodes$target) %>%
#   distinct(TF, target, "weight" = RP_score)
# 
# cistrome_igraph <- graph_from_data_frame(d = cistrome_edges, vertices = cistrome_nodes, directed = TRUE)
# 
# ggraph(cistrome_igraph, layout = "sugiyama") + 
#   geom_node_point(aes(size = sum_target)) +
#   geom_edge_link(aes(width = weight), 
#                  alpha = 0.8,
#                  arrow = arrow(length = unit(3, 'mm')), 
#                  start_cap = circle(2, 'mm'),
#                  end_cap = circle(3, 'mm')) + 
#   scale_edge_width(range = c(0.2, 2)) +
#   geom_node_text(aes(label = name), repel = TRUE) +
#   theme_graph()
```



## Reporter activity compared to background

Aim: Display reporter activities relative to background activity per tf and condition. Get an overview over all reporter activities per condition.

```{r reporter_activities, out.width= "80%", fig.align= "center"}
nt_df <- cDNA_df %>%
  dplyr::filter(condition %in% c("NT_1", "NT_2")) %>%
  dplyr::select(reporter_id, "nt_reporter_activity_minP" = reporter_activity_minP) %>%
  mutate(nt_reporter_activity_minP = ave(nt_reporter_activity_minP, reporter_id, FUN = mean)) %>%
  distinct()

cDNA_df2 <- merge(cDNA_df, nt_df, all = T) %>%
  #na.omit() %>%
  filter(neg_ctrls == "No") %>%
  mutate(reporter_FC = log2(reporter_activity_minP/nt_reporter_activity_minP)) %>%
  distinct(tf, reporter_id, condition, reporter_FC, commercial_reporter) %>%
  mutate(mean_tf_activity = ave(reporter_FC, tf, condition, FUN = mean))

## Reporter FC vs reporter activity in NT control - are the most changing reporters the ones with higher activity?
ggplot(merge(cDNA_df, nt_df, all = T) %>%
  filter(neg_ctrls == "No", tf == "SP1", condition == "SP1") %>%
  mutate(reporter_FC = log2(reporter_activity_minP/nt_reporter_activity_minP)) %>%
  distinct(tf, reporter_id, condition, reporter_FC, commercial_reporter, nt_reporter_activity_minP),
       aes(x = reporter_FC, y = nt_reporter_activity_minP)) +
  geom_point()

tf_kd_targets <- read.csv2("/DATA/usr/m.trauernicht/projects/SuRE-TF/data/gcf7027_KD_mES/tf_kd_targets_mES.csv") %>%
  #filter(condition %in% cDNA_df5$condition) %>%
  setnames("target", "tf") %>%
  mutate(is_target = T) %>%
  right_join(cDNA_df2, by = c("condition", "tf"))

tf_kd_targets$is_target[is.na(tf_kd_targets$is_target)] <- F
tf_kd_targets <- tf_kd_targets %>%
  filter(condition != "NT") %>%
  arrange(is_target)

# for (i in unique(tf_kd_targets$condition)) {
#   print(i)
#   p <- ggplot(tf_kd_targets %>%
#                 filter(condition == i),
#               mapping = aes(x = reorder(tf, -mean_tf_activity), y = reporter_FC, color = is_target)) +
#     geom_boxplot(aes(alpha = 0.7)) +
#     geom_quasirandom(aes(alpha = 0.7)) +
#     scale_color_manual(values = c(colors_diverse[1], colors_diverse[3])) +
#     ggtitle(i) +
#     theme_classic_lines_90()
#
#   print(p)
# }

tf_kd_targets <- tf_kd_targets %>%
  mutate(tf = gsub("_.*", "", tf))

tf_kd_targets2 <- tf_kd_targets %>%
  filter(tf %in% cDNA_df4$tf) %>%
  mutate(is_target = ifelse(commercial_reporter == "Yes", "commercial", is_target))

for (i in unique(tf_kd_targets2$condition)) {
  print(i)
  p <- ggplot(tf_kd_targets2 %>%
                filter(condition == i),
              mapping = aes(x = reorder(tf, -mean_tf_activity), y = reporter_FC, color = is_target, group = commercial_reporter)) +
    geom_hline(yintercept = 0, lty = 2) +
    geom_quasirandom(dodge.width = .75) +
    scale_color_manual(values = c("FALSE" = colors_diverse[1], "TRUE" = colors_diverse[3], "commercial" = colors_diverse[4])) +
    ggtitle(i) +
    theme_pubr(border = T, x.text.angle = 90)

  print(p)
}

for (i in unique(tf_kd_targets2$tf)) {
  print(i)
  p <- ggplot(tf_kd_targets2 %>%
                filter(tf == i),
              mapping = aes(x = reorder(condition, -mean_tf_activity), y = reporter_FC, color = is_target, group = commercial_reporter)) +
    geom_hline(yintercept = 0, lty = 2) +
    geom_quasirandom(dodge.width = .75, raster.dpi = 600) +
    scale_color_manual(values = c("FALSE" = colors_diverse[1], "TRUE" = colors_diverse[3], "commercial" = colors_diverse[4])) +
    ggtitle(i) +
    theme_pubr(border = T, x.text.angle = 90)

  print(p)
}


```

Conclusion: Some KDs work well, some synthetic reporters are better than the positive controls.

---







# Session Info
```{r}
paste("Run time: ",format(Sys.time()-StartTime))
getwd()
date()
sessionInfo()
```

