---
title: "ATACseq QC"
author: 
  - name: "Teun van den Brand"
    email: "t.vd.brand@nki.nl"
    affiliation: "Netherlands Cancer Institute"
date: '`r format(Sys.time(), "%d/%m/%Y")`'
output: 
  html_document:
    theme: united
    highlight: pygments
    fig_caption: yes
    code_folding: hide
    df_print: kable
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
bibliography: /DATA/users/t.vd.brand/bibliography/articles.bibtex
---

## Setup {.tabset}

<!-- little HTML script to do indentation of the table of contents -->
<script>
    $(document).ready(function() {
      $items = $('div#TOC li');
      $items.each(function(idx) {
        num_ul = $(this).parentsUntil('#TOC').length;
        $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
      });

    });
</script>

```{css, echo = FALSE}
div.sourceCode {
  overflow-x: hidden;
}
```

### Load packages

```{r load_packages, results = 'hide', warning = FALSE, message = FALSE}
library(DESeq2)
library(SummarizedExperiment)
library(GenomicRanges)

library(here)
library(ggplot2)
library(data.table)
library(scales)
library(ggh4x)

library(TxDb.Mmusculus.UCSC.mm10.knownGene)
```

### Load data

```{r load_data}
exp <- readRDS(here("rds", "experiment_timeseries.rds"))
```

### Setup Aesthetics

```{r setup_aes}
mycolour <- "#000000FF" # Opaque Black

theme_set(theme_gray())
theme_update(text = element_text(colour = mycolour),
             line = element_line(colour = mycolour),
             aspect.ratio = 1,
             axis.line  = element_line(colour = mycolour),
             axis.ticks = element_line(colour = mycolour),
             axis.text  = element_text(colour = mycolour),
             legend.key = element_blank(),
             panel.background = element_blank(),
             panel.grid.major = element_line(colour = "grey95"),
             panel.grid.minor = element_blank(),
             plot.background  = element_blank(),
             strip.background = element_blank(),
             strip.text = element_text(colour = mycolour))

rm(mycolour)

scale_colour_dtag <- function(...) {
  scale_colour_manual(
    name = "dTAG",
    values = c(
      "SOX2"  = "#e41a1c",
      "OCT4"  = "#377eb8",
      "NANOG" = "#4daf4a"
    ),
    ...
  )
}

scale_fill_dtag <- function(...) {
  scale_fill_manual(
    name = "dTAG",
    values = c(
      "SOX2"  = "#e41a1c",
      "OCT4"  = "#377eb8",
      "NANOG" = "#4daf4a"
    ),
    ...
  )
}

fill_shape_guide <- function(...) {
  guide_legend(override.aes = list(shape = 21), ...)
}
```

## Aim

Investigate some QC metrics for the timeseries ATACseq data.

## Data wrangling

Loop through the tabix files and calculate some metrics.

```{r}
peaks <- as(rowRanges(exp), "GNCList")

tss_width <- 100
flank_width <- 2000

tss <- resize(genes(TxDb.Mmusculus.UCSC.mm10.knownGene), 1)
tss_target <- resize(tss, tss_width, fix = "center")
tss_flank  <- c(
  flank(tss_target, flank_width / 2, start = TRUE),
  flank(tss_target, flank_width / 2, start = FALSE)
)
tss_target <- as(tss_target, "GNCList")
tss_flank  <- as(tss_flank, "GNCList")

qcstats <- lapply(seq_len(ncol(exp)), function(i) {
  dat <- fread(exp$tabix_file[[i]])
  dat <- with(dat, GRanges(V1, IRanges(V2, V3)))
  
  dat <- keepStandardChromosomes(dat, "Mus_musculus", "coarse")
  
  # Mitochondrial
  mito <- sum(seqnames(dat) == "chrM")
  dat  <- dat[seqnames(dat) != "chrM"]
  
  # Calculate fraction of reads in peaks
  frip <- mean(overlapsAny(dat, peaks))
  
  # Calculate TSS score
  target_tss   <- sum(overlapsAny(dat, tss_target)) / tss_width
  target_flank <- sum(overlapsAny(dat, tss_flank)) / flank_width
  
  # Calculate fragment length distribution
  fraglen <- table(width(dat))
  fraglen <- fraglen / sum(fraglen)
  
  DataFrame(
    id = i,
    useable_reads = length(dat),
    mitochondrial_reads = mito,
    frip = frip,
    tss_target = target_tss,
    tss_flank  = target_flank,
    fraglen = SimpleList(fraglen)
  )
})

qcstats <- do.call(rbind, qcstats)
```

## Statistics

### Read duplication

This is the percent duplication metric found in the Picard MarkDuplicates logs.

```{r}
df <- data.frame(
  run  = exp$run,
  tag  = exp$tag,
  time = exp$time,
  rep  = exp$rep,
  dup  = exp$logs$percent_duplication
)

ggplot(df, aes(interaction(run, time), dup)) +
  geom_point(aes(shape = time, fill = tag),
             size = 3) +
  facet_grid(~ tag, scales = "free_x", space = "free_x") +
  scale_y_continuous(labels = percent_format(1),
                     name = "Duplicated Reads",
                     limits = c(0, NA),
                     expand = c(0,0,0.1,0)) +
  scale_x_discrete(name = NULL) +
  scale_shape_manual(
    name = "Time",
    values = c(25, 21:24),
  ) +
  scale_fill_dtag(guide = fill_shape_guide()) +
  guides(x = guide_axis_nested()) +
  theme(aspect.ratio = NULL,
        axis.text.x = element_text(angle = 90, vjust = 0.5,
                                   margin = margin(t = 2.2, b = 2.2)))
```

### Mitochondrial reads

Counts of reads originating from 'chrM' versus not originating from 'chrM'.

```{r}
df <- data.frame(
  run  = exp$run,
  tag  = exp$tag,
  time = exp$time,
  rep  = exp$rep,
  mito = qcstats$mitochondrial_reads,
  useable = qcstats$useable_reads
)
df$perc <- df$mito / (df$mito + df$useable)

df2 <- tidyr::pivot_longer(df, c(mito, useable))

ggplot(df2, aes(y = interaction(run, time))) +
  geom_col(aes(x = value, fill = factor(name, c("useable", "mito"))), 
           width = 0.8) +
  facet_grid(tag ~ ., scales = "free_y", space = "free_y") +
  geom_text(
    data = df,
    aes(x = 0, label = percent(perc, 0.1)),
    size = 3, hjust = 1, nudge_x = -1e6
  ) +
  scale_fill_manual(
    name = "Read Origin",
    values = c("grey50", "#E41A1C"),
    labels = c("Nuclear", "Mitochondrial")
  ) +
  scale_x_continuous(
    labels = number_format(scale = 1e-6, suffix = "M"),
    name = "Reads",
    expand = c(0.07,0,0.1,0)
  ) +
  scale_y_discrete(name = NULL) +
  guides(y = guide_axis_nested()) +
  theme(aspect.ratio = NULL,
        axis.text.y = element_text(margin = margin(l = 2.2, r = 2.2)),
        legend.position = c(1,0),
        legend.justification = c(1, 0),
        legend.background = element_blank())
```

### TSS enrichment

Number of nuclear reads overlapping `r tss_width` bp around TSSs divided by number of nuclear reads overlapping `r flank_width` bp flanking those sites. Typical recommendation is that this metric is >4.

```{r}
df <- data.frame(
  run  = exp$run,
  tag  = exp$tag,
  time = exp$time,
  rep  = exp$rep,
  tss  = qcstats$tss_target / qcstats$tss_flank
)

ggplot(df, aes(interaction(run, time), tss)) +
  geom_point(aes(shape = time, fill = tag),
             size = 3) +
  geom_hline(yintercept = 4, colour = "purple", linetype = 2) +
  facet_grid(~ tag, scales = "free_x", space = "free_x") +
  scale_y_continuous(name = "Fold Enrichment",
                     limits = c(4, NA)) +
  scale_x_discrete(name = NULL) +
  scale_shape_manual(
    name = "Time",
    values = c(25, 21:24),
  ) +
  scale_fill_dtag(guide = fill_shape_guide()) +
  guides(x = guide_axis_nested()) +
  theme(aspect.ratio = NULL,
        axis.text.x = element_text(angle = 90, vjust = 0.5,
                                   margin = margin(t = 2.2, b = 2.2)))
```

### Fraction of Reads in Peaks

Number of nuclear reads that have any overlap with a peak as fraction of the total number of nuclear reads.

```{r}
df <- data.frame(
  run  = exp$run,
  tag  = exp$tag,
  time = exp$time,
  rep  = exp$rep,
  frip = qcstats$frip
)

ggplot(df, aes(interaction(run, time), frip)) +
  geom_point(aes(shape = time, fill = tag),
             size = 3) +
  facet_grid(~ tag, scales = "free_x", space = "free_x") +
  scale_y_continuous(name = "Fraction of Reads in Peaks",
                     limits = c(0, NA),
                     expand = c(0,0,0.1,0)) +
  scale_x_discrete(name = NULL) +
  scale_shape_manual(
    name = "Time",
    values = c(25, 21:24),
  ) +
  scale_fill_dtag(guide = fill_shape_guide()) +
  guides(x = guide_axis_nested()) +
  theme(aspect.ratio = NULL,
        axis.text.x = element_text(angle = 90, vjust = 0.5,
                                   margin = margin(t = 2.2, b = 2.2)))
```

### Fragment length distribution

Count for every unique value of nuclear read start in bp minus the read end in bp how often they occur and divide that by the total number of nuclear reads.

```{r}
fraglen <- lapply(qcstats$fraglen, as.data.frame)
fraglen <- rbindlist(fraglen, idcol = "id")
fraglen <- transform(
  fraglen,
  Var1 = as.numeric(as.character(fraglen$Var1)),
  tag  = exp$tag[id],
  time = exp$time[id],
  rep  = exp$rep[id],
  run  = exp$run[id]
)

ggplot(fraglen, aes(Var1, Freq * 1000)) +
  geom_area(aes(fill = tag, colour = tag),
            size = 0.1) +
  scale_x_continuous(
    name = "Fragment Length",
    limits = c(0, 600), oob = oob_keep,
    expand = c(0, 0),
    breaks = c(0, 200, 400)
  ) +
  scale_y_continuous(
    name = "Relative Frequency (x 1000)",
    limits = c(0, 0.01) * 1000, oob = oob_keep,
    expand = c(0, 0, 0.1, 0),
    breaks = c(0, 5, 10)
  ) +
  scale_fill_dtag(aesthetics = c("colour", "fill")) +
  facet_nested(time ~ tag + run, nest_line = TRUE) +
  theme(aspect.ratio = 2 / (1 + sqrt(5)),
        legend.position = "none")
```

## References

<div id="refs"></div>

## Session Info

<p>
  <a class="btn btn-primary" data-toggle="collapse" href="#sinfo" role="button" aria-expanded="false" aria-controls="sinfo">
    View
  </a>
</p>
<div class="collapse" id="sinfo">
  <div class="card card-body">

```{r session_info, include=TRUE, echo=FALSE, results = "asis"}
pkgs <- loadedNamespaces()
libTvdB::format_sessioninfo_html(pkgs)
```

  </div>
</div>
