---
title: "Motif analysis of the chosen TFs"
author: 
  - name: "Max Trauernicht"
    email: "m.trauernicht@nki.nl"
    affiliation: "Netherlands Cancer Institute - van Steensel lab"

date: '`r format(Sys.time(), "%d/%m/%Y")`'
output:
  html_document:
    theme: united
    highlight: pygments
    fig_caption: yes
    code_folding: hide
    df_print: kable
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
---

*knitr document van Steensel lab*

## Introduction
86 TFs were chosen to desing TF reporters for. In this analysis, I will analyse the motifs of the chosen TFs. The goal is to find out how unique the motifs are and how well they represent the motif landscape.

## Setup {.tabset}

<!-- little HTML script to do indentation of the table of contents -->
<script>
    $(document).ready(function() {
      $items = $('div#TOC li');
      $items.each(function(idx) {
        num_ul = $(this).parentsUntil('#TOC').length;
        $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
      });

    });
</script>

```{css, echo = FALSE}
div.sourceCode {
  overflow-x: hidden;
}
```

### Libraries
```{r setup, out.width= "100%", fig.align= "center", echo=FALSE, warning= FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
StartTime <-Sys.time()

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8) 
# libraries:
library(ggplot2)
library(seqinr)
library(seqLogo)
library(universalmotif)
library(Biostrings)
library(SimRAD)
library(gtools)
library(DNABarcodes)
library(phylotools)
library(ape)
library(magrittr)
library(dplyr)
library(readr)
library(stringr)
library(tidyr)
library(heatmaply)
library(pheatmap)
library(tibble)
library(ggseqlogo)
library(RColorBrewer)
library(data.table)
library(vwr)
library(ggbeeswarm)
library(ggrepel)
library(sjmisc)
library(RcmdrMisc)
library(LncFinder)
library(ENCODExplorer)
library(biomaRt)
library(ggpubr)
library(MotifDb)
library(umap)
```


### Collect motifs
```{r out.width= "80%", fig.align= "center", echo = FALSE, warning = FALSE, message=FALSE}
# Load Lambert et al. cluster information
tf_clusters <- read.csv2("/DATA/usr/m.trauernicht/projects/SuRE-TF/data/library_design/lambert_2018/lambert_tf_clusters.csv", header = T) %>%
  dplyr::select("motif_id" = CIS.BP.ID, "cluster" = Cluster.., "family" = TF.Family) %>%
  mutate(motif_id = paste(motif_id, ".txt", sep = ""))

lambert_all_motifs_df <- read.csv2("/DATA/usr/m.trauernicht/projects/SuRE-TF/data/library_design/lambert_2018/HumanTFs_Lambert_curated.csv", header = T)

lambert_all_motifs <- lambert_all_motifs_df %>%
  filter(best_motif == T) %>%
  mutate(CIS.BP_ID = paste(CIS.BP_ID, ".txt", sep = "")) %>%
  mutate(tf = gsub("(.*?)_.*", "\\1", motif_ID)) %>%
  mutate(tf = gsub("V\\$","", tf)) %>%
  mutate(tf = gsub("PBMC-", "", tf)) %>%
  dplyr::select("motif_id" = CIS.BP_ID, tf) %>%
  distinct() %>%
  group_by(tf) %>%
  slice_head(n = 1) ## For some TFs multiple motifs were selected as best - I will just keep one of those (randomly)

# Load motifs from library 1
gen1_motifs <- read.csv2("/DATA/usr/m.trauernicht/projects/SuRE-TF/data/library_design/parameter_files/TF_motifs_annotated.csv", header = T) %>%
  dplyr::select("tf" = TF, "motif_id" = `Source..CIS.BP...build.1.94.`) %>%
  mutate(tf = toupper(tf), motif_id = paste(motif_id, ".txt", sep = "")) %>%
  mutate(tf = ifelse(tf == "TRP53", "TP53", tf)) %>%
  mutate(tf = ifelse(tf == "TCFCP2L1", "TFCP2L1", tf)) 

gen1_motifs$tf[gen1_motifs$tf == "NR3C1(GR)"] <- "NR3C1"

# Load motifs from library 2
gen2_motifs <- read.csv("/DATA/usr/m.trauernicht/projects/SuRE-TF/data/library_design/lambert_2018/motifs_gen2.csv", header = T) %>%
  dplyr::select("tf" = TF, motif_id) %>%
  mutate()

# Combine all motifs
all_motifs <- rbind(lambert_all_motifs, gen1_motifs, gen2_motifs) %>%
  distinct() %>%
  mutate(motif_id = ifelse(tf == "SOX9", "M06107_1.94d.txt", motif_id)) %>%
  mutate(motif_id = ifelse(tf == "SOX2", "M06121_1.94d.txt", motif_id)) %>%
  mutate(selected = ifelse(motif_id %in% gen1_motifs$motif_id | motif_id %in% gen2_motifs$motif_id, "Yes", "No")) %>%
  filter(!motif_id %in% c("M10257_1.94d.txt", "M10277_1.94d.txt", "M10282_1.94d.txt", "M10303_1.94d.txt", 
                          "M00897_1.94d.txt", "M08847_1.94d.txt", "M01568_1.94d.txt", "M01343_1.94d.txt",
                          "M01775_1.94d.txt", "M01230_1.94d.txt", "M00877_1.94d.txt", "M01240_1.94d.txt",
                          "M01025_1.94d.txt", "M01278_1.94d.txt", "M01428_1.94d.txt", "M00497_1.94d.txt",
                          "M02157_1.94d.txt")) 
```


### Check which of those motifs are relevant in cancer

```{r}
# Load all lambert TFs
lambert_tfs <- read_csv("/DATA/usr/m.trauernicht/projects/SuRE-TF/data/lambert_all_tfs.csv") %>%
  filter(`...4` == "Yes") %>%
  dplyr::select("tf" = Name) %>%
  mutate(tf = toupper(tf)) %>%
  distinct()

# Load cancer relevant TFs
cosmic_cancer <- read.csv("/DATA/usr/m.trauernicht/projects/SuRE-TF/data/Census_all.csv", header = T) %>%
  dplyr::select("tf" = Gene.Symbol) %>%
  mutate(tf = toupper(tf)) %>%
  mutate(cancer = "Yes")

# Load prime TF reporters
prime_reporters <- read.csv("/DATA/usr/m.trauernicht/projects/SuRE-TF/data/best_reporters.csv", header = T) %>%
  filter(conf_level > 1) %>%
  mutate(tf = gsub("::RXRA", "", tf)) %>%
  mutate(tf = gsub("AHR::ARNT", "AHR", tf)) %>%
  mutate(tf = gsub("MAF::NFE2", "MAF", tf)) %>%
  mutate(tf = gsub("FOS::JUN", "FOS", tf)) %>% 
  mutate(tf = gsub("SMAD2::SMAD3::SMAD4", "SMAD2", tf)) %>%
  distinct(tf) %>%
  mutate(prime = "Yes")

# Load all selected tfs
all_tfs <- read.csv("/DATA/usr/m.trauernicht/projects/SuRE-TF/data/best_reporters.csv", header = T) %>%
  mutate(tf = gsub("::RXRA", "", tf)) %>%
  mutate(tf = gsub("AHR::ARNT", "AHR", tf)) %>%
  mutate(tf = gsub("MAF::NFE2", "MAF", tf)) %>%
  mutate(tf = gsub("FOS::JUN", "FOS", tf)) %>% 
  mutate(tf = gsub("SMAD2::SMAD3::SMAD4", "SMAD2", tf)) %>%
  distinct(tf) %>%
  mutate(selected = "Yes")

# Match with all motifs
all_motifs_cancer <- lambert_tfs %>%
  #mutate(tf = ifelse(tf == "AHR:ARNT", "AHR::ARNT", tf)) %>%
  #mutate(tf = ifelse(tf == "RAR:RXR", "RARA:RXRA", tf)) %>%
  #mutate(tf = ifelse(tf == "HIF1", "HIF1A", tf)) %>%
  #mutate(tf = toupper(tf)) %>%
  left_join(cosmic_cancer, by = "tf") %>%
  left_join(prime_reporters, by = "tf") %>%
  left_join(all_tfs, by = "tf") %>%
  mutate(cancer = ifelse(is.na(cancer), "No", "Yes")) %>%
  mutate(prime = ifelse(is.na(prime), "No", "Yes")) %>%
  mutate(selected = ifelse(is.na(selected), "No", "Yes")) %>%
  dplyr::select("tf", "cancer", "prime", "selected") %>%
  distinct()

```


### Import all motifs as PWMs
```{r warning = FALSE, message=FALSE}
seq_logos <- list()
for(i in 1:nrow(all_motifs)) {
  
  x <- all_motifs$motif_id[i]
  file_x <- paste("/DATA/usr/m.trauernicht/projects/SuRE-TF/data/library_design/lambert_2018/PWMs/", x, sep = "")
  
  if (file.exists(file_x) == T) {
    
    if (str_contains(x, "txt") == TRUE) {
    pwm <- read.table(file_x, header = T) %>% dplyr::select(-Pos) %>% t()
    seq_logos[i] <- list(pwm)
    }
    
    if (str_contains(x, "pcm") == TRUE)  {
    pwm <- read.table(file_x, header = F) %>% column_to_rownames("V1")
    pwm <- as.matrix(pwm)
    seq_logos[i] <- list(pwm)
    }
    
    if (str_contains(x, "jaspar") == TRUE) {
    pwm <- read_jaspar(file_x)
    pwm <- pwm@motif
    # Transform to relative values
    pwm <- colPercents(pwm)/100
    pwm <- pwm[1:4,]
    seq_logos[i] <- list(pwm)
    }
  }
}

# Rename pwms
names(seq_logos) <- all_motifs$motif_id

# Remove motifs for which pwms were not available
seq_logos <- seq_logos[!sapply(seq_logos,is.null)]

# Convert to universalmotif format
seq_logos_motif <- list()
for (i in names(seq_logos)) {
  seq_logos_motif[[i]] <- universalmotif::convert_motifs(seq_logos[[i]])
  seq_logos_motif[[i]]@name <- i
}

saveRDS(seq_logos, "/DATA/usr/m.trauernicht/projects/SuRE-TF/data/library_design/lambert_2018/seq_logos.rds")


# ## Alternative motif collection
# motifs <- convert_motifs(MotifDb)
# motifs <- filter_motifs(motifs,organism="Hsapiens")
```


### Compute pairwise similarities
```{r warning=FALSE, message=FALSE}
# Compute similarities using universalmotif function
motif_sim <- compare_motifs(seq_logos_motif, method = "PCC") ## Methods that work well: ALLR, ALLR_LL, PCC
```


## Fig 1A: UMAP visualization of chosen motifs within motif clusters
Aim: Show diversity of chosen motifs.
```{r warning=FALSE}
### UMAP analysis
set.seed(35425) 
motif_cor <- motif_sim %>%
  replace(is.na(.), 0)

motifs_selected <- rbind(gen1_motifs, gen2_motifs) %>%
  mutate(motif_id = ifelse(tf == "SOX9", "M06107_1.94d.txt", motif_id)) %>%
  mutate(motif_id = ifelse(tf == "SOX2", "M06121_1.94d.txt", motif_id))
saveRDS(motifs_selected, "/DATA/usr/m.trauernicht/projects/SuRE-TF/data/library_design/lambert_2018/motifs_selected.rds")

tf_umap <- umap(motif_cor) 
colnames(tf_umap$layout) <- c("A","B") 
tf_umap_plot <- tf_umap$layout %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "motif_id") %>% 
  left_join(all_motifs) %>%
  left_join(tf_clusters) %>%
  #left_join(lambert_clusters) %>%
  mutate(selected = ifelse(tf %in% motifs_selected$tf, "Yes", "No")) %>%
  mutate(cluster2 = ifelse(cluster <= 10, cluster, ">10")) %>%
  mutate(family2 = ifelse(family %in% c("bHLH", "bZIP", "Ets", "Forkhead", "GATA", "Homeodomain", "Nuclear receptor",
                                        "T-box", "C2H2 ZF"), family, "other"))

#Plot UMAP
ggplot(,aes(x = A, y = B)) +
  stat_density2d(data = tf_umap_plot, geom = "polygon", aes(alpha = ..level..), bins = 5) +
  geom_point(data = tf_umap_plot %>% filter(selected == "No"), size = 1 , color = "grey") + 
  geom_point(data = tf_umap_plot %>% filter(selected == "Yes"), size = 4 , color = "#E07858") + 
  theme_pubr() +
  scale_alpha(range = c(0.05,0.15)) +
  geom_text_repel(data = tf_umap_plot %>% filter(selected == "Yes"), aes(label = tf), force = 86) + 
  ylab("UMAP2") + 
  xlab("UMAP1")

ggplot(,aes(x = A, y = B)) +
  geom_point(data = tf_umap_plot %>% filter(selected == "No"), size = .4 , color = "grey") + 
  geom_point(data = tf_umap_plot %>% filter(selected == "Yes"), size = 1, color = "#E07858") + 
  theme_pubr() +
  ylab("UMAP2") + 
  xlab("UMAP1")

x <- ggplot(,aes(x = A, y = B, label = tf)) +
  geom_point(data = tf_umap_plot %>% filter(selected == "No"), size = .4 , color = "grey") + 
  geom_point(data = tf_umap_plot %>% filter(selected == "Yes"), size = 1, color = "#E07858") + 
  theme_pubr() +
  ylab("UMAP2") + 
  xlab("UMAP1")

ggplotly(x)

ggplot(,aes(x = A, y = B)) +
  stat_density2d(data = tf_umap_plot, geom = "polygon", aes(alpha = ..level..), bins = 5) +
  geom_point(data = tf_umap_plot %>% filter(selected == "No"), size = 1 , aes(color = family2)) + 
  geom_point(data = tf_umap_plot %>% filter(selected == "Yes"), size = 4 , color = "#E07858") + 
  theme_pubr() +
  scale_alpha(range = c(0.05,0.15)) +
  geom_text_repel(data = tf_umap_plot %>% filter(selected == "Yes"), aes(label = tf), force = 86) + 
  ylab("UMAP2") + 
  xlab("UMAP1")
```
---
Conclusion: Motifs represent large diversity within the complete motif collection. 


## Fig S1: Similarity heatmap of chosen motifs
```{r warning = F}
motif_similarity_df <- data.frame(motif_sim) %>%
  rownames_to_column("motif1") %>%
  pivot_longer(cols = -motif1, names_to = "motif2", values_to = "similarity") %>%
  mutate(similarity = as.numeric(similarity)) 

motif_similarity_df_selected <- motif_similarity_df %>%
  filter(motif1 %in% motifs_selected$motif_id & motif2 %in% motifs_selected$motif_id) %>%
  replace(is.na(.), 0)

motif_similarity_df_selected <- motif_similarity_df_selected %>%
  left_join(motifs_selected %>% dplyr::select("motif1" = motif_id, "tf1" = tf)) %>%
  left_join(motifs_selected %>% dplyr::select("motif2" = motif_id, "tf2" = tf)) %>%
  distinct(tf1, tf2, similarity)

motif_similarity_mat <- motif_similarity_df_selected %>%
  spread(key = "tf2", value = "similarity") %>%
  column_to_rownames("tf1")

ord <- hclust( dist(motif_similarity_mat, method = "euclidean"), method = "ward.D2" )$order

motif_similarity_df_selected <- motif_similarity_df_selected %>%
  mutate(tf1 = factor(tf1, levels = rownames(motif_similarity_mat)[ord]),
         tf2 = factor(tf2, levels = rownames(motif_similarity_mat)[ord]))

ggplot(motif_similarity_df_selected,
       aes(x = tf1, y = tf2, fill = similarity)) +
  geom_tile(size = .5) +
  coord_fixed() +
  theme_pubr(x.text.angle = 90, border = T) +
  scale_fill_gradient2(low = "#B7B7A4", mid = "white", high = "#E17B5C", midpoint = 0.5)

ggplot(motif_similarity_df_selected,
       aes(x = tf1, y = tf2, fill = similarity)) +
  geom_tile(size = .5) +
  coord_fixed() +
  theme_pubr(x.text.angle = 90, border = T) +
  scale_fill_gradient2(low = "white", high = "#E17B5C", midpoint = 0.4)
```
---
Conclusion: We have some TFs with overlapping motifs. However, those are mostly nuclear receptors that have specific stimulations.

# Session Info
```{r}
paste("Run time: ",format(Sys.time()-StartTime))
getwd()
date()
sessionInfo()
```

